### 2023-12-21 aanda.tex
Add files via upload

%                                                                 aa.dem
% AA vers. 9.1, LaTeX class for Astronomy & Astrophysics
% demonstration file
%                                                       (c) EDP Sciences
%-----------------------------------------------------------------------
%
%\documentclass[referee]{aa} % for a referee version
%\documentclass[onecolumn]{aa} % for a paper on 1 column  
%\documentclass[longauth]{aa} % for the long lists of affiliations 
%\documentclass[letter]{aa} % for the letters 
%\documentclass[bibyear]{aa} % if the references are not structured 
%                              according to the author-year natbib style

%
\documentclass{aa}

%
\usepackage{graphicx}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage{txfonts}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\usepackage[options]{hyperref}
% To add links in your PDF file, use the package "hyperref"
% with options according to your LaTeX or PDFLaTeX drivers.
%
\begin{document}

   \title{The periodicity of Betelgeuse explained by surface convection}

    \author{{ Q.~Pilate}\inst{1},{ A.~L{\'o}pez Ariste}\inst{1},{ A.~Lavail}\inst{1},{ Ph. Mathias}\inst{1} }

   \institute{IRAP, Universit\'e de Toulouse, CNRS, CNES, UPS.  14, Av. E. Belin. 31400 Toulouse, France 
             }

   \date{Received ...; accepted ...}

% \abstract{}{}{}{}{} 
% 5 {} token are mandatory

  \abstract

   \keywords{
               }

   \maketitle
%
%-------------------------------------------------------------------

\section{Introduction}
Betelgeuse is a red super giant star which has been observed by many astronomers for centuries. This star is known as a semi variable star. In the literature, the variability of the light curve of Betelgeuse is mainly explained by pulsation of the atmosphere. The most common periods found in Betelgeuse are the long secondary period (LSP) which is around $2000$ days and there are few known overtones such as $400$ days or $200$ days. Since the great dimming between the end of 2019 and beginning of 2020, those periods have changed. In this article, we attempt to recover the different periods of Betelgeuse using surface convection. The brightness of Betelgeuse is linked to the giant convective cells present at the surface. Hence, the light curve will be affected by the size and the number of convective cells at the surface. Betelgeuse has been observed almost weekly by the telescope Bernard Lyot at Pic du Midi in France, over a decade. We are also able to produce images of Betelgeuse using linear polarization. Those two points combined allowed us to produce a movie of the surface of Betelgeuse over the last ten years. Using this movie, we try to recover the periodicity of the light curve of Betelgeuse.

\section{Method}

Trying to find periods in an astrophysical context can be difficult due to the unevenly spaced observed data. To overcome this issue, we use the Lomb-Scargle (LS) periodogram. For a given set of data, it fits a periodic signal using least squares method, for each frequency between the first and the last observation. The better the periodic signal fits the data, the higher the power attributed to the frequency. We used the LS periodogram in the linear polarization and intensity data first. We can see in fig.\ref{LS Intensity} that we recover the LSP around 2000 days. For lower frequencies (i.e periods higher than 2000 days), the signal is attributed to cross correlation. It appears that there is a small signal around $0.003 d^{-1}$ ($\sim 333$ days) but it is hard to affirm that it corresponds a net signal or noise. The LS periodogram of stokes Q shows no LSP but rather shows two periods : one at 0.002 (500 days) and one at 0.005 (200 days). The second one may correspond to an overtone of the fundamental mode that has already been motioned in the literature but the one at 500 days has never been found. From the LS of stokes U, we recover the LSP but there is no clear signal for the other frequencies. Since the LSP is not present in stokes Q but only in stokes I and U, we attribute this frequency to stochastic events. To go further, we examine the LS periodogram of the displacement of the photo-center of Betelgeuse. From the polarimetric data of the TBL, we can produce images of Betelgeuse and follow its photo-center displacement, producing "movies" of Betelgeuse. However, our images are very dependant of the initial parameters to fit the observed profile. To overcome this issue, we computed movies of Betelgeuse 100 times, with a 100 initial different parameters.  Figure \ref{LS narval + neonarval} shows the LS of the data from narval and neo narval up to 2023 for the 100 movies, whereas fig.\ref{LS narval} shows the LS only for narval data. We see that by using only the narval data, we find two peaks in the LS periodogram: one at $\sim 330$ days and another one at $\sim 420$ days. However, those peaks are less clear with neo narval (which was added right after the great dimming of Betelgeuse).

\section{Convection to explain periodicities}

The disappearance of the periodicity of Betelgeuse after the dimming is consistent with the literature. However, our model only involves surface convection. This means that to explain the periodicity of Betelgeuse, we do not need the presence of pulsation of the atmosphere. Hence, it appears that surface convection is enough to explain the different periods of Betelgeuse. Concerning the stokes profile, we recover the LSP from stochastic effects. The linear polarization of Betelgeuse is also attributed to convective cells. Thus, our images of Betelgeuse and linear polarization can recover the periodicity of Betelgeuse by only involving surface dynamics which are linked to convection. To support the argument that the variability of Betelgeuse is due to the convection, we can also look at the timescale of convective structures. The surface of RSGs are made of giant convective cells, that live from months to years. Once again, the periodicity of $400$ days, or even the LSP of 2000 days often found in RSGs, are probably linked to the timescale where those super granules evolve. The LS of stokes Q shows a strong signal around 500 days around $\sim 50$ km/s. This signal can be explained by convective plumes that are beyond the limb. Those event happen occasionally in Betelgeuse, but might explain a periodicity. If a plume rises beyond the limb of Betelgeuse, more light is coming toward us, leading to an increase in luminosity. After a few month, the plume falls back and the luminosity of Betelgeuse decreases. Another explanation might be due to choc waves.

\begin{figure}
    \centering
    \includegraphics[width=0.4\textwidth]{LS Intensity.png}
    \caption{stokes I}
    \label{LS Intensity}
\end{figure}
\begin{figure}
    \centering
    \includegraphics[width=0.4\textwidth]{LS stokes Q.png}
    \caption{stokes Q}
    \label{LS Q}
\end{figure}

\begin{figure}
    \centering
    \includegraphics[width=0.4\textwidth]{LS stokes U.png}
    \caption{stokes U}
    \label{LS U}
\end{figure}

\begin{figure}
    \centering
    \includegraphics[width=0.4\textwidth]{LS photocentre Narval and Neo Narval.png}
    \caption{narval + neonarval}
    \label{LS narval + neonarval}
\end{figure}

\begin{figure}
    \centering
    \includegraphics[width=0.4\textwidth]{LS photocentre Narval.png}
    \caption{narval}
    \label{LS narval}
\end{figure}

\end{document}

### 2023-12-21 main.tex
Add files via upload

%                                                                 aa.dem
% AA vers. 9.1, LaTeX class for Astronomy & Astrophysics
% demonstration file
%                                                       (c) EDP Sciences
%-----------------------------------------------------------------------
%
%\documentclass[referee]{aa} % for a referee version
%\documentclass[onecolumn]{aa} % for a paper on 1 column  
%\documentclass[longauth]{aa} % for the long lists of affiliations 
%\documentclass[letter]{aa} % for the letters 
%\documentclass[bibyear]{aa} % if the references are not structured 
%                              according to the author-year natbib style

%
\documentclass{aa}

%
\usepackage{graphicx}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage{txfonts}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\usepackage[options]{hyperref}
% To add links in your PDF file, use the package "hyperref"
% with options according to your LaTeX or PDFLaTeX drivers.
%
\begin{document}

   \title{The periodicity of Betelgeuse explained by surface convection}

    \author{{ Q.~Pilate}\inst{1},{ A.~L{\'o}pez Ariste}\inst{1},{ A.~Lavail}\inst{1},{ Ph. Mathias}\inst{1} }

   \institute{IRAP, Universit\'e de Toulouse, CNRS, CNES, UPS.  14, Av. E. Belin. 31400 Toulouse, France 
             }

   \date{Received ...; accepted ...}

% \abstract{}{}{}{}{} 
% 5 {} token are mandatory

  \abstract

   \keywords{
               }

   \maketitle
%
%-------------------------------------------------------------------

\section{Introduction}
Betelgeuse is a red super giant star which has been observed by many astronomers for centuries. This star is known as a semi variable star. In the literature, the variability of the light curve of Betelgeuse is mainly explained by pulsation of the atmosphere. The most common periods found in Betelgeuse are the long secondary period (LSP) which is around $2000$ days and there are few known overtones such as $400$ days or $200$ days. Since the great dimming between the end of 2019 and beginning of 2020, those periods have changed. In this article, we attempt to recover the different periods of Betelgeuse using surface convection. The brightness of Betelgeuse is linked to the giant convective cells present at the surface. Hence, the light curve will be affected by the size and the number of convective cells at the surface and also by the timescale on those convective cells live. To explain to variability of Betelgeuse, we used the polarimetric data from the telescope Bernard Lyot, which observed Betelgeuse over a decade almost weekly. The linear polarization in Betelgeuse is associated to brightness inomogeneities at the surface of the star, due to huge convective cells at the surface. From the linear polarization signal, we were able to produce images of Betelgeuse. From the variability of the linear polarization and the intensity signal, we attempted to recover the variability of Betelgeuse. We also tried to recover the variability using the images produced by polarimetry. The photo-center displacement is affected by the size, number and timescale of convective cells. Using our images, we followed the displacement of the photo-center of Betelgeuse over 10 years to recover the variability of Betelgeuse.

\section{Method}

Trying to find periods in an astrophysical context can be difficult due to the unevenly spaced observed data. To overcome this issue, we used the Lomb-Scargle (LS) periodogram. For a given set of data, it fits a periodic signal using least squares method, for each frequency between the first and the last observation. The better the periodic signal fits the data, the higher the power attributed to the frequency. We used the LS periodogram in the linear polarization ($\sqrt{Q^2+U^2}$) and intensity signal first.

\begin{figure}[!h]
    \centering
    \includegraphics[width=0.5\textwidth]{LS Intensity.png}
    \caption{Lomb-Scargle periodogram of intensity over 10 years. Red regions represent high LS power (i.e most plausible frequencies) whereas blue regions stands for low LS power. }
    \label{LS Intensity}
\end{figure}

Figure \ref{LS Intensity} shows the LS periodogram of intensity for each frequency between -25 km/s and 70 km/s. The white dashed lines reprent the three main sequences found in the literature i.e the LSP and its overtones at 400 days and 250 days. From the intensity signal, we recover the LSP around 2000 days. For lower frequencies (i.e periods higher than 2000 days), the signal is attributed to cross correlation. It appears that there is a small signal around $0.003 d^{-1}$ ($\sim 333$ days) but it is hard to affirm that it corresponds a net signal or noise. \\

Figure \ref{LS Pl} shows the LS periodogram of linear polarization. We see that we find a strong power around a period of 400 days (second dashed line). In the linear polarization, the LSP does not appear in the LS periodogram. Hence, from the intensity and the linear polarization signal, we recover two of the three main periods found in Betelgeuse.Interestingly, the period around 250 days is present in the LS periodogram of stokes Q (see figure \ref{LS Q}) but not in stokes U. The nature of the differents periodicity is discussed in the next section, especially the one found in the linear polarization.

\begin{figure}[!h]
    \centering
    \includegraphics[width=0.5\textwidth]{LS Pl.png}
    \caption{Lomb-Scargle periodogram of linear polarization. The colour code is the same as figure \ref{LS Intensity}. }
    \label{LS Pl}
\end{figure}

\section{Convection to explain periodicities}

In the previous section, we have seen that we recover the different period of Betelgeuse from the intensity and linear polarization signal. While the nature of the LSP is still in debate, the nature of the lower periods is attributed to pulsation of the atmosphere of Betelgeuse. However, the mechanism behind the pulsation of Betelgeuse is less clear. In this section, we attempt to find an explanation to the 400 days periodicity of Betelgeuse. As mentioned before, the LS periodogram of linear polarization shows a periodicity around 400 days. Furthermore, the region where the LS power is high is around a wavelength of 40km/s. This wavelength correspond to the red wing of Betelgeuse. The polarization signal is supposed to stand between -20 and 40km/s. Signal above 40km/s is associated to plasma rising behind the limb of Betelgeuse. Hence, the strong power associated at this wavelength might be due to those convective plumes that are rising above the limb from time to time. From this scenario, we can explained the 400 days period in the light curve as follow: from time to time, Betelgeuse is ejecting plasma in the interstellar medium. Sometime, the plasma is ejected behind Betelgeuse, but rises so high that we can see the rising plume from Earth. This plume increases the number of photons that come to the Earth, leading to an increase of the brightness. After a few weeks or months, the plumes fall back to Betelgeuse, decreasing the brightness of Betelgeuse. This would explain why we recover the 400 days periodicity from linear polarization.

\begin{figure}[!h]
    \centering
    \includegraphics[width=0.5\textwidth]{LS stokes Q.png}
    \caption{stokes Q}
    \label{LS Q}
\end{figure}

\end{document}

### 2024-01-23 main.tex
Working on the introduction
Nouvelle version de l'introduction

@@
 
-   \title{The periodicity of Betelgeuse explained by surface convection}
+   \title{The variability of Betelgeuse explained by surface convection}
 
@@
 \section{Introduction}
-Betelgeuse is a red super giant star which has been observed by many astronomers for centuries.
-This star is known as a semi variable star.
-In the literature, the variability of the light curve of Betelgeuse is mainly explained by pulsation of the atmosphere.
-The most common periods found in Betelgeuse are the long secondary period (LSP) which is around $2000$ days and there are few known overtones such as $400$ days or $200$ days.
-Since the great dimming between the end of 2019 and beginning of 2020, those periods have changed.
-In this article, we attempt to recover the different periods of Betelgeuse using surface convection.
-The brightness of Betelgeuse is linked to the giant convective cells present at the surface.
-Hence, the light curve will be affected by the size and the number of convective cells at the surface and also by the timescale on those convective cells live.
-To explain to variability of Betelgeuse, we used the polarimetric data from the telescope Bernard Lyot, which observed Betelgeuse over a decade almost weekly.
-The linear polarization in Betelgeuse is associated to brightness inomogeneities at the surface of the star, due to huge convective cells at the surface.
-From the linear polarization signal, we were able to produce images of Betelgeuse.
-From the variability of the linear polarization and the intensity signal, we attempted to recover the variability of Betelgeuse.
-We also tried to recover the variability using the images produced by polarimetry.
-The photo-center displacement is affected by the size, number and timescale of convective cells.
-Using our images, we followed the displacement of the photo-center of Betelgeuse over 10 years to recover the variability of Betelgeuse.
+
+Betelgeuse is a prototypical red supergiant (RSG) known to be a semi-regular variable. Several periods can be found in the 
+literature usually clustered around the so-called long secondary period (LSP) of 2000 days and other more or less well defined 
+periods around 400 and 200 days. Such periods are often linked to radial pulsation modes,   and have been used to model 
+the caracteristics of the resonant cavity, hence the size of the star and eventually its evolutionary stage \cite{XXX}.
+
+At the end of 2019, Betelgeuse suffered a sudden and dramatic loss of brightness \cite{}, particularly in the blue colors. 
+\cite{} has demonstrated that such event was due to the emission by the star of a cloud of dust near to the line of sight of 
+Earth. In this event it was clear that a change in brightness was not due to any kind of pulsation but to a change on the 
+brightness distribution over the stellar disk due to the presence of large quantities of dust. While this may be seen as 
+a singular event, it brings up the question whether all changes in brightness cannot be due to similar changes in the brightness 
+distribution over the disk.
+
+Betelgeuse is known to present large convective cells with life times in the order of one to two years and measurable changes 
+in the span of one week. Bright convection cells near disk center will increase the integrated brightness of the star when 
+compared with other situations where such cells are found near the edges. So even without calling for the formation of 
+dust or other changes in opacity, one may argue that the simple convective patterns of the star may create a variability. Such 
+variability could be expected to be random, but will present quasi-periodicities related to the typical time scales of the 
+convective patterns.
+
+Since the end of the Great Dimming event, Betelgeuse has continue its random variation of brightness. But the classical periods 
+of variation cited above are not visible any more. This may be due to simply the short span of time, that does not afford for 
+the peaks to appear prominently in the variability spectra. But it has been sufficient to bring up that other explanation of the variability
+in terms  of convective patterns. In order to address this question we seek for the typical periods of 1000, 400 and 200 days in  
+observational proxies related to the convective activite but unrelated to pulsations nor directly linked to brightness?
+
+Linear polarization in the atomic lines of the spectra of Betelgeuse, discovered by \cite{}, has been interpreted as the joint action of 
+two mechanisms. First the depolarization of the continuum by atoms which absorb linearly polarized light and re-emit unpolarized light, the 
+continuum photons being polarized by Rayleigh scattering. This depolarization produces signals with an azimuth symmetry over the disk that 
+would  cancel out the net linear polarization on a homogeneous disk. Thus it must be combined with an inhomogenous disk to produce the net linear polarization 
+signal observed. This interpretation suggested the possibility of mapping those brightness inhomogeneities. This has been done by \cite{} 
+even producing 3-dimensional images of the atmosphere of Betelgeuse \cite{} when taking advantage of the different height of formation of 
+different lines in the spectrum of Betelgeuse. The images produced compare very well with cotemporaneous images made with interferometric 
+techniques and show clear convective patterns. Linear polarization in the atomic lines is therefore a proxy of convection, unrelated to radial 
+pulsations but linked to the brightness inhomogeneities due to the convective patterns in Betelgeuse. If one can find the aforementioned periods 
+in linear polarization, we must conclude that these periods are related to the convective activity which originates those linear polarization 
+and unrelated to any radial pulsation phenomena.
+
+Betelgeuse is a red super giant star which has been observed by many astronomers for centuries. This star is known as a 
+semi variable star. In the literature, the variability of the light curve of Betelgeuse is mainly explained by pulsation of 
+the atmosphere. The most common periods found in Betelgeuse are the long secondary period (LSP) which is around $2000$ days 
+and there are few known overtones such as $400$ days or $200$ days. Since the great dimming between the end of 2019 and 
+beginning of 2020, those periods have changed. In this article, we attempt to recover the different periods of Betelgeuse 
+using surface convection. The brightness of Betelgeuse is linked to the giant convective cells present at the surface. Hence, 
+the light curve will be affected by the size and the number of convective cells at the surface and also by the timescale on 
+those convective cells live. To explain to variability of Betelgeuse, we used the polarimetric data from the telescope Bernard 
+Lyot, which observed Betelgeuse over a decade almost weekly. The linear polarization in Betelgeuse is associated to brightness 
+inomogeneities at the surface of the star, due to huge convective cells at the surface. From the linear polarization signal, 
+we were able to produce images of Betelgeuse. From the variability of the linear polarization and the intensity signal, we 
+attempted to recover the variability of Betelgeuse. We also tried to recover the variability using the images produced by 
+polarimetry. The photo-center displacement is affected by the size, number and timescale of convective cells. Using our 
+images, we followed the displacement of the photo-center of Betelgeuse over 10 years to recover the variability of 
+Betelgeuse.
 

### 2024-01-29 main.tex
modified:   main.tex
v2

@@
 
-\section{Method}
+\section{Method: Searching for periodicities in the polarization spectra of Betelgeuse}
 
-Trying to find periods in an astrophysical context can be difficult due to the unevenly spaced observed data.
-To overcome this issue, we used the Lomb-Scargle (LS) periodogram.
-For a given set of data, it fits a periodic signal using least squares method, for each frequency between the first and the last observation.
-The better the periodic signal fits the data, the higher the power attributed to the frequency.
+Betelgeuse has been observed for the last 13 years with Narval and Neo-Narval at the Telescope Bernard Lyot. These instruments measure the polarization 
+over the visible spectra of Betelgeuese (390-1000 nm) with high spectral resolution (R=65000) and high polarimetric sensitivity. Despite 
+this sensitivity, the signal-to-noise ratios are not sufficient to measure the weak polarization signals in the individual atomic lines of the 
+spectrum. These amplitudes are today known to be of the order of $10^{-4}$ times the continuum intensity, and only excepcionnaly do they 
+reach amplitudes of $10^{-3}$ the continuum. In these exceptional observations, enough photons can be accumulated per spectral bin to 
+see the linear polarization signal above noise \cite{Auriere}. Most commonly, the amplitudes of linear polarization are below noise levels. In those 
+occassions we have to add up the signals of thousands of lines to reduce the noise and increase the signal-to-noise ratios. This line addition 
+is done through a technique called Least-Squares Deconvolution (LSD) \cite{Donati} and assumes that the linear polarization signal is similar 
+in all spectral lines up to scale factors in the sampling and amplitude. This technique has been successfully used in the past to measure 
+magnetic field distributions over stellar surfaces and it is now used to produce images of the brightness distributions in the photosphere 
+of Betelgeuse and other red superg
+
+Trying to find periods in an astrophysical context can be difficult due to the unevenly spaced observed data. To overcome this issue, 
+we used the Lomb-Scargle (LS) periodogram. For a given set of data, it fits a periodic signal using least squares method, for each frequency 
+between the first and the last observation. The better the periodic signal fits the data, the higher the power attributed to the frequency. 
 We used the LS periodogram in the linear polarization ($\sqrt{Q^2+U^2}$) and intensity signal first.

### 2024-01-29 main.tex
modified:   main.tex

@@
 
+\subsection{Lomb-Scargle periodogram}
+
 Trying to find periods in an astrophysical context can be difficult due to the unevenly spaced observed data. To overcome this issue, 
-we used the Lomb-Scargle (LS) periodogram. For a given set of data, it fits a periodic signal using least squares method, for each frequency 
-between the first and the last observation. The better the periodic signal fits the data, the higher the power attributed to the frequency. 
-We used the LS periodogram in the linear polarization ($\sqrt{Q^2+U^2}$) and intensity signal first.
+we used the Lomb-Scargle (LS) periodogram \citep{Lomb1976,Scargle1982}. For a given set of data, it fits a sinusoid function using least-squares at each 
+frequency between the first and the last observation. The better the fit, the higher the power attributed to the given frequency. Since Betelgeuse has 
+been observed by the Télescope Bennard Lyot at pic du midi since 2013, we have ten years of data. Betelgeuse was observed every two or three weeks by 
+the spectropolarimeters Narval and Neo-Narval, except during summer. From all these data, we can compute least-squared 
+deconvolution (LSD) \citep{Donati1997} profile of Stokes I, U and Q. The linear polarization present in Betelgeuse is attributed to brightness 
+inhomogeneities, which are due to the convective cells present at the surface \citep{LopezAriste2018}. To recover variability of Betelgeuse, 
+we computed the LS periodogram at each wavelength of the LSD profile, between -25 km/s and 80 km/s. We choose those velocity to span the entire 
+signal present in the LSD profile. In the work of \cite{LopezAriste2018}, the most blueshifted signal was present at -20 km/s, 
+corresponding to the maximal velocity of the rising plasma. The most redshifted signal was $\sim$ 40 km/s. This velocity corresponds 
+to the velocity of the star in our reference frame \citep{LopezAriste2018}.
 
-\begin{figure}[!h]
+\subsection{Before and after the dimming}
+
+At the beginning of 2020, Betelgeuse reached a historical minimum in its luminosity, called the great dimming \citep{Guinan2020}. 
+From interferometric data, we know that this event was caused by a cloud of gas in front of the line of sight \citep{Montargès2021}. 
+Interferometric images show a huge drop in luminosity in the southern hemisphere of Betelgeuse, leading to this dimming. Interestingly, 
+it has been shown by \cite{jadlovsky2022} that the periodicity of Betelgeuse has changed before and after the dimming. Using the light curve from AAVSO, 
+the authors showed that before the dimming, the dominant period of Betelgeuse was around 400 days. However, after the dimming, this period has changed and 
+is now shorter, around 200 days. It means that, after the dimming, there was a change in the behaviour of the atmosphere of Betelgeuse: the period is now
+shorter than before the dimming. Our analysis is based on this change of periodicity. By computing the Lomb-Scargle periodogram for each frequency 
+in the LSD profile, we expect to find periodicities around the LSP, but also on the 400 days and 200 days period. However, we expect a change
+in the periodicity after the dimming. This would lead to a decrease in the Lomb-Scargle power. 
+Figure \ref{LS intensity},\ref{LS Q},\ref{LS U} and \ref{LS linear polarization} show the Lomb-Scargle periodogram of the LSD profile of respectively 
+the intensity, stokes Q, stokes U and the linear polarization. The abscissa is the frequency, while the ordinate is the heliocentric radial velocity. 
+For each figure, the upper panel is the LS periodogram with data taken before the great dimming of Betelgeuse, that is from 2013 to 2019. 
+The lower panel is the LS periodogram for the data before and after the dimming: from 2013 to 2023. The white dashed lines represent the different 
+periods found in the literature: the LSP at 2000 days, the fondamental pressure mode at 400 days and its first overtone at 200 days. The Lomb-Scargle 
+power is normalised. The colorbar indicates the quality of the fit: red areas correspond to good fits from the LS periodogram, those areas are good 
+candidates for periodicities.
+
+\begin{figure*}[!h]
     \centering
-    \includegraphics[width=0.5\textwidth]{LS Intensity.png}
-    \caption{Lomb-Scargle periodogram of intensity over 10 years. Red regions represent high LS power (i.e most plausible frequencies) whereas blue regions stands for low LS power. }
-    \label{LS Intensity}
-\end{figure}
+    \includegraphics[width=\textwidth]{Lomb-Scargle Intensity.png}
+    \caption{Lomb-Scargle periodogram of the LSD profile of the intensity.
+The upper pannel is the LS periodogram for data before the great dimming.
+The lower panel is periodogram after the dimming.
+The ordinate refers to the velocity in the heliocentric radial velocity in km/s.
+The abscissa is the period in $\mathrm{days^{-1}}$.
+The three white dashed lines represent respectively (from left to right) the LSP at 2000 days, the 400 days period and its first overtone at 200 days.
+The Lomb-Scargle power refers to the quality of period found.
+}
+    \label{LS intensity}
+\end{figure*}
 
-Figure \ref{LS Intensity} shows the LS periodogram of intensity for each frequency between -25 km/s and 70 km/s.
-The white dashed lines reprent the three main sequences found in the literature i.e the LSP and its overtones at 400 days and 250 days.
-From the intensity signal, we recover the LSP around 2000 days.
-For lower frequencies (i.e periods higher than 2000 days), the signal is attributed to cross correlation.
-It appears that there is a small signal around $0.003 d^{-1}$ ($\sim 333$ days) but it is hard to affirm that it corresponds a net signal or noise.
-\\
+\begin{figure*}[!h]
+    \centering
+    \includegraphics[width=\textwidth]{Lomb-Scargle Stokes Q.png}
+    \caption{Same as figure \ref{LS intensity} but for stokes Q. }
+    \label{LS Q}
+\end{figure*}
 
-Figure \ref{LS Pl} shows the LS periodogram of linear polarization.
-We see that we find a strong power around a period of 400 days (second dashed line).
-In the linear polarization, the LSP does not appear in the LS periodogram.
-Hence, from the intensity and the linear polarization signal, we recover two of the three main periods found in Betelgeuse.Interestingly, the period around 250 days is present in the LS periodogram of stokes Q (see figure \ref{LS Q}) but not in stokes U.
-The nature of the differents periodicity is discussed in the next section, especially the one found in the linear polarization.
+\begin{figure*}[!h]
+    \centering
+    \includegraphics[width=\textwidth]{Lomb-Scargle Stokes U.png}
+    \caption{Same as figure \ref{LS intensity}, but for stokes U.}
+    \label{LS U}
+\end{figure*}
 
-\begin{figure}[!h]
+\begin{figure*}[!h]
     \centering
-    \includegraphics[width=0.5\textwidth]{LS Pl.png}
-    \caption{Lomb-Scargle periodogram of linear polarization. The colour code is the same as figure \ref{LS Intensity}. }
-    \label{LS Pl}
-\end{figure}
+    \includegraphics[width=\textwidth]{Lomb-Scargle linear polarization.png}
+    \caption{Same as figure \ref{LS intensity}, but for the linear polarization: $\sqrt{\mathrm{Q^2}+\mathrm{U^2}}$.}
+    \label{LS linear polarization}
+\end{figure*}
+
+\section{Recovering periodicity of Betelgeuse in the LSD profile}
+
+\subsection{Intensity}
+Figure \ref{LS intensity} shows the LS periodogram of the LSD of the intensity profile. Interestingly, we recover the LSP at 2000 days, 
+located at a HRV of $\sim$ 20 km/s. This periodicity seems to change after the dimming, where it is located at a HRV between 0 and 10 km/s. 
+There is also a weaker signal at the fundamental mode, around 400 days. This signal completely disappears after the dimming. It is hard to attribute 
+this LS power to a periodicity of Betelgeuse, however it might be a hint that this periodicity has changed after the dimming. Furthermore, our results 
+in the intensity profile are consistent with the one from \cite{Mathias2018} who used 2D fourier analysis and found that the dominant period was the LSP. 
+We can also see that there is a less stronger signal at 40 km/s in the LSP. We will discuss about it in the next section, but the fact that there is a weak 
+signal at 40 km/s, which is the velocity of the star in the heliocentric reference frame, could be a clue to understand the periodicity of Betelgeuse. 
+The intensity is sensitive to the whole star, hence the periodogram of the intensity is sensitive to global variation of the star. It is sensitive to 
+the timescale on which giant convective cells evolve. It has been observed that some convective cells can live for more than two years. Even though the
+LSP is around five years, we can attribute the LSP to the number of convective cells during a given time.
+
+\subsection{Stokes Q and U}
+
+Figure \ref{LS Q} and \ref{LS U} show the LS periodogram of both stokes Q and U. It is clear that the LS periodogram is completely different 
+from the intensity one. We can see a strong signal in fig. \ref{LS Q} around 200 days. Once again, this signal is much more weak after the dimming.
+Concerning stokes U in fig. \ref{LS U}, it appears that we only recover the LSP, which once again, is weaker after the great dimming. 
+Those periodograms rise a lot of questions: why is the 200 days period only present in stokes Q? Why not in stokes U or intensity? 
+We will discuss about those questions in the next question. However, it also appears that stokes Q is not sensitive to the LSP, or to be more precise, 
+it is much more sensitive to the 200 days period. From this, we can conclude that stokes Q evolves on a time scale much more closer to the first overtone 
+mode than the LSP or even the fundamental mode. Because stokes Q is sensitive only to shorter periods, the variation of stokes Q come from another mechanism 
+than the one from intensity. Because the linear polarization in Betelgeuse is due to convective cells, we attribute the signal present in the periodogram 
+as the lifetime of convective cells. This is what we observe in the LSD profile of Betelgeuse: from a week to another it does not change that much. 
+However, after a few month, the LSD profile of Betelgeuse has completely changed, due to the evolution of convective cells at the surface. 
+This would explain the strong signal present at 200 days in stokes Q.
+
+\subsection{The linear polarization}
+
+Figure \ref{LS linear polarization} shows the LS periodogram of the linear polarization of Betelgeuse: $\sqrt{\mathrm{Q^2+U^2}}$. 
+Interestingly, the LS periodogram shows a strong power at 400 days and 2000 days. This signal is much more weaker after the dimming.
+While the LSP is already present in the LS of intensity, here we recover the 400 days periodicity, often invoke in the literature. 
+Nonetheless, the period is present at a HRV of around 40 km/s. We recall that this velocity is the velocity of the star \cite{LopezAriste2018}. 
+Every signal above the velocity of the star is either due to the plasma sinking in Betelgeuse, or plumes of plasma rising above the limb of the star, 
+as seen in other red supergiant like $\mu$Cep \citep{LopezAriste2023}. Up to this point, we want to pinpoint the two elements that are a clue 
+to understand the variability of Betelgeuse: the 400 days period is present in the linear polarization signal and also at a HRV above 40 km/s.
+In any case, from the linear polarization and the intensity profile, we recover the different periods of Betelgeuse.
 
@@
 In the previous section, we have seen that we recover the different period of Betelgeuse from the intensity and linear polarization signal. 
-While the nature of the LSP is still in debate, the nature of the lower periods is attributed to pulsation of the atmosphere of Betelgeuse.
-However, the mechanism behind the pulsation of Betelgeuse is less clear.
-In this section, we attempt to find an explanation to the 400 days periodicity of Betelgeuse.
-As mentioned before, the LS periodogram of linear polarization shows a periodicity around 400 days.
-Furthermore, the region where the LS power is high is around a wavelength of 40km/s.
-This wavelength correspond to the red wing of Betelgeuse.
-The polarization signal is supposed to stand between -20 and 40km/s.
-Signal above 40km/s is associated to plasma rising behind the limb of Betelgeuse.
-Hence, the strong power associated at this wavelength might be due to those convective plumes that are rising above the limb from time to time.
-From this scenario, we can explained the 400 days period in the light curve as follow: from time to time, Betelgeuse is ejecting plasma in the interstellar medium.
-Sometime, the plasma is ejected behind Betelgeuse, but rises so high that we can see the rising plume from Earth.
-This plume increases the number of photons that come to the Earth, leading to an increase of the brightness.
-After a few weeks or months, the plumes fall back to Betelgeuse, decreasing the brightness of Betelgeuse.
-This would explain why we recover the 400 days periodicity from linear polarization.
-
-\begin{figure}[!h]
-    \centering
-    \includegraphics[width=0.5\textwidth]{LS stokes Q.png}
-    \caption{stokes Q}
-    \label{LS Q}
-\end{figure}
+While the nature of the LSP is still in debate, the nature of the lower periods is attributed to pulsation of the atmosphere of Betelgeuse 
+(e.g \cite{Kiss2006}). The mechanism behind the pulsation of Betelgeuse is attributed to the $\kappa$-mechanism. In this section, we attempt to 
+find an explanation to the 400 days periodicity of Betelgeuse. As mentioned before, the LS periodogram of linear polarization shows a periodicity 
+around 400 days. Furthermore, the region where the LS power is high around 40km/s. This wavelength correspond to the red wing of Betelgeuse.
+The polarization signal is supposed to stand between -20 and 40km/s. Signal above 40km/s is associated to plasma rising behind the limb of Betelgeuse. 
+This mechanism has been observed in the RSG $\mu$Cep and was introduced by \cite{LopezAriste2023} to explain the excess of linear polarization beyond 
+the velocity of the star. In $\mu$Cep, the linear polarization signal is very strong and often present, whereas in Betelgeuse, the signal above the 
+velocity of the star is less strong. However, because it is present from time to time, the LS periodogram finds this signal above the limb and finds a
+periodicity at this wavelength of 40 km/s. Hence, the strong power associated to this wavelength might be due to those convective plumes that are 
+rising above the limb from time to time. From this scenario, we can explained the 400 days period in the light curve as follow: from time to time,
+Betelgeuse is ejecting plasma in the interstellar medium due to winds. The plasma can be ejected from the hidden face of Betelgeuse. 
+It rises so high that we can see the rising plume from the Earth, because the plume rises above the limb of the star. Hence, from Earth, 
+the signal associated to this plume appears redshifted compared to the velocity of the star. This plume increases the number of photons that we receive,
+leading to an increase of the brightness. After a few weeks or months, the plume falls back to Betelgeuse, desapearing from our point of view, 
+leading to a decrease of the brightness of Betelgeuse. This is exactly what we see is the LSD profile of the linear polarization of Betelgeuse: 
+from time to time, there is a net signal above 40km/s, which is present for a few weeks, and then disappears. Since this signal is not always present, 
+it is found by the LS periodogram as a periodic event. However, it is not periodic at all: because the plumes rise at random moments, sometimes they 
+appear every years, sometime more often. This would explain why we only see it before the dimming, after the dimming, this kind of event never happened 
+for the moment. Thus, the LS periodogram, at that precise frequency, finds that there is a periodic signal before the dimming, but not after, leading to
+a lower LS power at this frequency. Concerning the 200 days period, it is caused by the timescale on which the stokes parameters evolve. For example, the 
+LSD profile of stokes Q or U will slightly change if we observe Betelgeuse every two weeks. However, on a longer timescale, for example a few month, 
+the LSD profile of stokes Q and U will change a lot. They are directly linked to the dynamics of the surface, hence,  their variation scale the lifetime 
+of convective cells. Some convective cells can live up to a few month, leading to this peak in the LS periodogram of stokes Q. As showed by
+\cite{LopezAriste2018}, some convective cells can live up to years, explaining also the LSP. One can notice that the peak is not present in stokes U. 
+However, if we try to explain the periodicity of Betelgeuse using pulsations, we would see this signal in either stokes U and the linear polarization. 
+To summarise what we have done so far: we were able to explain the variability of Betelgeuse by using surface convection only. The convective cells play 
+the most important role in the variability of Betelgeuse. Our work does not need pulsation to explain its variability. This does not mean that there are 
+no pulsation in Betelgeuse. What we claim is that the most important factor in the variability of Betelgeuse is the timescale on which the convective 
+cells evolve. Pulsation could exist on Betelgeuse, but probably on shorter timescale than the one observed.
 

### 2024-01-30 main.tex
Bib added

@@
 At the end of 2019, Betelgeuse suffered a sudden and dramatic loss of brightness \cite{}, particularly in the blue colors. 
-\cite{} has demonstrated that such event was due to the emission by the star of a cloud of dust near to the line of sight of 
+\cite{Miguel} has demonstrated that such event was due to the emission by the star of a cloud of dust near to the line of sight of 
 Earth. In this event it was clear that a change in brightness was not due to any kind of pulsation but to a change on the 
@@
 
-Linear polarization in the atomic lines of the spectra of Betelgeuse, discovered by \cite{}, has been interpreted as the joint action of 
+Linear polarization in the atomic lines of the spectra of Betelgeuse, discovered by \cite{auriere_discovery_2016}, has been interpreted as the joint action of 
 two mechanisms. First the depolarization of the continuum by atoms which absorb linearly polarized light and re-emit unpolarized light, the 
@@
 would  cancel out the net linear polarization on a homogeneous disk. Thus it must be combined with an inhomogenous disk to produce the net linear polarization 
-signal observed. This interpretation suggested the possibility of mapping those brightness inhomogeneities. This has been done by \cite{} 
-even producing 3-dimensional images of the atmosphere of Betelgeuse \cite{} when taking advantage of the different height of formation of 
+signal observed. This interpretation suggested the possibility of mapping those brightness inhomogeneities. This has been done by \cite{lopez_ariste_convective_2018} 
+even producing 3-dimensional images of the atmosphere of Betelgeuse \citep{lopez_ariste_three-dimensional_2022} when taking advantage of the different height of formation of 
 different lines in the spectrum of Betelgeuse. The images produced compare very well with cotemporaneous images made with interferometric 
@@
 reach amplitudes of $10^{-3}$ the continuum. In these exceptional observations, enough photons can be accumulated per spectral bin to 
-see the linear polarization signal above noise \cite{Auriere}. Most commonly, the amplitudes of linear polarization are below noise levels. In those 
+see the linear polarization signal above noise \citep{auriere_discovery_2016}. Most commonly, the amplitudes of linear polarization are below noise levels. In those 
 occassions we have to add up the signals of thousands of lines to reduce the noise and increase the signal-to-noise ratios. This line addition 
-is done through a technique called Least-Squares Deconvolution (LSD) \cite{Donati} and assumes that the linear polarization signal is similar 
+is done through a technique called Least-Squares Deconvolution (LSD) \citep{donati_spectropolarimetric_1997} and assumes that the 
+linear polarization signal is similar 
 in all spectral lines up to scale factors in the sampling and amplitude. This technique has been successfully used in the past to measure 
 magnetic field distributions over stellar surfaces and it is now used to produce images of the brightness distributions in the photosphere 
-of Betelgeuse and other red superg
+of Betelgeuse and other red supergiants as cited in the Introduction.
 
@@
 
+\begin{acknowledgements}
+    This work was supported by the "Programme National de Physique Stellaire" (PNPS) of CNRS/INSU co-funded by CEA and CNES.
+    We acknowledge support from the French National Research Agency (ANR)
+    funded project PEPPER (ANR-20-CE31-0002)
+    \end{acknowledgements}
+
+    \bibliographystyle{aa}
+    %\bibliographystyle{aa}
+
+    \bibliography{art75}
 \end{document}

### 2024-01-30 main.tex
Corrected bib
sect 2

@@
 of Betelgeuse and other red supergiants as cited in the Introduction.
+
+In this work we shall look into the signals produced through these line-addition techniques, and which produce a pseudo-spectral line in intensity 
+and linear polarization which does not belong to any atomic species in particular but to an average of all of them present and emitting in the 
+photosphere of Betelgeuse. These signals carry the coherent signals present in the photospheric lines, but the particularities of this or 
+that spectral line are erased. All these data has been presented before by \cite{auriere_discovery_2016}, \cite{mathias_evolution_2018} and 
+\cite{lopez_ariste_three-dimensional_2022}, which also give details on observing times and conditions as well as more detailed description 
+on the data reduction \footnote[1]{Beyond a 2-year proprietary embargo, and up to technical issues, all these data is available through 
+PolarBase (http://polarbase.irap.omp.eu/).}
 

### 2024-02-06 main.tex
modified:   art75.bib 	modified:   main.tex

@@
 
-Betelgeuse is a red super giant star which has been observed by many astronomers for centuries. This star is known as a 
-semi variable star. In the literature, the variability of the light curve of Betelgeuse is mainly explained by pulsation of 
-the atmosphere. The most common periods found in Betelgeuse are the long secondary period (LSP) which is around $2000$ days 
-and there are few known overtones such as $400$ days or $200$ days. Since the great dimming between the end of 2019 and 
-beginning of 2020, those periods have changed. In this article, we attempt to recover the different periods of Betelgeuse 
-using surface convection. The brightness of Betelgeuse is linked to the giant convective cells present at the surface. Hence, 
-the light curve will be affected by the size and the number of convective cells at the surface and also by the timescale on 
-those convective cells live. To explain to variability of Betelgeuse, we used the polarimetric data from the telescope Bernard 
-Lyot, which observed Betelgeuse over a decade almost weekly. The linear polarization in Betelgeuse is associated to brightness 
-inomogeneities at the surface of the star, due to huge convective cells at the surface. From the linear polarization signal, 
-we were able to produce images of Betelgeuse. From the variability of the linear polarization and the intensity signal, we 
-attempted to recover the variability of Betelgeuse. We also tried to recover the variability using the images produced by 
-polarimetry. The photo-center displacement is affected by the size, number and timescale of convective cells. Using our 
-images, we followed the displacement of the photo-center of Betelgeuse over 10 years to recover the variability of 
-Betelgeuse.
-
 \section{Method: Searching for periodicities in the polarization spectra of Betelgeuse}
@@
 Trying to find periods in an astrophysical context can be difficult due to the unevenly spaced observed data. To overcome this issue, 
-we used the Lomb-Scargle (LS) periodogram \citep{Lomb1976,Scargle1982}. For a given set of data, it fits a sinusoid function using least-squares at each 
+we used the Lomb-Scargle (LS) periodogram \citep{lomb_least-squares_1976,scargle_studies_1982}. For a given set of data, it fits a sinusoid function using least-squares at each 
 frequency between the first and the last observation. The better the fit, the higher the power attributed to the given frequency. Since Betelgeuse has 
 been observed by the Télescope Bennard Lyot at pic du midi since 2013, we have ten years of data. Betelgeuse was observed every two or three weeks by 
-the spectropolarimeters Narval and Neo-Narval, except during summer. From all these data, we can compute least-squared 
-deconvolution (LSD) \citep{Donati1997} profile of Stokes I, U and Q. The linear polarization present in Betelgeuse is attributed to brightness 
-inhomogeneities, which are due to the convective cells present at the surface \citep{LopezAriste2018}. To recover variability of Betelgeuse, 
+the spectropolarimeters Narval and Neo-Narval, except during summer.
+From all these data, we can compute the LSD profiles of Stokes I, U and Q.
+The linear polarization present in Betelgeuse is attributed to brightness
+inhomogeneities, which are due to the convective cells present at the surface \citep{lopez_ariste_convective_2018}. To recover variability of Betelgeuse, 
 we computed the LS periodogram at each wavelength of the LSD profile, between -25 km/s and 80 km/s. We choose those velocity to span the entire 
-signal present in the LSD profile. In the work of \cite{LopezAriste2018}, the most blueshifted signal was present at -20 km/s, 
+signal present in the LSD profile. In the work of \cite{lopez_ariste_convective_2018}, the most blueshifted signal was present at -20 km/s, 
 corresponding to the maximal velocity of the rising plasma. The most redshifted signal was $\sim$ 40 km/s. This velocity corresponds 
-to the velocity of the star in our reference frame \citep{LopezAriste2018}.
+to the velocity of the star in our reference frame.
 
@@
 
-At the beginning of 2020, Betelgeuse reached a historical minimum in its luminosity, called the great dimming \citep{Guinan2020}. 
-From interferometric data, we know that this event was caused by a cloud of gas in front of the line of sight \citep{Montargès2021}. 
+At the beginning of 2020, Betelgeuse reached a historical minimum in its luminosity, called the great dimming \citep{guinan_fall_2020}. 
+From interferometric data, we know that this event was caused by a cloud of gas in front of the line of sight \citep{montarges_dusty_2021}. 
 Interferometric images show a huge drop in luminosity in the southern hemisphere of Betelgeuse, leading to this dimming. Interestingly, 
-it has been shown by \cite{jadlovsky2022} that the periodicity of Betelgeuse has changed before and after the dimming. Using the light curve from AAVSO, 
+it has been shown by \cite{jadlovsky_analysis_2023} that the periodicity of Betelgeuse has changed before and after the dimming. Using the light curves from AAVSO, 
 the authors showed that before the dimming, the dominant period of Betelgeuse was around 400 days. However, after the dimming, this period has changed and 
@@
 periods found in the literature: the LSP at 2000 days, the fondamental pressure mode at 400 days and its first overtone at 200 days. The Lomb-Scargle 
-power is normalised. The colorbar indicates the quality of the fit: red areas correspond to good fits from the LS periodogram, those areas are good 
-candidates for periodicities.
+power is normalised. The colorbar indicates the quality of the fit: red areas correspond to good fits from the LS periodogram.
 
-\begin{figure*}[!h]
+\begin{figure}[!h]
     \centering
-    \includegraphics[width=\textwidth]{Lomb-Scargle Intensity.png}
-    \caption{Lomb-Scargle periodogram of the LSD profile of the intensity.
-The upper pannel is the LS periodogram for data before the great dimming.
-The lower panel is periodogram after the dimming.
+    \includegraphics[width=0.5\textwidth]{Lomb-Scargle Intensity.png}
+    \caption{Lomb-Scargle periodogram of the LSD profile of intensity.
+The upper pannel is the LS periodogram for data before the great dimming while the lower panel is the periodogram from the all dataset.
 The ordinate refers to the velocity in the heliocentric radial velocity in km/s.
@@
 The three white dashed lines represent respectively (from left to right) the LSP at 2000 days, the 400 days period and its first overtone at 200 days.
-The Lomb-Scargle power refers to the quality of period found.
+The Lomb-Scargle power refers to the quality of the fit from the LS periodogram.
 }
     \label{LS intensity}
-\end{figure*}
+\end{figure}
 
-\begin{figure*}[!h]
+\begin{figure}[!h]
     \centering
-    \includegraphics[width=\textwidth]{Lomb-Scargle Stokes Q.png}
+    \includegraphics[width=0.5\textwidth]{Lomb-Scargle Stokes Q.png}
     \caption{Same as figure \ref{LS intensity} but for stokes Q. }
     \label{LS Q}
-\end{figure*}
+\end{figure}
 
-\begin{figure*}[!h]
+\begin{figure}[!h]
     \centering
-    \includegraphics[width=\textwidth]{Lomb-Scargle Stokes U.png}
+    \includegraphics[width=0.5\textwidth]{Lomb-Scargle Stokes U.png}
     \caption{Same as figure \ref{LS intensity}, but for stokes U.}
     \label{LS U}
-\end{figure*}
+\end{figure}
 
-\begin{figure*}[!h]
+\begin{figure}[!h]
     \centering
-    \includegraphics[width=\textwidth]{Lomb-Scargle linear polarization.png}
+    \includegraphics[width=0.5\textwidth]{Lomb-Scargle linear polarization.png}
     \caption{Same as figure \ref{LS intensity}, but for the linear polarization: $\sqrt{\mathrm{Q^2}+\mathrm{U^2}}$.}
     \label{LS linear polarization}
-\end{figure*}
+\end{figure}
 
-\section{Recovering periodicity of Betelgeuse in the LSD profile}
+\section{Recovering the variability of Betelgeuse in the LSD profile}
 
@@
 this LS power to a periodicity of Betelgeuse, however it might be a hint that this periodicity has changed after the dimming. Furthermore, our results 
-in the intensity profile are consistent with the one from \cite{Mathias2018} who used 2D fourier analysis and found that the dominant period was the LSP. 
-We can also see that there is a less stronger signal at 40 km/s in the LSP. We will discuss about it in the next section, but the fact that there is a weak 
+in the intensity profile are consistent with the one from \cite{mathias_evolution_2018} who used 2D fourier analysis and found that the dominant period was the LSP. 
+We can also see that there is a less stronger signal at 40 km/s in the LSP. We will discuss about it in the next section, but this weak 
 signal at 40 km/s, which is the velocity of the star in the heliocentric reference frame, could be a clue to understand the periodicity of Betelgeuse.
-The intensity is sensitive to the whole star, hence the periodogram of the intensity is sensitive to global variation of the star. It is sensitive to 
-the timescale on which giant convective cells evolve. It has been observed that some convective cells can live for more than two years. Even though the
-LSP is around five years, we can attribute the LSP to the number of convective cells during a given time.
 
@@
 Figure \ref{LS Q} and \ref{LS U} show the LS periodogram of both stokes Q and U. It is clear that the LS periodogram is completely different 
-from the intensity one. We can see a strong signal in fig. \ref{LS Q} around 200 days. Once again, this signal is much more weak after the dimming.
+from the intensity one. We can see a strong signal in fig. \ref{LS Q} around 200 days. Once again, this signal is weaker after the dimming.
 Concerning stokes U in fig. \ref{LS U}, it appears that we only recover the LSP, which once again, is weaker after the great dimming. 
@@
 While the LSP is already present in the LS of intensity, here we recover the 400 days periodicity, often invoke in the literature. 
-Nonetheless, the period is present at a HRV of around 40 km/s. We recall that this velocity is the velocity of the star \cite{LopezAriste2018}. 
+Nonetheless, the period is present at a HRV of around 40 km/s. We recall that this velocity is the velocity of the star \cite{lopez_ariste_convective_2018}. 
 Every signal above the velocity of the star is either due to the plasma sinking in Betelgeuse, or plumes of plasma rising above the limb of the star, 
-as seen in other red supergiant like $\mu$Cep \citep{LopezAriste2023}. Up to this point, we want to pinpoint the two elements that are a clue 
+as seen in other red supergiant like $\mu$ Cep \citep{lopez_ariste_height_2023}. Up to this point, we want to pinpoint the two elements that are a clue 
 to understand the variability of Betelgeuse: the 400 days period is present in the linear polarization signal and also at a HRV above 40 km/s.
@@
 
+\subsection{The variability of Betelgeuse inferred from polarimetry imaging}
+
+Using linear polarization, \cite{lopez_ariste_convective_2018} has been able to reconstruct images of Betelgeuse that have been sucessfully compared to inteferometric images of \cite{montarges_close_2016}.
+The images are produced by finding the brightness distribution that better fits the observed linear polarization LSD profile using Marquardt-Levemberg minimisation \citep{lopez_ariste_convective_2018}.
+Because Betelgeuse is observed every two weeks by the TBL, we were able to follow its surface activity through polarimetry imaging.
+This allowed us to estimate the size and the life time of the convective cells at the surface.
+From our images, we computed a quantity called the photo-center, which is sensitive to the size and the number of convective cells.
+An homogeneous star will have a photo-center displacement close to the barycenter of the star, whereas a star with one or two huge convective cells will have a photo-center displacement more important, up to a few percent of the stellar radius.
+Because the photo-center is linked to surface convection, our idea was to compute the LS periodogram of the displacement of the photo-center.
+If we recover the different periods from our images, we would be able to explain them with only convection.
+However, before going further, we have to be carefull with the interpretation of our images.
+Linear polarization suffers from a $180 ^\circ$ degree ambiguity already mentioned in \cite{auriere_discovery_2016}.
+For that reason, our images can be rotated by $180^\circ$ degree, and the brightness distribution will still fit the observed LSD profile.
+Thus, to obtain a continuity between our images, for a given day, we use the brightness distribution of the previous day as the starting point of the fit.
+For the first image, we start with a random brightness distribution as the starting point.
+Even though we are able to produce images which are consistent with each other, it is important to keep in mind that the photo-center displacement will be affected by the choice of the first image.
+To overcome this issue, we decided to compute the photo-center displacement with 100 different initial parameters.
+So that we recover properties that are independent of the choice of the first image.
+
+\begin{figure}[!h]
+    \centering
+    \includegraphics[width=0.5\textwidth]{Lomb-Scargle Photo-center.png}
+    \caption{Lomb-Scargle periodogram of the 100 photo-center displacement of Betelgeuse.
+Each blue curve corresponds to a LS periodogram of one photo-center displacement.
+The black line is the average of the 100 LS periodogram.
+The black dashed lines represent the 2000, 400 and 200 days period respectively (from left to right).
+The upper panel is the LS from data before the dimming.
+The lower panel is the LS from the all dataset, including data after the great dimming.
+}
+    \label{LS photocenter}
+\end{figure}
+
+Figure \ref{LS photocenter} shows the LS periodogram of each photo-center displacement (blue lines) and the average photo-center displacement (black line).
+The upper panel is the LS from data before the great dimming of Betelgeuse, whereas the lower panel is the LS for the all set of data, including after the dimming.
+The black dashed lines are the 2000, 400 and 200 days period respectively.
+First, we find the the LS power is high around the LSP.
+We can also notice a decrease in the Lomb-Scargle power before and after the dimming.
+It appears that there are two smalls bumps around the 400 days period.
+Those bumps almost disappear after the dimming.
+Even though it is hard to attribute the 400 days variability to those bumps, from polarimetry imaging, we still find that there is a difference before and after the dimming.
+However, in our case, our images only involve surface convection.
+The photo-center displacement is pointing toward a variability explained by surface convection.
+
 \section{Convection to explain periodicities}
 
+\subsection{The 400 days period explained by rising plumes of plasma}
+
 In the previous section, we have seen that we recover the different period of Betelgeuse from the intensity and linear polarization signal. 
-While the nature of the LSP is still in debate, the nature of the lower periods is attributed to pulsation of the atmosphere of Betelgeuse 
-(e.g \cite{Kiss2006}). The mechanism behind the pulsation of Betelgeuse is attributed to the $\kappa$-mechanism. In this section, we attempt to 
+While the nature of the LSP is still in debate, the nature of the lower periods is attributed to pulsations of the pressure modes 
+(e.g \cite{kiss_variability_2006}). The mechanism behind the pulsation of Betelgeuse is attributed to the $\kappa$-mechanism. In this section, we attempt to 
 find an explanation to the 400 days periodicity of Betelgeuse. As mentioned before, the LS periodogram of linear polarization shows a periodicity 
-around 400 days. Furthermore, the region where the LS power is high around 40km/s. This wavelength correspond to the red wing of Betelgeuse.
+around 400 days. Furthermore, the region where the LS power is high is around 40km/s. This wavelength corresponds to most redshifted signal present in Betelgeuse.
 The polarization signal is supposed to stand between -20 and 40km/s. Signal above 40km/s is associated to plasma rising behind the limb of Betelgeuse. 
-This mechanism has been observed in the RSG $\mu$Cep and was introduced by \cite{LopezAriste2023} to explain the excess of linear polarization beyond 
-the velocity of the star. In $\mu$Cep, the linear polarization signal is very strong and often present, whereas in Betelgeuse, the signal above the 
-velocity of the star is less strong. However, because it is present from time to time, the LS periodogram finds this signal above the limb and finds a
-periodicity at this wavelength of 40 km/s. Hence, the strong power associated to this wavelength might be due to those convective plumes that are 
+This scenario has been observed in the RSG $\mu$ Cep and was introduced by \cite{lopez_ariste_height_2023} to explain the excess of linear polarization beyond 
+the velocity of the star. In $\mu$ Cep, the linear polarization signal is very strong and often present, whereas in Betelgeuse, the signal above the 
+velocity of the star is less strong. However, because it is present from time to time, the LS periodogram finds this signal above the limb and interprets it as a
+periodic signal at this wavelength of 40 km/s. Hence, the strong power associated to this wavelength might be due to those convective plumes that are 
 rising above the limb from time to time. From this scenario, we can explained the 400 days period in the light curve as follow: from time to time,
@@
 for the moment. Thus, the LS periodogram, at that precise frequency, finds that there is a periodic signal before the dimming, but not after, leading to
-a lower LS power at this frequency. Concerning the 200 days period, it is caused by the timescale on which the stokes parameters evolve. For example, the 
+a lower LS power at this frequency. Concerning the 200 days period, it is caused by the timescale on which the stokes parameters evolve. For instance, the 
 LSD profile of stokes Q or U will slightly change if we observe Betelgeuse every two weeks. However, on a longer timescale, for example a few month, 
@@
 of convective cells. Some convective cells can live up to a few month, leading to this peak in the LS periodogram of stokes Q. As showed by
-\cite{LopezAriste2018}, some convective cells can live up to years, explaining also the LSP. One can notice that the peak is not present in stokes U. 
+\cite{lopez_ariste_convective_2018}, some convective cells can live up to years, explaining also the LSP. One can notice that the peak is not present in stokes U. 
 However, if we try to explain the periodicity of Betelgeuse using pulsations, we would see this signal in either stokes U and the linear polarization. 
-To summarise what we have done so far: we were able to explain the variability of Betelgeuse by using surface convection only. The convective cells play 
+To summarise what we have done so far: we were able to explain the variability of Betelgeuse using surface convection only. The convective cells play 
 the most important role in the variability of Betelgeuse. Our work does not need pulsation to explain its variability. This does not mean that there are 
@@
 cells evolve. Pulsation could exist on Betelgeuse, but probably on shorter timescale than the one observed.
+
+\subsection{The light curve of Betelgeuse}
+
+In the literature, it is reported that before the great dimming, the main period of Betelgeuse was 400 days \citep{kiss_variability_2006}.
+This pulsation is associated to the fundamental model of pressure modes.
+After the dimming, this period is shorter and is now around 200 days, associated to the first overtone of the fundamental mode.
+However, there are no explanations concerning this change of variability due to the dimming.
+Using the data from AAVSO, we studied the light curve of Betelgeuse.
+Figure \ref{light curve Betelgeuse} shows the light curve of Betelgeuse in the visible, during and after the great dimming.
+After the dimming, the variability of Betelgeuse has changed.
+It is clear from fig.
+\ref{light curve Betelgeuse} that the variability of Betelgeuse is evolving on a timescale shorter than one year.
+However, from middle 2023 up to now, the magnitude of Betelgeuse remains almost constant, which is not consistent with a pulsating scenario.
+The constant magnitude of Betelgeuse points toward another scenario than pulsation.
+Using surface convection, we can fully explain the variability of Betelgeuse.
+The oscillations in the light curve are due to two mechanisms: surface convection and rising plumes of plasma beyond the limb of the star.
+This last mechanism being random, it happens from time to time.
+The timescale on which convective cells evolve is between one and two years, which is also consistent with the 400 days period found in the literature.
+After the dimming, the change of periodicity in the light curve is explained by surface convection, the random variation of brightness distribution could lead to the 200 days periodicity found in the literature.
+The constant brightness of Betelgeuse for the last 6 months is explainable by the same random brightness distribution: the surface of Betelgeuse is calm, leading to this constant brightness.
+In the future, we expect to see the light curve of Betelgeuse to change, maybe with another main period, but which is in fact caused by surface convection.
+Thus, the main mechanism in the variability of Betelgeuse is surface convection, which is sufficient to explain the variation in the magnitude of Betelgeuse.
+
+\begin{figure}[!h]
+    \centering
+    \includegraphics[width=0.5\textwidth]{Light_curve_Betelgeuse.png}
+    \caption{Light curve of Betelgeuse from AAVSO in the visible band.}
+    \label{light curve Betelgeuse}
+\end{figure}
+
+\section{Conclusion}
 

### 2024-02-08 main.tex
modified:   main.tex
modified:   art75.bib 	modified:   main.tex

@@
 Trying to find periods in an astrophysical context can be difficult due to the unevenly spaced observed data. To overcome this issue, 
-we used the Lomb-Scargle (LS) periodogram \citep{lomb_least-squares_1976,scargle_studies_1982}. For a given set of data, it fits a sinusoid function using least-squares at each 
-frequency between the first and the last observation. The better the fit, the higher the power attributed to the given frequency. Since Betelgeuse has 
-been observed by the Télescope Bennard Lyot at pic du midi since 2013, we have ten years of data. Betelgeuse was observed every two or three weeks by 
-the spectropolarimeters Narval and Neo-Narval, except during summer.
-From all these data, we can compute the LSD profiles of Stokes I, U and Q.
-The linear polarization present in Betelgeuse is attributed to brightness
-inhomogeneities, which are due to the convective cells present at the surface \citep{lopez_ariste_convective_2018}. To recover variability of Betelgeuse, 
+we used the Lomb-Scargle (hereafter LS) periodogram \citep{lomb_least-squares_1976,scargle_studies_1982}. For a given set of data, it fits sinusoid functions 
+using least-squares at each frequency between the first and the last observation. The better the fit, the higher the power attributed to the frequency. 
+Since Betelgeuse has been observed by the Télescope Bernard Lyot since 2013, we have ten years of data. Betelgeuse was observed every two or three weeks by 
+the spectropolarimeters Narval and Neo-Narval, except during summer. From all these data, we computed the LSD profiles of Stokes I, U and Q. 
+The linear polarization present in Betelgeuse is attributed to brightness inhomogeneities, which are due to the convective cells 
+present at the surface \citep{lopez_ariste_convective_2018}. Hence, if we are able to find periodicities in the LS periogogram of the LSD profiles of Betelgeuse,
+the periodicities would be associated to surface convection.
+
+To recover variability of Betelgeuse, 
 we computed the LS periodogram at each wavelength of the LSD profile, between -25 km/s and 80 km/s. We choose those velocity to span the entire 
 signal present in the LSD profile. In the work of \cite{lopez_ariste_convective_2018}, the most blueshifted signal was present at -20 km/s, 
-corresponding to the maximal velocity of the rising plasma. The most redshifted signal was $\sim$ 40 km/s. This velocity corresponds 
-to the velocity of the star in our reference frame.
+corresponding to the maximal velocity of the rising plasma. The most redshifted signal was 40 km/s. This velocity corresponds 
+to the velocity of the star in our reference frame. This last velocity is very important in our study. Signal above 40 km/s is attributed to either huge
+convective cells falling back to Betelgeuse, or to rising plumes of plasma located in the hidden face of Betelgeuse,
+that rise so high that they appear above the limb.
 
@@
 Interferometric images show a huge drop in luminosity in the southern hemisphere of Betelgeuse, leading to this dimming. Interestingly, 
-it has been shown by \cite{jadlovsky_analysis_2023} that the periodicity of Betelgeuse has changed before and after the dimming. Using the light curves from AAVSO, 
+it has been shown by \cite{jadlovsky_analysis_2023} and \cite{dupree_great_2022} that the periodicity of Betelgeuse has changed after the dimming. Using the light curves from AAVSO, 
 the authors showed that before the dimming, the dominant period of Betelgeuse was around 400 days. However, after the dimming, this period has changed and 
 is now shorter, around 200 days. It means that, after the dimming, there was a change in the behaviour of the atmosphere of Betelgeuse: the period is now
-shorter than before the dimming. Our analysis is based on this change of periodicity. By computing the Lomb-Scargle periodogram for each frequency 
-in the LSD profile, we expect to find periodicities around the LSP, but also on the 400 days and 200 days period. However, we expect a change
-in the periodicity after the dimming. This would lead to a decrease in the Lomb-Scargle power. 
+shorter than before the dimming. Our analysis is based on this change of periodicity.
+
+By computing the LS periodogram for each frequency 
+in the LSD profile, we expect to find periodicities around the LSP, but also on the 400 days and 200 days period. If the LS periodogram finds a periodicity at a
+given wavelength, it would appear as a strong LS power. However, we expect a change in the periodicity after the dimming. 
+This would lead to a decrease in the Lomb-Scargle power. By comparing LS periodogram before and after the dimming, we expect to see a change in the LS power,
+especially at the 400 days period.
+
 Figure \ref{LS intensity},\ref{LS Q},\ref{LS U} and \ref{LS linear polarization} show the Lomb-Scargle periodogram of the LSD profile of respectively 
-the intensity, stokes Q, stokes U and the linear polarization. The abscissa is the frequency, while the ordinate is the heliocentric radial velocity. 
-For each figure, the upper panel is the LS periodogram with data taken before the great dimming of Betelgeuse, that is from 2013 to 2019. 
+the intensity, stokes Q, stokes U and the linear polarization. The abscissa is the frequency (in $\mathrm{days^{-1}}$), while the ordinate is the heliocentric radial velocity in km/s. 
+For each figure, the upper panel is the LS periodogram with data taken before the great dimming of Betelgeuse, that is from 2013 to the beginning of 2019. 
 The lower panel is the LS periodogram for the data before and after the dimming: from 2013 to 2023. The white dashed lines represent the different 
-periods found in the literature: the LSP at 2000 days, the fondamental pressure mode at 400 days and its first overtone at 200 days. The Lomb-Scargle 
+periods found in the literature: the LSP at 2000 days, the fundamental pressure mode at 400 days and its first overtone at 200 days. The Lomb-Scargle 
 power is normalised. The colorbar indicates the quality of the fit: red areas correspond to good fits from the LS periodogram.
@@
     The upper pannel is the LS periodogram for data before the great dimming while the lower panel is the periodogram from the all dataset.
-The ordinate refers to the velocity in the heliocentric radial velocity in km/s.
-The abscissa is the period in $\mathrm{days^{-1}}$.
+    The ordinate refers to the velocity in the heliocentric radial velocity in km/s. The abscissa is the period in $\mathrm{days^{-1}}$. 
     The three white dashed lines represent respectively (from left to right) the LSP at 2000 days, the 400 days period and its first overtone at 200 days. 
-The Lomb-Scargle power refers to the quality of the fit from the LS periodogram.
-}
+    The Lomb-Scargle power refers to the quality of the fit from the LS periodogram. }
     \label{LS intensity}
@@
     \includegraphics[width=0.5\textwidth]{Lomb-Scargle Stokes Q.png}
-    \caption{Same as figure \ref{LS intensity} but for stokes Q. }
+    \caption{Same as figure \ref{LS intensity} for stokes Q. }
     \label{LS Q}
@@
     \includegraphics[width=0.5\textwidth]{Lomb-Scargle Stokes U.png}
-    \caption{Same as figure \ref{LS intensity}, but for stokes U.}
+    \caption{Same as figure \ref{LS intensity} for stokes U.}
     \label{LS U}
@@
     \includegraphics[width=0.5\textwidth]{Lomb-Scargle linear polarization.png}
-    \caption{Same as figure \ref{LS intensity}, but for the linear polarization: $\sqrt{\mathrm{Q^2}+\mathrm{U^2}}$.}
+    \caption{Same as figure \ref{LS intensity} for the linear polarization: $\sqrt{\mathrm{Q^2}+\mathrm{U^2}}$.}
     \label{LS linear polarization}
@@
 in the intensity profile are consistent with the one from \cite{mathias_evolution_2018} who used 2D fourier analysis and found that the dominant period was the LSP.
-We can also see that there is a less stronger signal at 40 km/s in the LSP. We will discuss about it in the next section, but this weak 
-signal at 40 km/s, which is the velocity of the star in the heliocentric reference frame, could be a clue to understand the periodicity of Betelgeuse.
 
@@
 Figure \ref{LS Q} and \ref{LS U} show the LS periodogram of both stokes Q and U. It is clear that the LS periodogram is completely different 
-from the intensity one. We can see a strong signal in fig. \ref{LS Q} around 200 days. Once again, this signal is weaker after the dimming.
+from the intensity one. A strong signal is present in fig. \ref{LS Q} around 200 days. Once again, this signal is weaker after the dimming.
 Concerning stokes U in fig. \ref{LS U}, it appears that we only recover the LSP, which once again, is weaker after the great dimming. 
 Those periodograms rise a lot of questions: why is the 200 days period only present in stokes Q? Why not in stokes U or intensity? 
-We will discuss about those questions in the next question. However, it also appears that stokes Q is not sensitive to the LSP, or to be more precise, 
+We will discuss about those questions in the next section. However, it also appears that stokes Q is not sensitive to the LSP, or to be more precise, 
 it is much more sensitive to the 200 days period. From this, we can conclude that stokes Q evolves on a time scale much more closer to the first overtone 
 mode than the LSP or even the fundamental mode. Because stokes Q is sensitive only to shorter periods, the variation of stokes Q come from another mechanism 
-than the one from intensity. Because the linear polarization in Betelgeuse is due to convective cells, we attribute the signal present in the periodogram 
+than the one from intensity. Linear polarization in Betelgeuse is due to convective cells, we attribute the signal present in the periodogram 
 as the lifetime of convective cells. This is what we observe in the LSD profile of Betelgeuse: from a week to another it does not change that much. 
@@
 While the LSP is already present in the LS of intensity, here we recover the 400 days periodicity, often invoke in the literature. 
-Nonetheless, the period is present at a HRV of around 40 km/s. We recall that this velocity is the velocity of the star \cite{lopez_ariste_convective_2018}. 
-Every signal above the velocity of the star is either due to the plasma sinking in Betelgeuse, or plumes of plasma rising above the limb of the star, 
+Nonetheless, the period is present at a HRV of around 40 km/s. We recall that this velocity is the velocity of the star. 
+Every signal above this velocity is either due to the plasma sinking in Betelgeuse, or plumes of plasma rising above the limb of the star, 
 as seen in other red supergiant like $\mu$ Cep \citep{lopez_ariste_height_2023}. Up to this point, we want to pinpoint the two elements that are a clue 
 to understand the variability of Betelgeuse: the 400 days period is present in the linear polarization signal and also at a HRV above 40 km/s.
-In any case, from the linear polarization and the intensity profile, we recover the different periods of Betelgeuse.
+In any case, from the linear polarization and the intensity profile, we recover the different periods of Betelgeuse. Since linear polarization is 
+associated to convection and because we recover the periods of Betelgeuse, the main mechanism behind the variability of Betelgeuse points toward surface convection.
 
@@
 
-Using linear polarization, \cite{lopez_ariste_convective_2018} has been able to reconstruct images of Betelgeuse that have been sucessfully compared to inteferometric images of \cite{montarges_close_2016}.
-The images are produced by finding the brightness distribution that better fits the observed linear polarization LSD profile using Marquardt-Levemberg minimisation \citep{lopez_ariste_convective_2018}.
-Because Betelgeuse is observed every two weeks by the TBL, we were able to follow its surface activity through polarimetry imaging.
-This allowed us to estimate the size and the life time of the convective cells at the surface.
-From our images, we computed a quantity called the photo-center, which is sensitive to the size and the number of convective cells.
-An homogeneous star will have a photo-center displacement close to the barycenter of the star, whereas a star with one or two huge convective cells will have a photo-center displacement more important, up to a few percent of the stellar radius.
-Because the photo-center is linked to surface convection, our idea was to compute the LS periodogram of the displacement of the photo-center.
-If we recover the different periods from our images, we would be able to explain them with only convection.
-However, before going further, we have to be carefull with the interpretation of our images.
-Linear polarization suffers from a $180 ^\circ$ degree ambiguity already mentioned in \cite{auriere_discovery_2016}.
-For that reason, our images can be rotated by $180^\circ$ degree, and the brightness distribution will still fit the observed LSD profile.
-Thus, to obtain a continuity between our images, for a given day, we use the brightness distribution of the previous day as the starting point of the fit.
-For the first image, we start with a random brightness distribution as the starting point.
-Even though we are able to produce images which are consistent with each other, it is important to keep in mind that the photo-center displacement will be affected by the choice of the first image.
-To overcome this issue, we decided to compute the photo-center displacement with 100 different initial parameters.
-So that we recover properties that are independent of the choice of the first image.
+Using linear polarization, \cite{lopez_ariste_convective_2018} have been able to reconstruct images of Betelgeuse that have been sucessfully compared 
+to inteferometric images of \cite{montarges_close_2016}. The images are produced by finding the brightness distribution that better fits the observed linear 
+polarization LSD profile using Marquardt-Levemberg minimisation.
+
+Betelgeuse is observed every two weeks by the TBL, hence,
+we were able to follow its surface activity through polarimetry imaging. This allowed us to estimate the size and the life time of the convective cells
+at the surface. From our images, we computed a quantity called the photo-center, which is sensitive to the size and the number of convective cells. 
+An homogeneous star will have a photo-center displacement close to the barycenter of the star, whereas a star with one or two huge convective cells will have a
+photo-center displacement more important, up to a few percent of the stellar radius in the case of RSG \citep{chiavassa_probing_2022}.
+
+Since the photo-center is linked to surface convection, 
+we computed the LS periodogram of the displacement of the photo-center. If we recover the different periods from our images, we would
+be able to explain them by involving only convection, since our imaging model does not involve pulsations. However, before going further, we have to be carefull with the interpretation of our images. 
+Linear polarization suffers from a $180 ^\circ$ degree ambiguity already mentioned in \cite{auriere_discovery_2016}. For that reason, our images can be 
+rotated by $180^\circ$ degree, and the brightness distribution will still fit the observed LSD profile. Thus, to obtain a continuity between our images, 
+for a given day, we use the brightness distribution of the previous day to start fitting the LSD profile.
+This is mandatory to obtain a continuity between the images.
+For the first image, we start with a
+random brightness distribution as the beginning of the fit. Even though we are able to produce images which are consistent with each other, it is important 
+to keep in mind that the photo-center displacement will be affected by the choice of the first image. To overcome this issue, we decided to compute
+the photo-center displacement from 100 different initial parameters. So that we recover properties that are independent of the choice of the first image.
 
@@
     \includegraphics[width=0.5\textwidth]{Lomb-Scargle Photo-center.png}
-    \caption{Lomb-Scargle periodogram of the 100 photo-center displacement of Betelgeuse.
-Each blue curve corresponds to a LS periodogram of one photo-center displacement.
-The black line is the average of the 100 LS periodogram.
-The black dashed lines represent the 2000, 400 and 200 days period respectively (from left to right).
-The upper panel is the LS from data before the dimming.
-The lower panel is the LS from the all dataset, including data after the great dimming.
-}
+    \caption{Lomb-Scargle periodogram of the 100 photo-center displacement of Betelgeuse. Each blue curve corresponds to a LS periodogram of one photo-center
+     displacement. The black line is the average of the 100 LS periodogram. The black dashed lines represent the 2000, 400 and 200 days period respectively 
+     (from left to right). The upper panel is the LS from data before the dimming. The lower panel is the LS from the all dataset, 
+     including data after the great dimming. }
     \label{LS photocenter}
@@
 
-Figure \ref{LS photocenter} shows the LS periodogram of each photo-center displacement (blue lines) and the average photo-center displacement (black line).
-The upper panel is the LS from data before the great dimming of Betelgeuse, whereas the lower panel is the LS for the all set of data, including after the dimming.
-The black dashed lines are the 2000, 400 and 200 days period respectively.
-First, we find the the LS power is high around the LSP.
-We can also notice a decrease in the Lomb-Scargle power before and after the dimming.
-It appears that there are two smalls bumps around the 400 days period.
-Those bumps almost disappear after the dimming.
-Even though it is hard to attribute the 400 days variability to those bumps, from polarimetry imaging, we still find that there is a difference before and after the dimming.
-However, in our case, our images only involve surface convection.
-The photo-center displacement is pointing toward a variability explained by surface convection.
+Figure \ref{LS photocenter} shows the LS periodogram of each photo-center displacement (blue lines) and the average LS periodogram of the photo-center displacements (black line). 
+The upper panel is the LS from the images before the great dimming of Betelgeuse, whereas the lower panel is the LS for the whole set of data, 
+including after the dimming. The black dashed lines are the 2000, 400 and 200 days period respectively. First, we find the the LS power is high around the LSP.
+We also notice a decrease in the Lomb-Scargle power before and after the dimming. It appears that there are two smalls bumps around the 400 days period. 
+Those bumps almost disappear after the dimming. Even though it is hard to attribute the 400 days variability to those bumps, from polarimetry imaging,
+we still find that there is a difference before and after the dimming. However, in our case, our images only involve surface convection.
+The LS of the photo-center displacement point toward a variability explained by convection only. The photo-center only is not enough to attribute the variability
+of Betelgeuse to surface convection, but what we found from polarimetry imaging is consistent with what we obtained from linear polarization.
+Both the photo-center displacement and the LS of the linear polarization point toward a variability explained by surface convection.
 
-\section{Convection to explain periodicities}
+\section{Surface convection to explain variability}
 
@@
 around 400 days. Furthermore, the region where the LS power is high is around 40km/s. This wavelength corresponds to most redshifted signal present in Betelgeuse.
-The polarization signal is supposed to stand between -20 and 40km/s. Signal above 40km/s is associated to plasma rising behind the limb of Betelgeuse. 
+
+The polarization signal is supposed to remain between -20 and 40km/s. Signal above 40km/s is associated to plasma rising behind the limb of Betelgeuse. 
 This scenario has been observed in the RSG $\mu$ Cep and was introduced by \cite{lopez_ariste_height_2023} to explain the excess of linear polarization beyond 
@@
 periodic signal at this wavelength of 40 km/s. Hence, the strong power associated to this wavelength might be due to those convective plumes that are 
-rising above the limb from time to time. From this scenario, we can explained the 400 days period in the light curve as follow: from time to time,
+rising above the limb from time to time. From this scenario, we can explained the 400 days period in the light curve. From time to time,
 Betelgeuse is ejecting plasma in the interstellar medium due to winds. The plasma can be ejected from the hidden face of Betelgeuse. 
-It rises so high that we can see the rising plume from the Earth, because the plume rises above the limb of the star. Hence, from Earth, 
+It rises so high that we see the rising plume from the Earth, because the plume rises above the limb of the star, contributing to the linear polarization signal. Hence, from Earth, 
 the signal associated to this plume appears redshifted compared to the velocity of the star. This plume increases the number of photons that we receive,
-leading to an increase of the brightness. After a few weeks or months, the plume falls back to Betelgeuse, desapearing from our point of view, 
-leading to a decrease of the brightness of Betelgeuse. This is exactly what we see is the LSD profile of the linear polarization of Betelgeuse: 
+leading to an increase of the brightness. After a few weeks or months, the plume falls back to Betelgeuse, disappearing from our point of view, 
+leading to a decrease of the brightness.
+This is exactly what we see in the LSD profile of the linear polarization of Betelgeuse: 
 from time to time, there is a net signal above 40km/s, which is present for a few weeks, and then disappears. Since this signal is not always present, 
 it is found by the LS periodogram as a periodic event. However, it is not periodic at all: because the plumes rise at random moments, sometimes they 
-appear every years, sometime more often. This would explain why we only see it before the dimming, after the dimming, this kind of event never happened 
+appear every years, sometime they do not. This would explain why we only see it before the dimming. After the dimming, this kind of event never happened 
 for the moment. Thus, the LS periodogram, at that precise frequency, finds that there is a periodic signal before the dimming, but not after, leading to
-a lower LS power at this frequency. Concerning the 200 days period, it is caused by the timescale on which the stokes parameters evolve. For instance, the 
+a lower LS power at this frequency.
+
+Concerning the 200 days period, it is caused by the timescale on which the stokes parameters evolve. For instance, the 
 LSD profile of stokes Q or U will slightly change if we observe Betelgeuse every two weeks. However, on a longer timescale, for example a few month, 
@@
 of convective cells. Some convective cells can live up to a few month, leading to this peak in the LS periodogram of stokes Q. As showed by
-\cite{lopez_ariste_convective_2018}, some convective cells can live up to years, explaining also the LSP. One can notice that the peak is not present in stokes U. 
+\cite{lopez_ariste_convective_2018}, some convective cells can live up to years, explaining also the LSP.
+One can notice that the peak is not present in stokes U, which might be due to stochastic event.
 However, if we try to explain the periodicity of Betelgeuse using pulsations, we would see this signal in either stokes U and the linear polarization.
+
 To summarise what we have done so far: we were able to explain the variability of Betelgeuse using surface convection only. The convective cells play 
@@
 no pulsation in Betelgeuse. What we claim is that the most important factor in the variability of Betelgeuse is the timescale on which the convective 
-cells evolve. Pulsation could exist on Betelgeuse, but probably on shorter timescale than the one observed.
+cells evolve. Pulsation could exist on Betelgeuse, but they are not the main contributor to the variability.
 
-\subsection{The light curve of Betelgeuse}
+\subsection{The variability of the light curve of Betelgeuse}
 
 In the literature, it is reported that before the great dimming, the main period of Betelgeuse was 400 days \citep{kiss_variability_2006}. 
-This pulsation is associated to the fundamental model of pressure modes.
-After the dimming, this period is shorter and is now around 200 days, associated to the first overtone of the fundamental mode.
-However, there are no explanations concerning this change of variability due to the dimming.
-Using the data from AAVSO, we studied the light curve of Betelgeuse.
-Figure \ref{light curve Betelgeuse} shows the light curve of Betelgeuse in the visible, during and after the great dimming.
-After the dimming, the variability of Betelgeuse has changed.
+This pulsation is associated to the fundamental model of pressure modes. After the dimming, this period is shorter and is now around 200 days, 
+associated to the first overtone of the fundamental mode. However, there are no explanations concerning this change of variability due to the dimming.
+
+Using the data from AAVSO, we studied the light curve of Betelgeuse. Figure \ref{light curve Betelgeuse} shows the light curve of Betelgeuse in the visible, 
+from 2014 up to now.
+Before the great dimming, the magnitude of Betelgeuse evolves on a timescale of a year, whereas after the dimming, the variability of Betelgeuse has changed.
 It is clear from fig.
-\ref{light curve Betelgeuse} that the variability of Betelgeuse is evolving on a timescale shorter than one year.
-However, from middle 2023 up to now, the magnitude of Betelgeuse remains almost constant, which is not consistent with a pulsating scenario.
-The constant magnitude of Betelgeuse points toward another scenario than pulsation.
-Using surface convection, we can fully explain the variability of Betelgeuse.
-The oscillations in the light curve are due to two mechanisms: surface convection and rising plumes of plasma beyond the limb of the star.
-This last mechanism being random, it happens from time to time.
-The timescale on which convective cells evolve is between one and two years, which is also consistent with the 400 days period found in the literature.
-After the dimming, the change of periodicity in the light curve is explained by surface convection, the random variation of brightness distribution could lead to the 200 days periodicity found in the literature.
-The constant brightness of Betelgeuse for the last 6 months is explainable by the same random brightness distribution: the surface of Betelgeuse is calm, leading to this constant brightness.
-In the future, we expect to see the light curve of Betelgeuse to change, maybe with another main period, but which is in fact caused by surface convection.
-Thus, the main mechanism in the variability of Betelgeuse is surface convection, which is sufficient to explain the variation in the magnitude of Betelgeuse.
+\ref{light curve Betelgeuse} that the
+variability of Betelgeuse is evolving on a timescale shorter than one year. However, from middle 2023 up to now, the magnitude of Betelgeuse remains almost 
+constant, which is not consistent with a pulsating scenario. The constant magnitude of Betelgeuse points toward another scenario than pulsation. Using surface 
+convection, we can fully explain the variability of Betelgeuse. The oscillations in the light curve are due to two mechanisms: surface convection and rising 
+plumes of plasma beyond the limb of the star. This last mechanism being random, it happens from time to time. The timescale on which convective cells evolve
+is between one and two years, which is also consistent with the 400 days period found in the literature.  After the dimming, the change of periodicity in the
+light curve is explained by surface convection, the random variation of brightness distribution could lead to the 200 days periodicity found in the literature. 
+The constant brightness of Betelgeuse for the last 6 months is explainable by the same random brightness distribution: the surface of Betelgeuse is calm, 
+leading to this constant brightness. In the future, we expect to see the light curve of Betelgeuse to change, maybe with another main period, but which 
+is in fact caused by surface convection. Thus, the main mechanism in the variability of Betelgeuse is surface convection, which is sufficient to explain the 
+variation in the magnitude of Betelgeuse.
 
@@
     \includegraphics[width=0.5\textwidth]{Light_curve_Betelgeuse.png}
-    \caption{Light curve of Betelgeuse from AAVSO in the visible band.}
+    \caption{Light curve of Betelgeuse from AAVSO in the V-band.}
     \label{light curve Betelgeuse}

### 2024-02-27 main.tex
a effacer
Changes by ALA

@@
 periods around 400 and 200 days. Such periods are often linked to radial pulsation modes,   and have been used to model 
-the caracteristics of the resonant cavity, hence the size of the star and eventually its evolutionary stage \cite{XXX}.
+the characteristics of the resonant cavity, hence the size of the star and eventually its evolutionary stage \cite{XXX}.
 
-At the end of 2019, Betelgeuse suffered a sudden and dramatic loss of brightness \cite{}, particularly in the blue colors. 
-\cite{Miguel} has demonstrated that such event was due to the emission by the star of a cloud of dust near to the line of sight of 
-Earth. In this event it was clear that a change in brightness was not due to any kind of pulsation but to a change on the 
+At the end of 2019, Betelgeuse reached a historical minimum in its luminosity, called the great dimming \citep{guinan_fall_2020}. 
+From interferometric data, we know that this event was caused by a cloud of dust close to the line of sight \citep{montarges_dusty_2021}. 
+Interferometric images show a huge drop in luminosity in the southern hemisphere of Betelgeuse, leading to this dimming.
+In this event it was clear that a change in brightness was not due to any kind of pulsation but to a change on the
 brightness distribution over the stellar disk due to the presence of large quantities of dust. While this may be seen as 
 a singular event, it brings up the question whether all changes in brightness cannot be due to similar changes in the brightness 
-distribution over the disk.
+distribution over the disk, unrelated to any pulsation phenomenon.
 
@@
 compared with other situations where such cells are found near the edges. So even without calling for the formation of 
-dust or other changes in opacity, one may argue that the simple convective patterns of the star may create a variability. Such 
+dust or other changes in opacity, one may argue that the simple evolution of convective patterns of the star may create a variability. Such 
 variability could be expected to be random, but will present quasi-periodicities related to the typical time scales of the 
@@
 
-Since the end of the Great Dimming event, Betelgeuse has continue its random variation of brightness. But the classical periods 
-of variation cited above are not visible any more. This may be due to simply the short span of time, that does not afford for 
-the peaks to appear prominently in the variability spectra. But it has been sufficient to bring up that other explanation of the variability
-in terms  of convective patterns. In order to address this question we seek for the typical periods of 1000, 400 and 200 days in  
-observational proxies related to the convective activite but unrelated to pulsations nor directly linked to brightness?
+Since the end of the Great Dimming event, Betelgeuse has continue its random variation of brightness. 
+Interestingly, 
+it has been shown by \cite{jadlovsky_analysis_2023} and \cite{dupree_great_2022} that the periodicity of Betelgeuse has changed after the dimming. Using the light curves from AAVSO, 
+the authors showed that before the dimming, the dominant period of Betelgeuse was around 400 days. However, after the dimming, this period has changed and 
+is now shorter, around 200 days. It means that, after the dimming, there was a change in the behaviour of the atmosphere of Betelgeuse: the period is now
+shorter than before the dimming. Our analysis is based on this change of periodicity. 
+Faced with the dramatic change in stellar parameters and age that such change in a pulsation period would 
+require, our interest in an alternative explanation of the variability
+in terms  of convective patterns arose. In order to address this question we shall seek for such typical periods of 1000, 400 and 200 days in  
+observational proxies related to the convective activite but unrelated to pulsations and which cannot be directly linked to brightness?
 
@@
 different lines in the spectrum of Betelgeuse. The images produced compare very well with cotemporaneous images made with interferometric 
-techniques and show clear convective patterns. Linear polarization in the atomic lines is therefore a proxy of convection, unrelated to radial 
+techniques and show clear convective patterns, akin to solar granulation.
+
+Linear polarization observed in the atomic lines is therefore a proxy of convection, unrelated to radial 
 pulsations but linked to the brightness inhomogeneities due to the convective patterns in Betelgeuse. If one can find the aforementioned periods 
-in linear polarization, we must conclude that these periods are related to the convective activity which originates those linear polarization 
-and unrelated to any radial pulsation phenomena.
+in linear polarization, we must conclude that these periods are related to the convective activity which originates those linear polarization signals
+and unrelated to any radial pulsation phenomena. This is the purpose of the present work.
 
-\section{Method: Searching for periodicities in the polarization spectra of Betelgeuse}
+\section{Searching for periodicities in the polarization spectra of Betelgeuse}
 
-Betelgeuse has been observed for the last 13 years with Narval and Neo-Narval at the Telescope Bernard Lyot. These instruments measure the polarization 
+Betelgeuse has been observed for the last 13 years with Narval and Neo-Narval at the Telescope Bernard Lyot. These two instruments measure the polarization 
 over the visible spectra of Betelgeuese (390-1000 nm) with high spectral resolution (R=65000) and high polarimetric sensitivity. Despite 
-this sensitivity, the signal-to-noise ratios are not sufficient to measure the weak polarization signals in the individual atomic lines of the 
+this high sensitivity, the signal-to-noise ratios are not sufficient to measure the weak polarization signals in individual atomic lines of the 
 spectrum. These amplitudes are today known to be of the order of $10^{-4}$ times the continuum intensity, and only excepcionnaly do they 
@@
 and linear polarization which does not belong to any atomic species in particular but to an average of all of them present and emitting in the 
-photosphere of Betelgeuse. These signals carry the coherent signals present in the photospheric lines, but the particularities of this or 
+photosphere of Betelgeuse. After such procedures of line addition, these signals carry the coherent signals present in the photospheric lines, but the particularities of this or 
 that spectral line are erased. All these data has been presented before by \cite{auriere_discovery_2016}, \cite{mathias_evolution_2018} and 
@@
 
-\subsection{Lomb-Scargle periodogram}
+%\subsection{Lomb-Scargle periodograms of the profiles}
 
-Trying to find periods in an astrophysical context can be difficult due to the unevenly spaced observed data. To overcome this issue, 
+Trying to find periods in an astrophysical context carries the hurdle of unevenly spaced observed data. To overcome this issue, 
 we used the Lomb-Scargle (hereafter LS) periodogram \citep{lomb_least-squares_1976,scargle_studies_1982}. For a given set of data, it fits sinusoid functions 
 using least-squares at each frequency between the first and the last observation. The better the fit, the higher the power attributed to the frequency. 
-Since Betelgeuse has been observed by the Télescope Bernard Lyot since 2013, we have ten years of data. Betelgeuse was observed every two or three weeks by 
-the spectropolarimeters Narval and Neo-Narval, except during summer. From all these data, we computed the LSD profiles of Stokes I, U and Q. 
-The linear polarization present in Betelgeuse is attributed to brightness inhomogeneities, which are due to the convective cells 
-present at the surface \citep{lopez_ariste_convective_2018}. Hence, if we are able to find periodicities in the LS periogogram of the LSD profiles of Betelgeuse,
-the periodicities would be associated to surface convection.
+We apply this technique to the  LSD profiles of Stokes I, U and Q of Betelgeuse observed at the TBL during 
+that span of 13 years.
 
-To recover variability of Betelgeuse, 
-we computed the LS periodogram at each wavelength of the LSD profile, between -25 km/s and 80 km/s. We choose those velocity to span the entire 
-signal present in the LSD profile. In the work of \cite{lopez_ariste_convective_2018}, the most blueshifted signal was present at -20 km/s, 
-corresponding to the maximal velocity of the rising plasma. The most redshifted signal was 40 km/s. This velocity corresponds 
-to the velocity of the star in our reference frame. This last velocity is very important in our study. Signal above 40 km/s is attributed to either huge
-convective cells falling back to Betelgeuse, or to rising plumes of plasma located in the hidden face of Betelgeuse,
-that rise so high that they appear above the limb.
+We first computed the LS periodogram at each wavelength of the LSD profile, between -25 km/s and 80 km/s to span the entire 
+signal present in the LSD profile. \cite{lopez_ariste_convective_2018} found the most blueshifted signal towards -20 km/s, 
+and interpreted it as  the maximum velocity of the rising plasma, seen from the Earth. The most redshifted signal was found at 40 km/s. 
+This velocity is interpreted
+as the rest velocity of the star in our reference frame. Seldom, signals at velocities larger than 40 km/s can be found 
+and can be interpreted as either dark and cold plasma falling back to Betelgeuse, 
+or to rising plumes of plasma located in the hidden face of Betelgeuse,
+and rising so high that they appear above the limb.
 
-\subsection{Before and after the dimming}
+% The LS periodogram  computed for each frequency 
+% in the LSD profile must show periodicities around the LSP, but also on the 400 days and 200 days period. If the LS periodogram finds a periodicity at a
+% given wavelength, it would appear as a strong LS power. However, we expect a change in the periodicity after the dimming. 
+% This would lead to a decrease in the Lomb-Scargle power. By comparing LS periodogram before and after the dimming, we expect to see a change in the LS power,
+% especially at the 400 days period.
 
-At the beginning of 2020, Betelgeuse reached a historical minimum in its luminosity, called the great dimming \citep{guinan_fall_2020}. 
-From interferometric data, we know that this event was caused by a cloud of gas in front of the line of sight \citep{montarges_dusty_2021}. 
-Interferometric images show a huge drop in luminosity in the southern hemisphere of Betelgeuse, leading to this dimming. Interestingly, 
-it has been shown by \cite{jadlovsky_analysis_2023} and \cite{dupree_great_2022} that the periodicity of Betelgeuse has changed after the dimming. Using the light curves from AAVSO, 
-the authors showed that before the dimming, the dominant period of Betelgeuse was around 400 days. However, after the dimming, this period has changed and 
-is now shorter, around 200 days. It means that, after the dimming, there was a change in the behaviour of the atmosphere of Betelgeuse: the period is now
-shorter than before the dimming. Our analysis is based on this change of periodicity.
-
-By computing the LS periodogram for each frequency 
-in the LSD profile, we expect to find periodicities around the LSP, but also on the 400 days and 200 days period. If the LS periodogram finds a periodicity at a
-given wavelength, it would appear as a strong LS power. However, we expect a change in the periodicity after the dimming. 
-This would lead to a decrease in the Lomb-Scargle power. By comparing LS periodogram before and after the dimming, we expect to see a change in the LS power,
-especially at the 400 days period.
-
-Figure \ref{LS intensity},\ref{LS Q},\ref{LS U} and \ref{LS linear polarization} show the Lomb-Scargle periodogram of the LSD profile of respectively 
-the intensity, stokes Q, stokes U and the linear polarization. The abscissa is the frequency (in $\mathrm{days^{-1}}$), while the ordinate is the heliocentric radial velocity in km/s. 
+Figures \ref{LS intensity},\ref{LS Q},\ref{LS U} and \ref{LS linear polarization} show the Lomb-Scargle periodogram in normalised power of the LSD profile of respectively 
+the intensity, stokes Q, stokes U and the linear polarization. 
+%The abscissa is the frequency (in $\mathrm{days^{-1}}$), while the ordinate is the heliocentric radial velocity in km/s. 
 For each figure, the upper panel is the LS periodogram with data taken before the great dimming of Betelgeuse, that is from 2013 to the beginning of 2019. 
-The lower panel is the LS periodogram for the data before and after the dimming: from 2013 to 2023. The white dashed lines represent the different 
-periods found in the literature: the LSP at 2000 days, the fundamental pressure mode at 400 days and its first overtone at 200 days. The Lomb-Scargle 
-power is normalised. The colorbar indicates the quality of the fit: red areas correspond to good fits from the LS periodogram.
+The lower panel is the LS periodogram for the whole period before and after the dimming: from 2013 to 2023. The white dashed lines mark the different 
+periods found in the literature: the LSP at 2000 days, the fundamental pressure mode at 400 days and its first overtone at 200 days.
 
@@
 
-\section{Recovering the variability of Betelgeuse in the LSD profile}
+%\section{Recovering the variability of Betelgeuse in the LSD profile}
 
-\subsection{Intensity}
+%\subsection{Intensity}
 Figure \ref{LS intensity} shows the LS periodogram of the LSD of the intensity profile. Interestingly, we recover the LSP at 2000 days, 
-located at a HRV of $\sim$ 20 km/s. This periodicity seems to change after the dimming, where it is located at a HRV between 0 and 10 km/s. 
-There is also a weaker signal at the fundamental mode, around 400 days. This signal completely disappears after the dimming. It is hard to attribute 
-this LS power to a periodicity of Betelgeuse, however it might be a hint that this periodicity has changed after the dimming. Furthermore, our results 
+but located at a HRV of $\sim$ 20 km/s rather than over the whole profile. This periodicity is still present after the  dimming, but  it is located at a HRV between 0 and 10 km/s. 
+There is also a weaker signal at the fundamental mode, around 400 days before the dimming. But this signal completely disappears after the dimming. From these observations
+we find  hard to attribute 
+the observed LS power to a global period of Betelgeuse.
+%, however it might be a hint that this periodicity has changed after the dimming. 
+Our results 
 in the intensity profile are consistent with the one from \cite{mathias_evolution_2018} who used 2D fourier analysis and found that the dominant period was the LSP.
@@
 Those periodograms rise a lot of questions: why is the 200 days period only present in stokes Q? Why not in stokes U or intensity? 
-We will discuss about those questions in the next section. However, it also appears that stokes Q is not sensitive to the LSP, or to be more precise, 
+We will discuss about those questions in the next section. It also appears that stokes Q is not sensitive to the LSP, or to be more precise, 
 it is much more sensitive to the 200 days period. From this, we can conclude that stokes Q evolves on a time scale much more closer to the first overtone 
-mode than the LSP or even the fundamental mode. Because stokes Q is sensitive only to shorter periods, the variation of stokes Q come from another mechanism 
-than the one from intensity. Linear polarization in Betelgeuse is due to convective cells, we attribute the signal present in the periodogram 
-as the lifetime of convective cells. This is what we observe in the LSD profile of Betelgeuse: from a week to another it does not change that much. 
-However, after a few month, the LSD profile of Betelgeuse has completely changed, due to the evolution of convective cells at the surface. 
-This would explain the strong signal present at 200 days in stokes Q.
+mode than the LSP or even the fundamental mode. Since Stokes Q is sensitive  to different periods, we must conclude that it is due to other mechanisms
+than the one behind the periodicitys observed in intensity. Linear polarization in Betelgeuse is interpreted to be a proxy of  convective cells,
+so we can speculate that the signal present in the periodogram 
+may be related to  the lifetime of convective cells. Such hypothesis is comforted by what we see in the 
+LSD profiles of  Betelgeuse  whose shape is overhauled  in typical time scales of  a few months,  
+Such time scales observed in the Stokes profiles and attributed to the convective time scales 
+appear to explain the strong signal present at 200 days in the LS periodogram of Stokes Q.
 
@@
 
-Figure \ref{LS linear polarization} shows the LS periodogram of the linear polarization of Betelgeuse: $\sqrt{\mathrm{Q^2+U^2}}$. 
+Figure \ref{LS linear polarization} shows the LS periodogram of the total linear polarization of Betelgeuse: $\sqrt{\mathrm{Q^2+U^2}}$. 
 Interestingly, the LS periodogram shows a strong power at 400 days and 2000 days. This signal is much more weaker after the dimming.
-While the LSP is already present in the LS of intensity, here we recover the 400 days periodicity, often invoke in the literature. 
-Nonetheless, the period is present at a HRV of around 40 km/s. We recall that this velocity is the velocity of the star. 
-Every signal above this velocity is either due to the plasma sinking in Betelgeuse, or plumes of plasma rising above the limb of the star, 
-as seen in other red supergiant like $\mu$ Cep \citep{lopez_ariste_height_2023}. Up to this point, we want to pinpoint the two elements that are a clue 
-to understand the variability of Betelgeuse: the 400 days period is present in the linear polarization signal and also at a HRV above 40 km/s.
-In any case, from the linear polarization and the intensity profile, we recover the different periods of Betelgeuse. Since linear polarization is 
-associated to convection and because we recover the periods of Betelgeuse, the main mechanism behind the variability of Betelgeuse points toward surface convection.
+While the LSP is already present in the LS of intensity, here we recover the 400 days periodicity, often invoked in the literature. 
+Nonetheless, the power of this period peaks at a HRV of around 40 km/s. We recall that this velocity is the rest velocity of the star. 
+Every signal beyond this velocity is either due to  plasma sinking in Betelgeuse, or to plumes of plasma rising beyond and above the limb of the star. 
+Such plumes from the back hemisphere haven been often seen in other red supergiants like $\mu$ Cep \citep{lopez_ariste_height_2023}. 
+This suggests that both signals at these period of 400 days found in intensity and total linear polarisation 
+could well be due to the periodic appearance of plumes in the back himisphere rising high enough to appear above
+the visible limb and generating a supplementary source of brightness. Once again we find a plausible explanation 
+of these periods unrelated to pulsations, but rather to the convective dynamics of the star.
+
+% Up to this point, we want to pinpoint the two elements that are a clue 
+% to understand the variability of Betelgeuse: the 400 days period is present in the linear polarization signal and also at a HRV above 40 km/s.
+% In any case, from the linear polarization and the intensity profile, we recover the different periods of Betelgeuse. Since linear polarization is 
+% associated to convection and because we recover the periods of Betelgeuse, the main mechanism behind the variability of Betelgeuse points toward surface convection.
 
@@
 to inteferometric images of \cite{montarges_close_2016}. The images are produced by finding the brightness distribution that better fits the observed linear 
-polarization LSD profile using Marquardt-Levemberg minimisation.
+polarization LSD profile using a Marquardt-Levemberg minimisation.
 
-Betelgeuse is observed every two weeks by the TBL, hence,
-we were able to follow its surface activity through polarimetry imaging. This allowed us to estimate the size and the life time of the convective cells
-at the surface. From our images, we computed a quantity called the photo-center, which is sensitive to the size and the number of convective cells. 
-An homogeneous star will have a photo-center displacement close to the barycenter of the star, whereas a star with one or two huge convective cells will have a
+Betelgeuse is observed in average every other weeks by the TBL, and hence
+we are able to follow its surface activity through polarimetry imaging. This has  allowed in the past the estimatation of  the size and the
+ life time of the convective cells
+at the surface. From these images, we can compute a photo-center, which is sensitive to the size and the number of convective cells. 
+An homogeneous star will have a photo-center coinciding with barycenter of the star, whereas a star with one or two huge convective cells will have a
 photo-center displacement more important, up to a few percent of the stellar radius in the case of RSG \citep{chiavassa_probing_2022}.
 
-Since the photo-center is linked to surface convection, 
-we computed the LS periodogram of the displacement of the photo-center. If we recover the different periods from our images, we would
-be able to explain them by involving only convection, since our imaging model does not involve pulsations. However, before going further, we have to be carefull with the interpretation of our images. 
-Linear polarization suffers from a $180 ^\circ$ degree ambiguity already mentioned in \cite{auriere_discovery_2016}. For that reason, our images can be 
-rotated by $180^\circ$ degree, and the brightness distribution will still fit the observed LSD profile. Thus, to obtain a continuity between our images, 
-for a given day, we use the brightness distribution of the previous day to start fitting the LSD profile.
-This is mandatory to obtain a continuity between the images.
-For the first image, we start with a
-random brightness distribution as the beginning of the fit. Even though we are able to produce images which are consistent with each other, it is important 
-to keep in mind that the photo-center displacement will be affected by the choice of the first image. To overcome this issue, we decided to compute
-the photo-center displacement from 100 different initial parameters. So that we recover properties that are independent of the choice of the first image.
+Since the photo-center is linked to surface convection, it is worth checking for periods in its 
+dynamics over the 13 years of observations of Betelgeuse at the TBL. Such periods will not 
+be completely unrelated to the ones presented in the previous section, but will capture 
+other aspects of the phenomena at work.
+
+We computed the LS periodogram of the displacement of the photo-center. But, before going any further, we have to warn
+ about the interpretation of such images. 
+Linear polarization suffers from a $180 ^\circ$ degree ambiguity already mentioned in \cite{auriere_discovery_2016}. For that reason, our images 
+can be 
+rotated by $180^\circ$ degree, and the brightness distribution will still fit the observed LSD polarization profiles. 
+If a particular observation were to be treated independently of previous observations, the solution found by 
+the algorithm could be any of the ambigous solutions possible. To obtain a continuity 
+between the series images, 
+for a given day we use the brightness distribution of the previous day as initial point 
+of the  fitting iteration; and for the first image, we start with a 
+random brightness distribution as the beginning of the fit. 
+Even though we are able to produce a series of images  consistent with each other, it is important 
+to keep in mind that the photo-center displacement computed from the series 
+will be affected by the choice of the first image. 
+To overcome this issue, we decided to compute
+the photo-center displacement from 100 different initial parameters and their subsequent series 
+of images in an effort to recover from the ensemble
+properties that are independent of the choice of the first image.
 
@@
 
-Figure \ref{LS photocenter} shows the LS periodogram of each photo-center displacement (blue lines) and the average LS periodogram of the photo-center displacements (black line). 
-The upper panel is the LS from the images before the great dimming of Betelgeuse, whereas the lower panel is the LS for the whole set of data, 
-including after the dimming. The black dashed lines are the 2000, 400 and 200 days period respectively. First, we find the the LS power is high around the LSP.
-We also notice a decrease in the Lomb-Scargle power before and after the dimming. It appears that there are two smalls bumps around the 400 days period. 
-Those bumps almost disappear after the dimming. Even though it is hard to attribute the 400 days variability to those bumps, from polarimetry imaging,
-we still find that there is a difference before and after the dimming. However, in our case, our images only involve surface convection.
-The LS of the photo-center displacement point toward a variability explained by convection only. The photo-center only is not enough to attribute the variability
-of Betelgeuse to surface convection, but what we found from polarimetry imaging is consistent with what we obtained from linear polarization.
-Both the photo-center displacement and the LS of the linear polarization point toward a variability explained by surface convection.
-
-\section{Surface convection to explain variability}
-
-\subsection{The 400 days period explained by rising plumes of plasma}
-
-In the previous section, we have seen that we recover the different period of Betelgeuse from the intensity and linear polarization signal. 
-While the nature of the LSP is still in debate, the nature of the lower periods is attributed to pulsations of the pressure modes 
-(e.g \cite{kiss_variability_2006}). The mechanism behind the pulsation of Betelgeuse is attributed to the $\kappa$-mechanism. In this section, we attempt to 
-find an explanation to the 400 days periodicity of Betelgeuse. As mentioned before, the LS periodogram of linear polarization shows a periodicity 
-around 400 days. Furthermore, the region where the LS power is high is around 40km/s. This wavelength corresponds to most redshifted signal present in Betelgeuse.
-
-The polarization signal is supposed to remain between -20 and 40km/s. Signal above 40km/s is associated to plasma rising behind the limb of Betelgeuse. 
-This scenario has been observed in the RSG $\mu$ Cep and was introduced by \cite{lopez_ariste_height_2023} to explain the excess of linear polarization beyond 
-the velocity of the star. In $\mu$ Cep, the linear polarization signal is very strong and often present, whereas in Betelgeuse, the signal above the 
-velocity of the star is less strong. However, because it is present from time to time, the LS periodogram finds this signal above the limb and interprets it as a
-periodic signal at this wavelength of 40 km/s. Hence, the strong power associated to this wavelength might be due to those convective plumes that are 
-rising above the limb from time to time. From this scenario, we can explained the 400 days period in the light curve. From time to time,
-Betelgeuse is ejecting plasma in the interstellar medium due to winds. The plasma can be ejected from the hidden face of Betelgeuse. 
-It rises so high that we see the rising plume from the Earth, because the plume rises above the limb of the star, contributing to the linear polarization signal. Hence, from Earth, 
-the signal associated to this plume appears redshifted compared to the velocity of the star. This plume increases the number of photons that we receive,
-leading to an increase of the brightness. After a few weeks or months, the plume falls back to Betelgeuse, disappearing from our point of view, 
-leading to a decrease of the brightness.
-This is exactly what we see in the LSD profile of the linear polarization of Betelgeuse: 
-from time to time, there is a net signal above 40km/s, which is present for a few weeks, and then disappears. Since this signal is not always present, 
-it is found by the LS periodogram as a periodic event. However, it is not periodic at all: because the plumes rise at random moments, sometimes they 
-appear every years, sometime they do not. This would explain why we only see it before the dimming. After the dimming, this kind of event never happened 
-for the moment. Thus, the LS periodogram, at that precise frequency, finds that there is a periodic signal before the dimming, but not after, leading to
-a lower LS power at this frequency.
-
-Concerning the 200 days period, it is caused by the timescale on which the stokes parameters evolve. For instance, the 
-LSD profile of stokes Q or U will slightly change if we observe Betelgeuse every two weeks. However, on a longer timescale, for example a few month, 
-the LSD profile of stokes Q and U will change a lot. They are directly linked to the dynamics of the surface, hence,  their variation scale the lifetime 
-of convective cells. Some convective cells can live up to a few month, leading to this peak in the LS periodogram of stokes Q. As showed by
-\cite{lopez_ariste_convective_2018}, some convective cells can live up to years, explaining also the LSP.
-One can notice that the peak is not present in stokes U, which might be due to stochastic event.
-However, if we try to explain the periodicity of Betelgeuse using pulsations, we would see this signal in either stokes U and the linear polarization.
-
-To summarise what we have done so far: we were able to explain the variability of Betelgeuse using surface convection only. The convective cells play 
-the most important role in the variability of Betelgeuse. Our work does not need pulsation to explain its variability. This does not mean that there are 
-no pulsation in Betelgeuse. What we claim is that the most important factor in the variability of Betelgeuse is the timescale on which the convective 
-cells evolve. Pulsation could exist on Betelgeuse, but they are not the main contributor to the variability.
+Figure \ref{LS photocenter} shows the LS periodogram of the photo-center displacement  of each one of those 100 series (blue lines) and the average LS periodogram (black line). 
+As before, the upper panel shows the LS from images before the great dimming of Betelgeuse only, whereas the lower panel is the LS for the whole set of 
+data,  including images after the dimming. The black dashed lines mark the 2000, 400 and 
+200 days periods respectively. First, we find the the LS power is high around the LSP, though
+uncomfortably close to the auto-correlation peak.
+We also notice a decrease in the Lomb-Scargle power from  before to after the dimming, in 
+concurrence to what was seen in the LS periodograms of the LSD profiles. Before it, there are two smalls bumps around the 400 days period. 
+But those bumps almost completely disappear after the dimming. Even though it is 
+hard to identify the 400 days period from those bumps alone,
+we still find that there is a difference before and after the dimming. 
+And this is concluded  from images that exclusively involve surface convection.
+The LS of the photo-center displacement points towards a variability explained by convection only. 
+And this is consistent with what we obtained from linear polarization.
+%Both the photo-center displacement and the LS of the linear polarization point toward a variability explained by surface convection.
 
@@
 
-In the literature, it is reported that before the great dimming, the main period of Betelgeuse was 400 days \citep{kiss_variability_2006}. 
-This pulsation is associated to the fundamental model of pressure modes. After the dimming, this period is shorter and is now around 200 days, 
-associated to the first overtone of the fundamental mode. However, there are no explanations concerning this change of variability due to the dimming.
+After examining the periods found by the Lomb-Scargle technique on the polarization data, and its derivatives, obtained by the TBL over the last 
+13 years, it is worth putting them in the context of the periods found traditionally on light curves over the same period of time. Two aspects 
+are of our interest in this comparison: the behaviour of the light curve before and after the great dimming, and the amplitudes of the peaks 
+at the main cited periods.
 
-Using the data from AAVSO, we studied the light curve of Betelgeuse. Figure \ref{light curve Betelgeuse} shows the light curve of Betelgeuse in the visible, 
-from 2014 up to now.
+% In the literature, it is reported that before the great dimming the main period of Betelgeuse was of 400 days \citep{kiss_variability_2006}. 
+% This pulsation is often associated to the fundamental pressure mode. After the dimming, however this period appears to have disappeared 
+% and only the period at 200 days,  associated to the first overtone of the fundamental mode, is visible since. 
+% No explanations based upon pulsations have been ofered concerning this change of variability after or perhaps because of the dimming. But this 
+% is a behaviour that
+
+We have recovered the light curve of Betelgeuse in the visible from the AAVSO database for the last 13 years (Figure \ref{light curve Betelgeuse}).
 Before the great dimming, the magnitude of Betelgeuse evolves on a timescale of a year, whereas after the dimming, the variability of Betelgeuse has changed. 
-It is clear from fig.
-\ref{light curve Betelgeuse} that the
-variability of Betelgeuse is evolving on a timescale shorter than one year. However, from middle 2023 up to now, the magnitude of Betelgeuse remains almost 
-constant, which is not consistent with a pulsating scenario. The constant magnitude of Betelgeuse points toward another scenario than pulsation. Using surface 
-convection, we can fully explain the variability of Betelgeuse. The oscillations in the light curve are due to two mechanisms: surface convection and rising 
-plumes of plasma beyond the limb of the star. This last mechanism being random, it happens from time to time. The timescale on which convective cells evolve
-is between one and two years, which is also consistent with the 400 days period found in the literature.  After the dimming, the change of periodicity in the
-light curve is explained by surface convection, the random variation of brightness distribution could lead to the 200 days periodicity found in the literature. 
-The constant brightness of Betelgeuse for the last 6 months is explainable by the same random brightness distribution: the surface of Betelgeuse is calm, 
-leading to this constant brightness. In the future, we expect to see the light curve of Betelgeuse to change, maybe with another main period, but which 
-is in fact caused by surface convection. Thus, the main mechanism in the variability of Betelgeuse is surface convection, which is sufficient to explain the 
-variation in the magnitude of Betelgeuse.
+It is clear from fig. \ref{light curve Betelgeuse} that the 
+variability of Betelgeuse is evolving on a timescale shorter than one year. This qualitative change has been pointed out before.
+In the literature, it is reported that before the great dimming the main period of Betelgeuse was of 400 days \citep{kiss_variility_2006}. 
+This period is often associated to the fundamental pressure mode. After the dimming, however this period appears to have disappeared 
+and only the period at 200 days,  associated to the first overtone of the fundamental mode, is visible since \citep{XXX}.
+No explanations based upon pulsations have been ofered concerning this change of variability after or perhaps because of the dimming. But this 
+is a behaviour that fully concurs with what we find in the polarization data, and which cannot be related to any stellar pulsation but rather 
+convective dynamics. Beyond the actual values of the periods found in a Lomb-Scargle periodogram, we also see a correspondence between polarization 
+and the light curve in these changes in the observed periods before and after the great dimming. Such concurrence points to an explanation 
+in terms of convective dynamics also for the variability of the light curve.
+
+It may be argued that the periodograms presented in the previous sections and, in particular, those  of the photo-center displacements, 
+present low amplitude peaks at the referred main periods. One may wonder whether we are over-interpreting mere fluctuations of the periodograms, 
+while the peaks seen in the literature from analysis of the light curve are unmistakable. It is therefore worth to take advantage of the 
+availability of the AAVSO data to perform a Lomb-Scargle periodogram of the light curve in the same period covered by our polarization data and,
+furthermore, taking into consideration data around the dates for which TBL data is available. Periodograms  with such constraints are 
+shown in Fig.\ref{LSAAVSO}. Even when the full dataset for the period is taken into account, the periods present small amplitudes, and are 
+fully and favourably comparable with the periodograms presented above. This is even more dramatic when the periodogram is made from data at the exclusive 
+dates at which TBL data is available. It stands out that the low amplitude peaks are rather due to both the sparse cadence and short span of 
+time of the TBL data series rather than to a physical absence. Actually, the present tests puts more emphasis on the importance of the 
+strong peaks seen at selected wavelengths in the LSD profiles, strong when compared to the peaks at similar periods arising from the light curve.
+
+% constant, which is not consistent with a pulsating scenario. The constant magnitude of Betelgeuse points toward another scenario than pulsation. Using surface 
+% convection, we can fully explain the variability of Betelgeuse. The oscillations in the light curve are due to two mechanisms: surface convection and rising 
+% plumes of plasma beyond the limb of the star. This last mechanism being random, it happens from time to time. The timescale on which convective cells evolve
+% is between one and two years, which is also consistent with the 400 days period found in the literature.  After the dimming, the change of periodicity in the
+% light curve is explained by surface convection, the random variation of brightness distribution could lead to the 200 days periodicity found in the literature. 
+% The constant brightness of Betelgeuse for the last 6 months is explainable by the same random brightness distribution: the surface of Betelgeuse is calm, 
+% leading to this constant brightness. In the future, we expect to see the light curve of Betelgeuse to change, maybe with another main period, but which 
+% is in fact caused by surface convection. Thus, the main mechanism in the variability of Betelgeuse is surface convection, which is sufficient to explain the 
+% variation in the magnitude of Betelgeuse.
 
@@
 
+\section{Surface convection suffices to explain the observed variability}
+
+Before summing up the conclusions of this work, we can afford proposing alternative scenarios for the observed periodicities in the light curve, 
+the LSD profiles and the displacement of the photo-center of the inferred photospheric image reconstructions. The presence of the same 
+periods and, after taking into consideration the time span and sparsity of the TBL data series, the larger amplitude of the peaks in
+polarization data  than in the light curve point to a common origin, one that is rather based upon convective dynamics and not to radial or 
+other global pulsations.
+
+% We recall that while the nature of the LSP is still in debate, the nature of the lower periods is quite commonly 
+% attributed to  pressure modes 
+% (e.g \cite{kiss_variability_2006}). The mechanism behind the pulsation of Betelgeuse is attributed to the $\kappa$-mechanism. In this 
+% section, we attempt to 
+% find an explanation to the 400 days periodicity of Betelgeuse. 
+% As mentioned before, the LS periodogram of linear polarization shows a periodicity 
+% around 400 days. Furthermore, the region where the LS power is high is around 40km/s. 
+% This wavelength corresponds to most redshifted signal present in Betelgeuse.
+
+The 400-day period appears prominently in both the intensity and the total polarization profiles. And in both cases, it is mostly present 
+at wavelengths in the red wing, often beyond the 40 km/s value.
+We recall that in the present scenario for the formation of the polarization signal, this signal is supposed to remain between 
+-20 and 40km/s. Signals above 40km/s are uncommon and can be associated to
+ plasma rising behind the limb of Betelgeuse. 
+Such scenario has been observed in the RSG $\mu$ Cep and was introduced by \cite{lopez_ariste_height_2023} to explain the excess of 
+linear polarization beyond  the rest velocity of the star, which for Betelgeuse is fixed at +40km/s. 
+In $\mu$ Cep, such  linear polarization signal beyond the limit of the rest velocity of the star is very strong and particularly 
+common, two features that prompted a dedicated explanation. In Betelgeuse, such signals, when present, are weaker. However, 
+they are visible from time to time, with regularities that are captured by  the LS periodogram at 400 days.
+Without further justification, we may propose that convective plumes in the hidden hemisphere of the star and 
+rising above the limb from time to time are responsible for the  400 days period in the light curve. When such plumes rise high enough 
+to be geometrically visible above the limb they show signals in intensity and polarization at redshifted wavelengths. They also
+ increase the emitting surface of the star and  the number of photons that we receive,
+leading to an increase of the brightness visible in the light curve. After a few weeks or months, the plume falls back to Betelgeuse, 
+disappearing from our point of view,  leading to a decrease of the brightness.
+
+The appearance of such high-rising plumes is not necessarily periodic. And the star can easily shift from active periods when such 
+plumes are a common feature to calmer periods when no plume is visible for months. $\mu$ Cep may be in one such active period presently,
+while Betelgeuse may have experienced recently one great such event in the direction of the Earth that, after dust formed, became the great 
+dimming, but since then no other high-rising plume has appeared, explaining the absence of the 400-day period since the great dimming. 
+From the point of view of convective dynamics we can offer an explanation of the 400-day period which not only explains the observation of 
+this period both in the light curve and in the polarization profiles, but also its presence at redshifted wavelengths and its present (and probably 
+temporary) disappearance since the great dimming.
+
+% This is exactly what we see in the LSD profile of the linear polarization of Betelgeuse: 
+% from time to time, there is a net signal above 40km/s, which is present for a few weeks, and then disappears. Since this signal is not always present, 
+% it is found by the LS periodogram as a periodic event. However, it is not periodic at all: because the plumes rise at random moments, sometimes they 
+% appear every years, sometime they do not. This would explain why we only see it before the dimming. After the dimming, this kind of event never happened 
+% for the moment. Thus, the LS periodogram, at that precise frequency, finds that there is a periodic signal before the dimming, but not after, leading to
+% a lower LS power at this frequency.
+
+Concerning the 200 days period, we notice that it coincides with  the typical timescales on which the Stokes parameters evolve. 
+The 
+LSD profiles of stokes Q or U will slightly change when we observe Betelgeuse every two weeks. But they are completely overhauled 
+on time scales of several months. The Lomb-Scargle perioograms appear to capture this spectral dynamic around periods of 200 days.
+ Such changes in the profiles are directly linked to the dynamics of the surface,  to  the evolution and movements 
+of convective cells over the visible hemisphere. So the 200-day period appears to be related to the evolution of the convective patterns 
+on the surface of Betelgeuse. 
+ As showed by
+\cite{lopez_ariste_convective_2018}, some of those convective cells can live up for years. We may speculate
+that the LSP may be related to those long living structures. However the present span of data from the TBL is insufficient to do more than
+just speculate. It should be noticed that  
+ the peak is not present in stokes U, and this may point to a more stochastic event than just long-living structures.
+
+% To summarise what we have done so far: we were able to explain the variability of Betelgeuse using surface convection only. The convective cells play 
+% the most important role in the variability of Betelgeuse. Our work does not need pulsation to explain its variability. This does not mean that there are 
+% no pulsation in Betelgeuse. What we claim is that the most important factor in the variability of Betelgeuse is the timescale on which the convective 
+% cells evolve. Pulsation could exist on Betelgeuse, but they are not the main contributor to the variability.
+
 \section{Conclusion}
+
+The main observation  presented in this work is that the very same periods identified in the light curve of Betelgeuse are also seen in the intensity and polarization
+spectra measured with the TBL. Periods in the light curve have often been traditionally associated to pulsation due to pressure modes. But 
+such pulsation would in no case affect the polarisation signals nor would explain the fact that the periods can be observed with stronger 
+amplitude at selected wavelengths but no others.  Pulsation appears to offer no coherent explanation to these observations, are hardly to the 
+observed changes in the variability since the great dimming.  The concurrence of observational facts, actual period values, evolution before 
+and after the dimming, the larger amplitude of the period peaks in polarizarion data, points towards a common origin for those variabilities. 
+And such origin cannot be pulsations.
+
+The interpretaion of the linear polarization profiles is made in terms of convective structures in the atmosphere of Betelgeuse. We suggest 
+that it is in these convective structures and their temporal evolution that we will find the true reason for the observed periodicities both 
+in the light curve and in the spectropolarimetric observations. This is the main conclusion of this work: the observed periods are related 
+to convective dynamics and not to radial pulsations of the star.
+
+Although of secondary importance, we may suggest with more or less ease what phenomena are at the origin of the main periods. The 400-day period 
+appears to be strongly related to the appearence of plumes in the back hemisphere of the star rising high enough to be seen above the limb.
+Such plumes have been described in another RSG, $\mu$ Cep, and while less common or prominent, they are also present in Betelgeuse. The irregularity 
+of their presence justifies the 400-day period, which is captured by the Lomb-Scargle periodogram in the polarization signals attributed to 
+this phenomenon. The redshifted wavelengths at which this periodicity is found are justified by the plume moving away of the star and of the Earth while 
+raising in the back hemisphere, while the brightness increase is justified by the extra surface of emitting plasma visible. The 200-day period is less 
+prominent but appears to be correlated to the typical time scales of the evolution of the convective structures themselves on the visible hemisphere. 
+The LSP origin is harder to pinpoint given the insufficient time length of our time series. One can speculate if it may be related to the 
+longest living photospheric structures, but this explanation does not adequately explain all the features associated to this period. Longer time 
+series of spectropolarimetric observations of Betelgeuse will be required to shed light on this.
 
@@
     funded project PEPPER (ANR-20-CE31-0002)
+    An acknowledgement to the AAVSO may be required??????????????????
     \end{acknowledgements}

### 2024-03-07 main.tex
modified:   art75.bib 	modified:   main.tex

@@
 periods around 400 and 200 days. Such periods are often linked to radial pulsation modes,   and have been used to model 
-the characteristics of the resonant cavity, hence the size of the star and eventually its evolutionary stage \cite{XXX}.
+the characteristics of the resonant cavity, hence the size of the star and eventually its evolutionary stage \citep{saio_evolutionary_2023}.
 
 At the end of 2019, Betelgeuse reached a historical minimum in its luminosity, called the great dimming \citep{guinan_fall_2020}. 
-From interferometric data, we know that this event was caused by a cloud of dust close to the line of sight \citep{montarges_dusty_2021}. 
+From interferometric data, we speculate that this event was caused by a cloud of dust close to the line of sight \citep{montarges_dusty_2021}. 
 Interferometric images show a huge drop in luminosity in the southern hemisphere of Betelgeuse, leading to this dimming.
@@
 
-Betelgeuse is known to present large convective cells with life times in the order of one to two years and measurable changes 
+Betelgeuse is known to present large convective cells with lifetimes in the order of one to two years and measurable changes 
 in the span of one week. Bright convection cells near disk center will increase the integrated brightness of the star when 
@@
 it has been shown by \cite{jadlovsky_analysis_2023} and \cite{dupree_great_2022} that the periodicity of Betelgeuse has changed after the dimming. Using the light curves from AAVSO, 
-the authors showed that before the dimming, the dominant period of Betelgeuse was around 400 days. However, after the dimming, this period has changed and 
-is now shorter, around 200 days. It means that, after the dimming, there was a change in the behaviour of the atmosphere of Betelgeuse: the period is now
-shorter than before the dimming. Our analysis is based on this change of periodicity. 
+the authors showed that before the dimming, the dominant period of Betelgeuse was the 400 days period. However, after the dimming, this period has changed and 
+is now shorter, around 200 days.
+Therefore, after the dimming, there was a change in the behaviour of the atmosphere of Betelgeuse: the time scales on which Betelgeuse evolves are now shorter.
+Our analysis is based on this change of periodicity.
 Faced with the dramatic change in stellar parameters and age that such change in a pulsation period would 
 require, our interest in an alternative explanation of the variability
-in terms  of convective patterns arose. In order to address this question we shall seek for such typical periods of 1000, 400 and 200 days in  
-observational proxies related to the convective activite but unrelated to pulsations and which cannot be directly linked to brightness?
+in terms  of convective patterns arose. In order to address this question we shall seek for such typical periods of 2000, 400 and 200 days in  
+observational proxies related to the convective activity but unrelated to pulsations and which cannot be directly linked to brightness.
 
@@
 even producing 3-dimensional images of the atmosphere of Betelgeuse \citep{lopez_ariste_three-dimensional_2022} when taking advantage of the different height of formation of 
-different lines in the spectrum of Betelgeuse. The images produced compare very well with cotemporaneous images made with interferometric 
-techniques and show clear convective patterns, akin to solar granulation.
+different lines in the spectrum of Betelgeuse. The images produced compare very well with contemporaneous images made with interferometric 
+techniques \citep{montarges_close_2016} and show clear convective patterns, akin to solar granulation.
 
@@
 
-Betelgeuse has been observed for the last 13 years with Narval and Neo-Narval at the Telescope Bernard Lyot. These two instruments measure the polarization 
+Betelgeuse has been observed for the last 13 years with Narval and Neo-Narval at the Telescope Bernard Lyot (TBL). These two instruments measure the polarization 
 over the visible spectra of Betelgeuese (390-1000 nm) with high spectral resolution (R=65000) and high polarimetric sensitivity. Despite 
 this high sensitivity, the signal-to-noise ratios are not sufficient to measure the weak polarization signals in individual atomic lines of the 
-spectrum. These amplitudes are today known to be of the order of $10^{-4}$ times the continuum intensity, and only excepcionnaly do they 
+spectrum. These amplitudes are today known to be of the order of $10^{-4}$ times the continuum intensity, and only exceptionally do they 
 reach amplitudes of $10^{-3}$ the continuum. In these exceptional observations, enough photons can be accumulated per spectral bin to 
 see the linear polarization signal above noise \citep{auriere_discovery_2016}. Most commonly, the amplitudes of linear polarization are below noise levels. In those 
-occassions we have to add up the signals of thousands of lines to reduce the noise and increase the signal-to-noise ratios. This line addition 
+occasions we have to add up the signals of thousands of lines to reduce the noise and increase the signal-to-noise ratios. This line addition 
 is done through a technique called Least-Squares Deconvolution (LSD) \citep{donati_spectropolarimetric_1997} and assumes that the 
@@
 \cite{lopez_ariste_three-dimensional_2022}, which also give details on observing times and conditions as well as more detailed description 
-on the data reduction \footnote[1]{Beyond a 2-year proprietary embargo, and up to technical issues, all these data is available through 
+on the data reduction. \footnote[1]{Beyond a 2-year proprietary embargo, and up to technical issues, all these data is available through 
 PolarBase (http://polarbase.irap.omp.eu/).}
@@
 Trying to find periods in an astrophysical context carries the hurdle of unevenly spaced observed data. To overcome this issue, 
-we used the Lomb-Scargle (hereafter LS) periodogram \citep{lomb_least-squares_1976,scargle_studies_1982}. For a given set of data, it fits sinusoid functions 
+we used the Lomb-Scargle periodogram \citep{lomb_least-squares_1976,scargle_studies_1982}. For a given set of data, it fits sinusoidal 
 using least-squares at each frequency between the first and the last observation. The better the fit, the higher the power attributed to the frequency. 
-We apply this technique to the  LSD profiles of Stokes I, U and Q of Betelgeuse observed at the TBL during 
+We applied this technique to the  LSD profiles of Stokes I, U and Q of Betelgeuse observed at the TBL during 
 that span of 13 years.
 
-We first computed the LS periodogram at each wavelength of the LSD profile, between -25 km/s and 80 km/s to span the entire 
+We first computed the Lomb-Scargle periodogram at each wavelength of the LSD profile, between -25 km/s and 80 km/s to span the entire 
 signal present in the LSD profile. \cite{lopez_ariste_convective_2018} found the most blueshifted signal towards -20 km/s, 
@@
 %The abscissa is the frequency (in $\mathrm{days^{-1}}$), while the ordinate is the heliocentric radial velocity in km/s. 
-For each figure, the upper panel is the LS periodogram with data taken before the great dimming of Betelgeuse, that is from 2013 to the beginning of 2019. 
-The lower panel is the LS periodogram for the whole period before and after the dimming: from 2013 to 2023. The white dashed lines mark the different 
-periods found in the literature: the LSP at 2000 days, the fundamental pressure mode at 400 days and its first overtone at 200 days.
+For each figure, the upper panel is the Lomb-Scargle periodogram with data taken before the great dimming of Betelgeuse, that is from 2013 to the beginning of 2019. 
+The lower panel is the Lomb-Scargle periodogram for the whole dataset, before and after the dimming: from 2013 to 2023. The white dashed lines mark the different 
+periods found in the literature: the LSP at 2000 days, the fundamental pressure mode at 400 days and its first overtone mode at 200 days.
 
@@
     \caption{Lomb-Scargle periodogram of the LSD profile of intensity. 
-    The upper pannel is the LS periodogram for data before the great dimming while the lower panel is the periodogram from the all dataset.
-    The ordinate refers to the velocity in the heliocentric radial velocity in km/s. The abscissa is the period in $\mathrm{days^{-1}}$. 
-    The three white dashed lines represent respectively (from left to right) the LSP at 2000 days, the 400 days period and its first overtone at 200 days. 
-    The Lomb-Scargle power refers to the quality of the fit from the LS periodogram. }
+    The upper panel is the Lomb-Scargle periodogram for data before the great dimming while the lower panel is the periodogram from the whole dataset.
+    The ordinate refers to the velocity in the heliocentric radial velocity in km/s.
+    The three white dashed lines represent respectively (from left to right) the LSP at 2000 days, the 400 days period and its first overtone at 200 days.  }
     \label{LS intensity}
@@
     \includegraphics[width=0.5\textwidth]{Lomb-Scargle linear polarization.png}
-    \caption{Same as figure \ref{LS intensity} for the linear polarization: $\sqrt{\mathrm{Q^2}+\mathrm{U^2}}$.}
+    \caption{Same as figure \ref{LS intensity} for linear polarization: $\sqrt{\mathrm{Q^2}+\mathrm{U^2}}$.}
     \label{LS linear polarization}
@@
 %\subsection{Intensity}
-Figure \ref{LS intensity} shows the LS periodogram of the LSD of the intensity profile. Interestingly, we recover the LSP at 2000 days, 
-but located at a HRV of $\sim$ 20 km/s rather than over the whole profile. This periodicity is still present after the  dimming, but  it is located at a HRV between 0 and 10 km/s. 
+Figure \ref{LS intensity} shows the Lomb-Scargle periodogram of the LSD of the intensity profile. Interestingly, we recover the LSP at 2000 days, 
+but located at a HRV of 20 km/s rather than over the whole profile. This periodicity is still present after the  dimming, but  it is located at a HRV between 0 and 10 km/s. 
 There is also a weaker signal at the fundamental mode, around 400 days before the dimming. But this signal completely disappears after the dimming. From these observations
 we find  hard to attribute 
-the observed LS power to a global period of Betelgeuse.
+the observed Lomb-Scargle power to a global period of Betelgeuse.
 %, however it might be a hint that this periodicity has changed after the dimming. 
 Our results 
-in the intensity profile are consistent with the one from \cite{mathias_evolution_2018} who used 2D fourier analysis and found that the dominant period was the LSP.
+in the intensity profile are consistent with the one from \cite{mathias_evolution_2018} who used 2D Fourier analysis and found that the dominant period was the LSP.
 
@@
 
-Figure \ref{LS Q} and \ref{LS U} show the LS periodogram of both stokes Q and U. It is clear that the LS periodogram is completely different 
+Figure \ref{LS Q} and \ref{LS U} show the Lomb-Scargle periodogram of both stokes Q and U. It is clear that the periodogram is completely different 
 from the intensity one. A strong signal is present in fig. \ref{LS Q} around 200 days. Once again, this signal is weaker after the dimming.
@@
 Those periodograms rise a lot of questions: why is the 200 days period only present in stokes Q? Why not in stokes U or intensity? 
-We will discuss about those questions in the next section. It also appears that stokes Q is not sensitive to the LSP, or to be more precise, 
-it is much more sensitive to the 200 days period. From this, we can conclude that stokes Q evolves on a time scale much more closer to the first overtone 
+We will discuss about those questions in the next section. It also appears that stokes Q is
+much more sensitive to the 200 days period than the LSP. From this, we can conclude that stokes
+Q evolves on a time scale much more closer to the first overtone 
 mode than the LSP or even the fundamental mode. Since Stokes Q is sensitive to different periods, we must conclude that it is due to other mechanisms
-than the one behind the periodicitys observed in intensity. Linear polarization in Betelgeuse is interpreted to be a proxy of  convective cells,
+than the one behind the periodicities observed in intensity. Linear polarization in Betelgeuse is interpreted to be a proxy of  convective cells,
 so we can speculate that the signal present in the periodogram 
 may be related to  the lifetime of convective cells. Such hypothesis is comforted by what we see in the 
-LSD profiles of  Betelgeuse  whose shape is overhauled  in typical time scales of  a few months,  
+LSD profiles of  Betelgeuse, whose shape is overhauled  in typical time scales of  a few months.
 Such time scales observed in the Stokes profiles and attributed to the convective time scales 
-appear to explain the strong signal present at 200 days in the LS periodogram of Stokes Q.
+appear to explain the strong signal present at 200 days in the Lomb-Scargle periodogram of Stokes Q.
 
@@
 
-Figure \ref{LS linear polarization} shows the LS periodogram of the total linear polarization of Betelgeuse: $\sqrt{\mathrm{Q^2+U^2}}$. 
-Interestingly, the LS periodogram shows a strong power at 400 days and 2000 days. This signal is much more weaker after the dimming.
-While the LSP is already present in the LS of intensity, here we recover the 400 days periodicity, often invoked in the literature. 
+Figure \ref{LS linear polarization} shows the Lomb-Scargle periodogram of the total linear polarization of Betelgeuse: $\sqrt{\mathrm{Q^2+U^2}}$. 
+Interestingly, the periodogram shows a strong power at 400 days and 2000 days. This signal is much more weaker after the dimming.
+While the LSP is already present in fig. \ref{LS intensity}, here we recover the 400 days periodicity, often invoked in the literature. 
 Nonetheless, the power of this period peaks at a HRV of around 40 km/s. We recall that this velocity is the rest velocity of the star. 
 Every signal beyond this velocity is either due to  plasma sinking in Betelgeuse, or to plumes of plasma rising beyond and above the limb of the star. 
-Such plumes from the back hemisphere haven been often seen in other red supergiants like $\mu$ Cep \citep{lopez_ariste_height_2023}. 
+Such plumes from the back hemisphere have been often seen in other red supergiants like $\mu$ Cep \citep{lopez_ariste_height_2023}. 
 This suggests that both signals at these period of 400 days found in intensity and total linear polarisation 
-could well be due to the periodic appearance of plumes in the back himisphere rising high enough to appear above
+could well be due to the periodic appearance of plumes in the back hemisphere rising high enough to appear above
 the visible limb and generating a supplementary source of brightness. Once again we find a plausible explanation 
@@
 
-Using linear polarization, \cite{lopez_ariste_convective_2018} have been able to reconstruct images of Betelgeuse that have been sucessfully compared 
+Using linear polarization, \cite{lopez_ariste_convective_2018} have been able to reconstruct images of Betelgeuse that have been successfully compared 
 to inteferometric images of \cite{montarges_close_2016}. The images are produced by finding the brightness distribution that better fits the observed linear 
-polarization LSD profile using a Marquardt-Levemberg minimisation.
+polarization LSD profile using a Marquardt-Levenberg minimisation.
 
-Betelgeuse is observed in average every other weeks by the TBL, and hence
-we are able to follow its surface activity through polarimetry imaging. This has  allowed in the past the estimatation of  the size and the
- life time of the convective cells
-at the surface. From these images, we can compute a photo-center, which is sensitive to the size and the number of convective cells. 
-An homogeneous star will have a photo-center coinciding with barycenter of the star, whereas a star with one or two huge convective cells will have a
-photo-center displacement more important, up to a few percent of the stellar radius in the case of RSG \citep{chiavassa_probing_2022}.
+Betelgeuse is observed in average every two or three weeks by the TBL, and hence
+we are able to follow its surface activity through polarimetry imaging. This has  allowed in the past the estimation of  the size and the
+ lifetime of the convective cells
+at the surface. From these images, we computed a photo-center, which is sensitive to the size and the number of convective cells. 
+An homogeneous star will have a photo-center displacement coinciding with barycenter of the star, whereas a star with one or two huge convective cells will have a
+photo-center displacement more important, up to a few percent of the stellar radius in the case of red supergiants \citep{chiavassa_probing_2022}.
 
@@
 
-We computed the LS periodogram of the displacement of the photo-center. But, before going any further, we have to warn
+%We computed the Lomb-Scargle periodogram of the displacement of the photo-center.
+But, before going any further, we have to warn
  about the interpretation of such images. 
@@
 If a particular observation were to be treated independently of previous observations, the solution found by 
-the algorithm could be any of the ambigous solutions possible. To obtain a continuity 
-between the series images, 
-for a given day we use the brightness distribution of the previous day as initial point 
+the algorithm could be any of the ambiguous solutions possible. To obtain a continuity 
+between the series of images, 
+for a given day we used the brightness distribution of the previous day as initial point 
 of the  fitting iteration; and for the first image, we start with a 
@@
     \includegraphics[width=0.5\textwidth]{Lomb-Scargle Photo-center.png}
-    \caption{Lomb-Scargle periodogram of the 100 photo-center displacement of Betelgeuse. Each blue curve corresponds to a LS periodogram of one photo-center
-     displacement. The black line is the average of the 100 LS periodogram. The black dashed lines represent the 2000, 400 and 200 days period respectively 
-     (from left to right). The upper panel is the LS from data before the dimming. The lower panel is the LS from the all dataset, 
-     including data after the great dimming. }
+    \caption{Lomb-Scargle periodogram of the 100 photo-center displacement of Betelgeuse. Each blue curve corresponds to a Lomb-Scargle periodogram of one photo-center
+     displacement. The black line is the average of the 100 periodogram. The black dashed lines represent the 2000, 400 and 200 days period respectively 
+     (from left to right). In the upper panel, only data before the dimming were used to compute the Lomb-Scargle periodogram, while for the lower panel, the whole dataset was used. }
     \label{LS photocenter}
@@
 
-Figure \ref{LS photocenter} shows the LS periodogram of the photo-center displacement  of each one of those 100 series (blue lines) and the average LS periodogram (black line). 
-As before, the upper panel shows the LS from images before the great dimming of Betelgeuse only, whereas the lower panel is the LS for the whole set of 
-data,  including images after the dimming. The black dashed lines mark the 2000, 400 and 
-200 days periods respectively. First, we find the the LS power is high around the LSP, though
+Figure \ref{LS photocenter} shows the Lomb-Scargle periodogram of the photo-center displacement of each one of those 100 series (blue lines) and the average Lomb-Scargle periodogram (black line). 
+As before, the upper panel shows the Lomb-Scargle periodogram from data before the great dimming only, whereas we used the whole set of data for the lower panel.
+The black dashed lines mark the 2000, 400 and 
+200 days periods respectively. First, we find that the Lomb-Scargle power is high around the LSP, though
 uncomfortably close to the auto-correlation peak.
 We also notice a decrease in the Lomb-Scargle power from  before to after the dimming, in 
-concurrence to what was seen in the LS periodograms of the LSD profiles. Before it, there are two smalls bumps around the 400 days period. 
+concurrence to what was seen in the periodograms of the LSD profiles. Before it, there are two smalls bumps around the 400 days period. 
 But those bumps almost completely disappear after the dimming. Even though it is 
@@
 And this is concluded  from images that exclusively involve surface convection.
-The LS of the photo-center displacement points towards a variability explained by convection only. 
-And this is consistent with what we obtained from linear polarization.
+The Lomb-Scargle periodogram of the photo-center displacement points towards a variability explained by convection only,
+and this is consistent with what we obtained from linear polarization.
 %Both the photo-center displacement and the LS of the linear polarization point toward a variability explained by surface convection.
@@
 variability of Betelgeuse is evolving on a timescale shorter than one year. This qualitative change has been pointed out before.
-In the literature, it is reported that before the great dimming the main period of Betelgeuse was of 400 days \citep{kiss_variility_2006}. 
+In the literature, it is reported that before the great dimming the main period of Betelgeuse was of 400 days \citep{kiss_variability_2006}. 
 This period is often associated to the fundamental pressure mode. After the dimming, however this period appears to have disappeared 
-and only the period at 200 days,  associated to the first overtone of the fundamental mode, is visible since \citep{XXX}.
-No explanations based upon pulsations have been ofered concerning this change of variability after or perhaps because of the dimming. But this 
+and only the period at 200 days,  associated to the first overtone of the fundamental mode, is visible since \citep{dupree_great_2022}.
+No explanations based upon pulsations have been offered concerning this change of variability after or perhaps because of the dimming. But this 
 is a behaviour that fully concurs with what we find in the polarization data, and which cannot be related to any stellar pulsation but rather 
@@
 furthermore, taking into consideration data around the dates for which TBL data is available. Periodograms  with such constraints are 
-shown in Fig.\ref{LSAAVSO}. Even when the full dataset for the period is taken into account, the periods present small amplitudes, and are 
+shown in Fig.\ref{LSAAVSO}. Even when the full dataset for the period is taken into account (red line), the periods present small amplitudes, and are 
 fully and favourably comparable with the periodograms presented above. This is even more dramatic when the periodogram is made from data at the exclusive 
-dates at which TBL data is available. It stands out that the low amplitude peaks are rather due to both the sparse cadence and short span of 
+dates at which TBL data is available (blue line). It stands out that the low amplitude peaks are rather due to both the sparse cadence and short span of 
 time of the TBL data series rather than to a physical absence. Actually, the present tests puts more emphasis on the importance of the 
 strong peaks seen at selected wavelengths in the LSD profiles, strong when compared to the peaks at similar periods arising from the light curve.
+
+\begin{figure}[!h]
+    \centering
+    \includegraphics[width=0.5\textwidth]{Light_curve_Betelgeuse.png}
+    \caption{Light curve of Betelgeuse from AAVSO in the V-band.}
+    \label{light curve Betelgeuse}
+\end{figure}
+
+\begin{figure}[!h]
+    \centering
+    \includegraphics[width=\linewidth]{Lomb-Scargle-TBL.png}
+    \caption{Lomb-Scargle periodogram from the AAVSO data between 2013 and 2023 (red line) and from the AAVSO data taken near TBL observations (blue line).}
+    \label{LSAAVSO}
+\end{figure}
 
@@
 % variation in the magnitude of Betelgeuse.
-
-\begin{figure}[!h]
-    \centering
-    \includegraphics[width=0.5\textwidth]{Light_curve_Betelgeuse.png}
-    \caption{Light curve of Betelgeuse from AAVSO in the V-band.}
-    \label{light curve Betelgeuse}
-\end{figure}
 
@@
 We recall that in the present scenario for the formation of the polarization signal, this signal is supposed to remain between 
--20 and 40km/s. Signals above 40km/s are uncommon and can be associated to
+-20 and +40km/s. Signals above 40km/s are uncommon and can be associated to
  plasma rising behind the limb of Betelgeuse. 
@@
 common, two features that prompted a dedicated explanation. In Betelgeuse, such signals, when present, are weaker. However, 
-they are visible from time to time, with regularities that are captured by  the LS periodogram at 400 days.
+they are visible from time to time, with regularities that are captured by  the Lomb-Scargle periodogram at 400 days.
 Without further justification, we may propose that convective plumes in the hidden hemisphere of the star and 
@@
 LSD profiles of stokes Q or U will slightly change when we observe Betelgeuse every two weeks. But they are completely overhauled 
-on time scales of several months. The Lomb-Scargle perioograms appear to capture this spectral dynamic around periods of 200 days.
+on time scales of several months. The Lomb-Scargle periodograms appear to capture this spectral dynamic around periods of 200 days.
  Such changes in the profiles are directly linked to the dynamics of the surface,  to  the evolution and movements 
@@
 
-The interpretaion of the linear polarization profiles is made in terms of convective structures in the atmosphere of Betelgeuse. We suggest 
+The interpretation of the linear polarization profiles is made in terms of convective structures in the atmosphere of Betelgeuse. We suggest 
 that it is in these convective structures and their temporal evolution that we will find the true reason for the observed periodicities both 
@@
 Although of secondary importance, we may suggest with more or less ease what phenomena are at the origin of the main periods. The 400-day period 
-appears to be strongly related to the appearence of plumes in the back hemisphere of the star rising high enough to be seen above the limb.
+appears to be strongly related to the appearance of plumes in the back hemisphere of the star rising high enough to be seen above the limb.
 Such plumes have been described in another RSG, $\mu$ Cep, and while less common or prominent, they are also present in Betelgeuse. The irregularity 
@@
     We acknowledge support from the French National Research Agency (ANR)
-    funded project PEPPER (ANR-20-CE31-0002)
-    An acknowledgement to the AAVSO may be required??????????????????
+    funded project PEPPER (ANR-20-CE31-0002).
+    We acknowledge the observers of the AAVSO International Database who provided more data than enough to produce this work.
+
     \end{acknowledgements}

### 2024-04-16 main.tex
Add files via upload

@@
 Betelgeuse is a prototypical red supergiant (RSG) known to be a semi-regular variable. Several periods can be found in the 
-literature usually clustered around the so-called long secondary period (LSP) of 2000 days and other more or less well defined 
-periods around 400 and 200 days. Such periods are often linked to radial pulsation modes,   and have been used to model 
-the characteristics of the resonant cavity, hence the size of the star and eventually its evolutionary stage \citep{saio_evolutionary_2023}.
+literature usually clustered around the so-called long secondary period (LSP) of 2000 days, and other shorter periods of 400 and 200 days.
+The origin of the LSP remains poorly known, some tried to explain the LSP by the lifetime of the large convective cells present at the surface of RSG (e.g \cite{stothers_giant_2010}) or by the magnetic field \citep{wood_long_2004}.
+The shorter periods of 400 and 200 days seem to have a different origin than the LSP.
+By studying a large sample of RSG, \cite{kiss_variability_2006} attributed the 400 days period to radial pulsation modes, the 200 days period being attributed to the first overtone \citep{joyce_standing_2020}, which was in agreement with the historical work of \cite{stothers_pulsation_1969}.\\
 
-At the end of 2019, Betelgeuse reached a historical minimum in its luminosity, called the great dimming \citep{guinan_fall_2020}. 
-From interferometric data, we speculate that this event was caused by a cloud of dust close to the line of sight \citep{montarges_dusty_2021}. 
-Interferometric images show a huge drop in luminosity in the southern hemisphere of Betelgeuse, leading to this dimming.
-In this event it was clear that a change in brightness was not due to any kind of pulsation but to a change on the
-brightness distribution over the stellar disk due to the presence of large quantities of dust. While this may be seen as 
-a singular event, it brings up the question whether all changes in brightness cannot be due to similar changes in the brightness 
-distribution over the disk, unrelated to any pulsation phenomenon.
-
-Betelgeuse is known to present large convective cells with lifetimes in the order of one to two years and measurable changes 
+Besides radial pulsation, surface activity leading to random brightness variation was also mentioned to explain the variability of Betelgeuse, being firstly theorised by \cite{schwarzschild_scale_1975}.
+Betelgeuse is known to present large convective cells with lifetimes in the order of one to two years \citep{lopez_ariste_convective_2018} and measurable changes
 in the span of one week. Bright convection cells near disk center will increase the integrated brightness of the star when 
@@
 variability could be expected to be random, but will present quasi-periodicities related to the typical time scales of the 
-convective patterns.
+convective patterns. \\
 
+%more or less well defined 
+%periods around 400 and 200 days. 
+%Such periods are often linked to radial pulsation modes,   and have been used to model 
+%the characteristics of the resonant cavity, hence the size of the star and eventually its evolutionary stage \citep{saio_evolutionary_2023}.
+
+At the end of 2019, Betelgeuse reached a historical minimum in its luminosity, called the great dimming \citep{guinan_fall_2020}. 
+From interferometric data, it is speculated that this event was caused by a cloud of dust close to the line of sight \citep{montarges_dusty_2021}. 
+Interferometric images show a drop in luminosity in the southern hemisphere of Betelgeuse, that might be caused by a mass loss event and leading to this dimming \citep{dupree_great_2022}.
+Other theories such as a drop in temperature was also mentioned by \cite{harper_photospheric_2020}, or an increase in the molecular opacity \citep{kravchenko_atmosphere_2021}.
+In this event it was clear that a change in brightness was not due to any kind of pulsation but to a change on the
+brightness distribution over the stellar disk.
+%due to the presence of large quantities of dust.
 Since the end of the Great Dimming event, Betelgeuse has continue its random variation of brightness. 
@@
 it has been shown by \cite{jadlovsky_analysis_2023} and \cite{dupree_great_2022} that the periodicity of Betelgeuse has changed after the dimming. Using the light curves from AAVSO, 
-the authors showed that before the dimming, the dominant period of Betelgeuse was the 400 days period. However, after the dimming, this period has changed and 
-is now shorter, around 200 days.
-Therefore, after the dimming, there was a change in the behaviour of the atmosphere of Betelgeuse: the time scales on which Betelgeuse evolves are now shorter.
-Our analysis is based on this change of periodicity.
-Faced with the dramatic change in stellar parameters and age that such change in a pulsation period would 
-require, our interest in an alternative explanation of the variability
-in terms  of convective patterns arose. In order to address this question we shall seek for such typical periods of 2000, 400 and 200 days in  
-observational proxies related to the convective activity but unrelated to pulsations and which cannot be directly linked to brightness.
+the authors showed that before the dimming, the dominant period of Betelgeuse was the 400 days period, while after the great dimming the main periods of Betelgeuse have shortened, oscillating between 97 days and 230 days \citep{dupree_great_2022}, revealing a change in the behaviour of the atmosphere of Betelgeuse.
+While the great dimming may be seen as 
+a singular event, it brings up the question whether all changes in brightness cannot be due to similar changes in the brightness 
+distribution over the disk, unrelated to any pulsation phenomenon. 
+Our interest in an alternative explanation of the variability
+in terms  of convective patterns arose. In order to address this question we shall seek for such typical periods of 400 and 200 days in  
+observational proxies related to the convective activity.\\
+%but unrelated to pulsations and which cannot be directly linked to brightness.
+%Faced with the dramatic change in stellar parameters and age that such change in a pulsation period would 
+%require,
 
@@
 different lines in the spectrum of Betelgeuse. The images produced compare very well with contemporaneous images made with interferometric 
-techniques \citep{montarges_close_2016} and show clear convective patterns, akin to solar granulation.
+techniques \citep{montarges_close_2016} and show clear convective patterns, akin to solar granulation.\
 
@@
 pulsations but linked to the brightness inhomogeneities due to the convective patterns in Betelgeuse. If one can find the aforementioned periods 
-in linear polarization, we must conclude that these periods are related to the convective activity which originates those linear polarization signals
-and unrelated to any radial pulsation phenomena. This is the purpose of the present work.
+in linear polarization, we must conclude that these periods are related to the convective activity which originates those linear polarization signals.
+%and unrelated to any radial pulsation phenomena. 
+This is the purpose of the present work.
+In section \ref{Section 2} we describe the dataset of linear polarization obtained with the TBL and the Lomb-Scargle periodogram used to derive periods.
+In section \ref{section 3} we associate the 200 days period to convective timescale of the smallest granules and the 330 days periodicity to the largest granules, and we provide an explanation in the change of variability of Betelgeuse before and after the great dimming.
+In section \ref{Conclusion}, we summup our work.
 
 \section{Searching for periodicities in the polarization spectra of Betelgeuse}
+\label{Section 2}
 
-Betelgeuse has been observed for the last 13 years with Narval and Neo-Narval at the Telescope Bernard Lyot (TBL). These two instruments measure the polarization 
+Betelgeuse has been observed for the last 10 years with Narval and Neo-Narval at the Telescope Bernard Lyot (TBL). These two instruments measure the polarization 
 over the visible spectra of Betelgeuese (390-1000 nm) with high spectral resolution (R=65000) and high polarimetric sensitivity. Despite 
@@
 
+\subsection{The variability present in the stokes parameters}
+
 Trying to find periods in an astrophysical context carries the hurdle of unevenly spaced observed data. To overcome this issue, 
@@
 using least-squares at each frequency between the first and the last observation. The better the fit, the higher the power attributed to the frequency. 
-We applied this technique to the  LSD profiles of Stokes I, U and Q of Betelgeuse observed at the TBL during 
-that span of 13 years.
+We applied this technique to the  LSD profiles of Stokes $I$, $U$, $Q$ and to the total linear polarization of Betelgeuse observed by the TBL during 
+that span of 10 years.
 
-We first computed the Lomb-Scargle periodogram at each wavelength of the LSD profile, between -25 km/s and 80 km/s to span the entire 
+We first computed the Lomb-Scargle periodogram at each wavelength of the LSD profile, between -25 km/s and +60 km/s to span the entire 
 signal present in the LSD profile. \cite{lopez_ariste_convective_2018} found the most blueshifted signal towards -20 km/s, 
-and interpreted it as  the maximum velocity of the rising plasma, seen from the Earth. The most redshifted signal was found at 40 km/s. 
+and interpreted it as  the maximum velocity of the rising plasma, seen from the Earth. The most redshifted signal was found at +40 km/s. 
 This velocity is interpreted
-as the rest velocity of the star in our reference frame. Seldom, signals at velocities larger than 40 km/s can be found 
+as the rest velocity of the star in our reference frame. Seldom, signals at velocities larger than +40 km/s can be found 
 and can be interpreted as either dark and cold plasma falling back to Betelgeuse, 
@@
 
-Figures \ref{LS intensity},\ref{LS Q},\ref{LS U} and \ref{LS linear polarization} show the Lomb-Scargle periodogram in normalised power of the LSD profile of respectively 
-the intensity, stokes Q, stokes U and the linear polarization. 
+Figures \ref{LS intensity},\ref{LS Q},\ref{LS U} and \ref{LS linear polarization} show the Lomb-Scargle periodogram of the LSD profile of respectively 
+the intensity, stokes $Q$, stokes $U$ and the linear polarization. 
 %The abscissa is the frequency (in $\mathrm{days^{-1}}$), while the ordinate is the heliocentric radial velocity in km/s. 
-For each figure, the upper panel is the Lomb-Scargle periodogram with data taken before the great dimming of Betelgeuse, that is from 2013 to the beginning of 2019. 
-The lower panel is the Lomb-Scargle periodogram for the whole dataset, before and after the dimming: from 2013 to 2023. The white dashed lines mark the different 
-periods found in the literature: the LSP at 2000 days, the fundamental pressure mode at 400 days and its first overtone mode at 200 days.
+For each figure, the uper panel is the Lomb-Scargle periodogram of each velocity bin (black line) and the total average (red line).
+The green line corresponds to the window function.
+The lower panel is the Lomb-Scargle periodogram represented for each velocity bin, while the white dotted line refer to 400 and 200 days.
+The right panel is the mean profile of each stokes parameter for each velocity bin.
 
@@
     \centering
-    \includegraphics[width=0.5\textwidth]{Lomb-Scargle Intensity.png}
+    \includegraphics[width=0.5\textwidth]{Lomb-Scargle Intensity.pdf}
     \caption{Lomb-Scargle periodogram of the LSD profile of intensity.
-    The upper panel is the Lomb-Scargle periodogram for data before the great dimming while the lower panel is the periodogram from the whole dataset.
-    The ordinate refers to the velocity in the heliocentric radial velocity in km/s.
-    The three white dashed lines represent respectively (from left to right) the LSP at 2000 days, the 400 days period and its first overtone at 200 days.  }
+    The upper panel is the Lomb-Scargle periodogram for each velocity bin (black line) and the average (red line). The green line is the window function.
+    In the lower pannel, the two white dashed lines represent respectively the 400 days and the 200 days periods. The right panel is the average intensity profile for each velocity bin.}
     \label{LS intensity}
@@
     \centering
-    \includegraphics[width=0.5\textwidth]{Lomb-Scargle Stokes Q.png}
+    \includegraphics[width=0.5\textwidth]{Lomb-Scargle Stokes Q.pdf}
     \caption{Same as figure \ref{LS intensity} for stokes Q. }
@@
     \centering
-    \includegraphics[width=0.5\textwidth]{Lomb-Scargle Stokes U.png}
+    \includegraphics[width=0.5\textwidth]{Lomb-Scargle Stokes U.pdf}
     \caption{Same as figure \ref{LS intensity} for stokes U.}
@@
     \centering
-    \includegraphics[width=0.5\textwidth]{Lomb-Scargle linear polarization.png}
+    \includegraphics[width=0.5\textwidth]{Lomb-Scargle linear polarization.pdf}
     \caption{Same as figure \ref{LS intensity} for linear polarization: $\sqrt{\mathrm{Q^2}+\mathrm{U^2}}$.}
@@
 %\subsection{Intensity}
-Figure \ref{LS intensity} shows the Lomb-Scargle periodogram of the LSD of the intensity profile. Interestingly, we recover the LSP at 2000 days, 
-but located at a HRV of 20 km/s rather than over the whole profile. This periodicity is still present after the  dimming, but  it is located at a HRV between 0 and 10 km/s. 
-There is also a weaker signal at the fundamental mode, around 400 days before the dimming. But this signal completely disappears after the dimming. From these observations
-we find  hard to attribute 
-the observed Lomb-Scargle power to a global period of Betelgeuse.
-%, however it might be a hint that this periodicity has changed after the dimming. 
-Our results 
-in the intensity profile are consistent with the one from \cite{mathias_evolution_2018} who used 2D Fourier analysis and found that the dominant period was the LSP.
+Figure \ref{LS intensity} shows the Lomb-Scargle periodogram of the LSD of the intensity profile.
+The 400 and 200 days period seem to be recovered by the periodogram.
+However, the 400 days period is uncomfortably close to the peak of the window function, while the 200 days period is not.
+We also notice that the two periods are present at the same HRV.
+Interestingly, \cite{mathias_evolution_2018} also found this 200 days periodicity with a shorter observation period.
 
-\subsection{Stokes Q and U}
+Figure \ref{LS Q} and \ref{LS U} show the Lomb-Scargle periodogram of both stokes $Q$ and $U$. It is clear that the periodogram is completely different 
+from the intensity one.
+A strong signal is present in fig.
+\ref{LS Q} around 200 days.
+Concerning stokes U in fig.
+\ref{LS U}, it appears that the main period is around $0.003 \ \mathrm{days^{-1}}$ (330 days).
+Other period such as the one at 200 days or 250 days ($0.004 \ \mathrm{days^{-1}}$) are present but difficult to trust.
+From stokes $Q$ and $U$, we recover the 200 days period and the 330 days period is more or less close to the 400 days period found in the literature.
+Figure \ref{LS linear polarization} shows the Lomb-Scargle periodogram of the total linear polarization of Betelgeuse: $\sqrt{\mathrm{Q^2+U^2}}$. 
+Interestingly, the periodogram shows a strong power at both 330 days and 200 days, which are the period that we have found in both stokes $Q$ and $U$.
 
-Figure \ref{LS Q} and \ref{LS U} show the Lomb-Scargle periodogram of both stokes Q and U. It is clear that the periodogram is completely different 
-from the intensity one. A strong signal is present in fig. \ref{LS Q} around 200 days. Once again, this signal is weaker after the dimming.
-Concerning stokes U in fig. \ref{LS U}, it appears that we only recover the LSP, which once again, is weaker after the great dimming. 
-Those periodograms rise a lot of questions: why is the 200 days period only present in stokes Q? Why not in stokes U or intensity? 
-We will discuss about those questions in the next section. It also appears that stokes Q is
-much more sensitive to the 200 days period than the LSP. From this, we can conclude that stokes
-Q evolves on a time scale much more closer to the first overtone 
-mode than the LSP or even the fundamental mode. Since Stokes Q is sensitive to different periods, we must conclude that it is due to other mechanisms
-than the one behind the periodicities observed in intensity. Linear polarization in Betelgeuse is interpreted to be a proxy of  convective cells,
-so we can speculate that the signal present in the periodogram 
-may be related to  the lifetime of convective cells. Such hypothesis is comforted by what we see in the 
-LSD profiles of  Betelgeuse, whose shape is overhauled  in typical time scales of  a few months.
-Such time scales observed in the Stokes profiles and attributed to the convective time scales 
-appear to explain the strong signal present at 200 days in the Lomb-Scargle periodogram of Stokes Q.
-
-\subsection{The linear polarization}
-
-Figure \ref{LS linear polarization} shows the Lomb-Scargle periodogram of the total linear polarization of Betelgeuse: $\sqrt{\mathrm{Q^2+U^2}}$. 
-Interestingly, the periodogram shows a strong power at 400 days and 2000 days. This signal is much more weaker after the dimming.
-While the LSP is already present in fig. \ref{LS intensity}, here we recover the 400 days periodicity, often invoked in the literature. 
-Nonetheless, the power of this period peaks at a HRV of around 40 km/s. We recall that this velocity is the rest velocity of the star. 
-Every signal beyond this velocity is either due to  plasma sinking in Betelgeuse, or to plumes of plasma rising beyond and above the limb of the star. 
-Such plumes from the back hemisphere have been often seen in other red supergiants like $\mu$ Cep \citep{lopez_ariste_height_2023}. 
-This suggests that both signals at these period of 400 days found in intensity and total linear polarisation 
-could well be due to the periodic appearance of plumes in the back hemisphere rising high enough to appear above
-the visible limb and generating a supplementary source of brightness. Once again we find a plausible explanation 
-of these periods unrelated to pulsations, but rather to the convective dynamics of the star.
+%Nonetheless, the power of this period peaks at a HRV of around 40 km/s. We recall that this velocity is the rest velocity of the star. 
+%Every signal beyond this velocity is either due to  plasma sinking in Betelgeuse, or to plumes of plasma rising beyond and above the limb of the star. 
+%Such plumes from the back hemisphere have been often seen in other red supergiants like $\mu$ Cep \citep{lopez_ariste_height_2023}. 
+%This suggests that both signals at these period of 400 days found in intensity and total linear polarisation 
+%could well be due to the periodic appearance of plumes in the back hemisphere rising high enough to appear above
+%the visible limb and generating a supplementary source of brightness. Once again we find a plausible explanation 
+%of these periods unrelated to pulsations, but rather to the convective dynamics of the star.
 
@@
     \centering
-    \includegraphics[width=0.5\textwidth]{Lomb-Scargle Photo-center.png}
-    \caption{Lomb-Scargle periodogram of the 100 photo-center displacement of Betelgeuse. Each blue curve corresponds to a Lomb-Scargle periodogram of one photo-center
-     displacement. The black line is the average of the 100 periodogram. The black dashed lines represent the 2000, 400 and 200 days period respectively 
-     (from left to right). In the upper panel, only data before the dimming were used to compute the Lomb-Scargle periodogram, while for the lower panel, the whole dataset was used. }
+    \includegraphics[width=0.5\textwidth]{Lomb-Scargle Photo-center.pdf}
+    \caption{Lomb-Scargle periodogram of the 100 photo-center displacement of Betelgeuse. Each blue lines correspond to a Lomb-Scargle periodogram of one photo-center
+     displacement. The black line is the average of the 100 periodogram. The red dashed lines represent the 400 and 200 days period respectively 
+     (from left to right). As before, the green line is the window function.}
     \label{LS photocenter}
@@
 
-Figure \ref{LS photocenter} shows the Lomb-Scargle periodogram of the photo-center displacement of each one of those 100 series (blue lines) and the average Lomb-Scargle periodogram (black line). 
-As before, the upper panel shows the Lomb-Scargle periodogram from data before the great dimming only, whereas we used the whole set of data for the lower panel.
-The black dashed lines mark the 2000, 400 and 
-200 days periods respectively. First, we find that the Lomb-Scargle power is high around the LSP, though
-uncomfortably close to the auto-correlation peak.
-We also notice a decrease in the Lomb-Scargle power from  before to after the dimming, in 
-concurrence to what was seen in the periodograms of the LSD profiles. Before it, there are two smalls bumps around the 400 days period. 
-But those bumps almost completely disappear after the dimming. Even though it is 
-hard to identify the 400 days period from those bumps alone,
-we still find that there is a difference before and after the dimming. 
-And this is concluded  from images that exclusively involve surface convection.
-The Lomb-Scargle periodogram of the photo-center displacement points towards a variability explained by convection only,
-and this is consistent with what we obtained from linear polarization.
+Figure \ref{LS photocenter} shows the Lomb-Scargle periodogram of the photo-center displacement of each one of those 100 series (blue lines) and the average Lomb-Scargle periodogram (black line), 
+%As before, the upper panel shows the Lomb-Scargle periodogram from data before the great dimming only, whereas we used the whole set of data for the lower panel.
+while the black dashed lines mark the 400 and 
+200 days periods respectively.
+As before, the window function is represented by the green line.
+Interestingly, we find two bumps around the 400 days period.
+We cannot directly confirm that we recover the 400 days period from those small bumps, but since polarimetry imaging involves only surface convection, this is an argument that points toward a variability explained by surface convection.
+%First, we find that the Lomb-Scargle power is high around the LSP, though
+%uncomfortably close to the auto-correlation peak.
+%We also notice a decrease in the Lomb-Scargle power from  before to after the dimming, in 
+%concurrence to what was seen in the periodograms of the LSD profiles. Before it, there are two smalls bumps around the 400 days period. 
+%But those bumps almost completely disappear after the dimming. Even though it is 
+%hard to identify the 400 days period from those bumps alone,
+%we still find that there is a difference before and after the dimming. 
+%And this is concluded  from images that exclusively involve surface convection.
+%The Lomb-Scargle periodogram of the photo-center displacement points towards a variability explained by convection only,
+%and this is consistent with what we obtained from linear polarization.
 %Both the photo-center displacement and the LS of the linear polarization point toward a variability explained by surface convection.
 
-\subsection{The variability of the light curve of Betelgeuse}
+\section{The variability of the light curve of Betelgeuse}
+\label{section 3}
+%Before summing up the conclusions of this work, we can afford proposing alternative scenarios for the observed periodicities in the light curve, 
+%the LSD profiles and the displacement of the photo-center of the inferred photospheric image reconstructions. 
+The presence of the same 
+periods and, after taking into consideration the time span and sparsity of the TBL data series, the larger amplitude of the peaks in
+polarization data  than in the light curve point to a common origin, one that is rather based upon convective dynamics.
+Since we associate stokes $Q$ and $U$ to proxies of convection, we associate the 200 days period to the typical timescales on which the Stokes parameters evolve.
+The LSD profiles of stokes $Q$ or $U$ will slightly change when we observe Betelgeuse every month, but are completely overhauled 
+on time scales of several months. The Lomb-Scargle periodograms appear to capture this spectral dynamic around periods of 200 days.
+ Such changes in the profiles are directly linked to the dynamics of the surface,  to  the evolution and movements 
+of convective cells over the visible hemisphere. So the 200-day period appears to be related to the evolution of the convective patterns 
+on the surface of Betelgeuse.  The fact that this peak is strongly present in stokes $Q$ but less in $U$ points toward a more stochastic origin, which can fully be explained by convection.\
+
+The other period found in the linear polarization and stokes $U$, the 330 days periods, is more or less close to the 400 days period often present in the literature (e.g \cite{kiss_variability_2006} found a period $388 \pm 30$ days).
+It has been shown by \cite{lopez_ariste_convective_2018} that large convective cells can live up to one year, thus this period would be associated to the timescale on which the largest granules evolve.
+This typical timescale has also been observed in the numerical simulations and might play an important role in the determination of stellar distances through the displacement of the photo-center \citep{chiavassa_probing_2022}.
+Once again, the 330 days periodicity being more present in stokes $U$ than $Q$ unravels a stochastic process that is due to convective motions.
+\
+
+Finally, the intensity profile seems to match the 400 days periods, but is uncomfortably close to a peak of the window function of the TBL and a secondary peak close to 200 days.
+Since we are able to recover the 200 days period in the stokes paramters, we attribute this period to the convective timescales, but concerning the 400 days period, it could be a mix of the typical convective timescales of large granules associated with stochastic pulsations exited by convective cavities.
+Up to this point, the model developed to interpret linear polarization does not involve pulsations, that is why we interpret the 200 days period as the convective turnover timescale.
+The 400 days period involve the intensity profile, thus the 400 days period is probably caused by both convective timescales and stochastic pulsations.
+ %As showed by
+%\cite{lopez_ariste_convective_2018}, some of those convective cells can live up for years. We may speculate
+%that the LSP may be related to those long living structures. However the present span of data from the TBL is insufficient to do more than
+%just speculate. It should be noticed that  
+% the peak is not present in stokes U, and this may point to a more stochastic event than just long-living structures.
+
+\subsection{The light curve of Betelgeuse from the point of view of convection}
 
 After examining the periods found by the Lomb-Scargle technique on the polarization data, and its derivatives, obtained by the TBL over the last 
-13 years, it is worth putting them in the context of the periods found traditionally on light curves over the same period of time. Two aspects 
-are of our interest in this comparison: the behaviour of the light curve before and after the great dimming, and the amplitudes of the peaks 
-at the main cited periods.
+10 years, it is worth putting them in the context of the periods found traditionally on light curves over the same period of time. Two aspects 
+are of our interest in this comparison: the behaviour of the light curve before and after the great dimming.
+%and the amplitudes of the peaks 
+%at the main cited periods.
 
@@
 
-We have recovered the light curve of Betelgeuse in the visible from the AAVSO database for the last 13 years (Figure \ref{light curve Betelgeuse}).
-Before the great dimming, the magnitude of Betelgeuse evolves on a timescale of a year, whereas after the dimming, the variability of Betelgeuse has changed. 
+We have recovered the light curve of Betelgeuse in the visible from the AAVSO database for the last 10 years (Figure \ref{light curve Betelgeuse}).
+Before the great dimming, the magnitude of Betelgeuse evolves on a timescale of a year, whereas after the dimming, the variability of Betelgeuse has shortened. 
 It is clear from fig. \ref{light curve Betelgeuse} that the 
-variability of Betelgeuse is evolving on a timescale shorter than one year. This qualitative change has been pointed out before.
-In the literature, it is reported that before the great dimming the main period of Betelgeuse was of 400 days \citep{kiss_variability_2006}. 
-This period is often associated to the fundamental pressure mode. After the dimming, however this period appears to have disappeared 
-and only the period at 200 days,  associated to the first overtone of the fundamental mode, is visible since \citep{dupree_great_2022}.
-No explanations based upon pulsations have been offered concerning this change of variability after or perhaps because of the dimming. But this 
-is a behaviour that fully concurs with what we find in the polarization data, and which cannot be related to any stellar pulsation but rather 
-convective dynamics. Beyond the actual values of the periods found in a Lomb-Scargle periodogram, we also see a correspondence between polarization 
-and the light curve in these changes in the observed periods before and after the great dimming. Such concurrence points to an explanation 
-in terms of convective dynamics also for the variability of the light curve.
+variability of Betelgeuse is evolving on a timescale shorter than one year. This qualitative change has been pointed out before,
+before the great dimming the main period of Betelgeuse was of 400 days \citep{kiss_variability_2006}. 
+This period is often associated to the fundamental pressure mode. After the dimming however, this period appears to have disappeared 
+and only timescales shorter than 230 days are visible since \citep{dupree_great_2022}. 
+No explanations have been offered concerning this change of variability after or perhaps because of the dimming, thus we can afford proposing alternative scenarios for the observed periodicities in the light curve.
+\
 
-It may be argued that the periodograms presented in the previous sections and, in particular, those  of the photo-center displacements, 
-present low amplitude peaks at the referred main periods. One may wonder whether we are over-interpreting mere fluctuations of the periodograms, 
-while the peaks seen in the literature from analysis of the light curve are unmistakable. It is therefore worth to take advantage of the 
-availability of the AAVSO data to perform a Lomb-Scargle periodogram of the light curve in the same period covered by our polarization data and,
-furthermore, taking into consideration data around the dates for which TBL data is available. Periodograms  with such constraints are 
-shown in Fig.\ref{LSAAVSO}. Even when the full dataset for the period is taken into account (red line), the periods present small amplitudes, and are 
-fully and favourably comparable with the periodograms presented above. This is even more dramatic when the periodogram is made from data at the exclusive 
-dates at which TBL data is available (blue line). It stands out that the low amplitude peaks are rather due to both the sparse cadence and short span of 
-time of the TBL data series rather than to a physical absence. Actually, the present tests puts more emphasis on the importance of the 
-strong peaks seen at selected wavelengths in the LSD profiles, strong when compared to the peaks at similar periods arising from the light curve.
+RSGs encounter mass loss event, where due to stellar winds, plasma is ejected in the interstellar medium.
+Those events have already been seen in RSG like $\mu$Cep, where \cite{lopez_ariste_height_2023} found an excess of linear polarization beyond the limb of the star and associated it to rising plumes of plasma in the back hemisphere of the star.
+In the case of Betelgeuse, this kind of event happened only once or twice during the last 10 years, with no change in the variability of the star.
+However, since the great dimming, the variability of Betelgeuse has changed, meaning that this event was such different from the usual mass loss event, that it hustled the dynamics of the photosphere.
+\
+
+Considering a scenario where the variability of Betelgeuse is mostly governed by convection, before the great dimming, the variability of the star was linked to the typical timescale on which convection occurs, that is 200 days for the small structures and roughly 350 days for the large structures.
+The great dimming impacted the behaviour of the photosphere, such that the largest structures do not evolve on timescale of the order of 350 days but rather 200 days, that is the smallest convective structures.
+We hypothesise that since the dimming, the photosphere is not in equilibrium and the largest convective structures are destroyed by the turbulent motion of the photosphere.
+Therefore, it is hard to identify periods after the great dimming, which \cite{dupree_great_2022} noticed.
+We expect to recover the 400 days variability within the next years, which corresponds to the time for the photosphere to go back to equilibrium.
+%But this 
+%is a behaviour that fully concurs with what we find in the polarization data, and which cannot be related to any stellar pulsation but rather 
+%convective dynamics. Beyond the actual values of the periods found in a Lomb-Scargle periodogram, we also see a correspondence between polarization 
+%and the light curve in these changes in the observed periods before and after the great dimming. Such concurrence points to an explanation 
+%in terms of convective dynamics also for the variability of the light curve.
+
+%It may be argued that the periodograms presented in the previous sections and, in particular, those  of the photo-center displacements, 
+%present low amplitude peaks at the referred main periods. One may wonder whether we are over-interpreting mere fluctuations of the periodograms, 
+%while the peaks seen in the literature from analysis of the light curve are unmistakable. It is therefore worth to take advantage of the 
+%availability of the AAVSO data to perform a Lomb-Scargle periodogram of the light curve in the same period covered by our polarization data and,
+%furthermore, taking into consideration data around the dates for which TBL data is available. Periodograms  with such constraints are 
+%shown in Fig.\ref{LSAAVSO}. Even when the full dataset for the period is taken into account (red line), the periods present small amplitudes, and are 
+%fully and favourably comparable with the periodograms presented above. This is even more dramatic when the periodogram is made from data at the exclusive 
+%dates at which TBL data is available (blue line). It stands out that the low amplitude peaks are rather due to both the sparse cadence and short span of 
+%time of the TBL data series rather than to a physical absence. Actually, the present tests puts more emphasis on the importance of the 
+%strong peaks seen at selected wavelengths in the LSD profiles, strong when compared to the peaks at similar periods arising from the light curve.
 
@@
 
-\begin{figure}[!h]
-    \centering
-    \includegraphics[width=\linewidth]{Lomb-Scargle-TBL.png}
-    \caption{Lomb-Scargle periodogram from the AAVSO data between 2013 and 2023 (red line) and from the AAVSO data taken near TBL observations (blue line).}
-    \label{LSAAVSO}
-\end{figure}
+%\begin{figure}[!h]
+%    \centering
+%    \includegraphics[width=\linewidth]{Lomb-Scargle-TBL.png}
+%    \caption{Lomb-Scargle periodogram from the AAVSO data between 2013 and 2023 (red line) and from the AAVSO data taken near TBL observations (blue line).}
+%    \label{LSAAVSO}
+%\end{figure}
 
@@
 
-\section{Surface convection suffices to explain the observed variability}
-
-Before summing up the conclusions of this work, we can afford proposing alternative scenarios for the observed periodicities in the light curve, 
-the LSD profiles and the displacement of the photo-center of the inferred photospheric image reconstructions. The presence of the same 
-periods and, after taking into consideration the time span and sparsity of the TBL data series, the larger amplitude of the peaks in
-polarization data  than in the light curve point to a common origin, one that is rather based upon convective dynamics and not to radial or 
-other global pulsations.
+%and not to radial or 
+%other global pulsations.
 
@@
 
-The 400-day period appears prominently in both the intensity and the total polarization profiles. And in both cases, it is mostly present 
-at wavelengths in the red wing, often beyond the 40 km/s value.
-We recall that in the present scenario for the formation of the polarization signal, this signal is supposed to remain between 
--20 and +40km/s. Signals above 40km/s are uncommon and can be associated to
- plasma rising behind the limb of Betelgeuse. 
-Such scenario has been observed in the RSG $\mu$ Cep and was introduced by \cite{lopez_ariste_height_2023} to explain the excess of 
-linear polarization beyond  the rest velocity of the star, which for Betelgeuse is fixed at +40km/s. 
-In $\mu$ Cep, such  linear polarization signal beyond the limit of the rest velocity of the star is very strong and particularly 
-common, two features that prompted a dedicated explanation. In Betelgeuse, such signals, when present, are weaker. However, 
-they are visible from time to time, with regularities that are captured by  the Lomb-Scargle periodogram at 400 days.
-Without further justification, we may propose that convective plumes in the hidden hemisphere of the star and 
-rising above the limb from time to time are responsible for the  400 days period in the light curve. When such plumes rise high enough 
-to be geometrically visible above the limb they show signals in intensity and polarization at redshifted wavelengths. They also
- increase the emitting surface of the star and  the number of photons that we receive,
-leading to an increase of the brightness visible in the light curve. After a few weeks or months, the plume falls back to Betelgeuse, 
-disappearing from our point of view,  leading to a decrease of the brightness.
+%The 400-day period appears prominently in both the intensity and the total polarization profiles. And in both cases, it is mostly present 
+%at wavelengths in the red wing, often beyond the 40 km/s value.
+%We recall that in the present scenario for the formation of the polarization signal, this signal is supposed to remain between 
+%-20 and +40km/s. Signals above 40km/s are uncommon and can be associated to
+ %plasma rising behind the limb of Betelgeuse. 
+%Such scenario has been observed in the RSG $\mu$ Cep and was introduced by \cite{lopez_ariste_height_2023} to explain the excess of 
+%linear polarization beyond  the rest velocity of the star, which for Betelgeuse is fixed at +40km/s. 
+%In $\mu$ Cep, such  linear polarization signal beyond the limit of the rest velocity of the star is very strong and particularly 
+%common, two features that prompted a dedicated explanation. In Betelgeuse, such signals, when present, are weaker. However, 
+%they are visible from time to time, with regularities that are captured by  the Lomb-Scargle periodogram at 400 days.
+%Without further justification, we may propose that convective plumes in the hidden hemisphere of the star and 
+%rising above the limb from time to time are responsible for the  400 days period in the light curve. When such plumes rise high enough 
+%to be geometrically visible above the limb they show signals in intensity and polarization at redshifted wavelengths. They also
+ %increase the emitting surface of the star and  the number of photons that we receive,
+%leading to an increase of the brightness visible in the light curve. After a few weeks or months, the plume falls back to Betelgeuse, 
+%disappearing from our point of view,  leading to a decrease of the brightness.
 
-The appearance of such high-rising plumes is not necessarily periodic. And the star can easily shift from active periods when such 
-plumes are a common feature to calmer periods when no plume is visible for months. $\mu$ Cep may be in one such active period presently,
-while Betelgeuse may have experienced recently one great such event in the direction of the Earth that, after dust formed, became the great 
-dimming, but since then no other high-rising plume has appeared, explaining the absence of the 400-day period since the great dimming. 
-From the point of view of convective dynamics we can offer an explanation of the 400-day period which not only explains the observation of 
-this period both in the light curve and in the polarization profiles, but also its presence at redshifted wavelengths and its present (and probably 
-temporary) disappearance since the great dimming.
+%The appearance of such high-rising plumes is not necessarily periodic. And the star can easily shift from active periods when such 
+%plumes are a common feature to calmer periods when no plume is visible for months. $\mu$ Cep may be in one such active period presently,
+%while Betelgeuse may have experienced recently one great such event in the direction of the Earth that, after dust formed, became the great 
+%dimming, but since then no other high-rising plume has appeared, explaining the absence of the 400-day period since the great dimming. 
+%From the point of view of convective dynamics we can offer an explanation of the 400-day period which not only explains the observation of 
+%this period both in the light curve and in the polarization profiles, but also its presence at redshifted wavelengths and its present (and probably 
+%temporary) disappearance since the great dimming.
 
@@
 
-Concerning the 200 days period, we notice that it coincides with  the typical timescales on which the Stokes parameters evolve. 
-The 
-LSD profiles of stokes Q or U will slightly change when we observe Betelgeuse every two weeks. But they are completely overhauled 
-on time scales of several months. The Lomb-Scargle periodograms appear to capture this spectral dynamic around periods of 200 days.
- Such changes in the profiles are directly linked to the dynamics of the surface,  to  the evolution and movements 
-of convective cells over the visible hemisphere. So the 200-day period appears to be related to the evolution of the convective patterns 
-on the surface of Betelgeuse. 
- As showed by
-\cite{lopez_ariste_convective_2018}, some of those convective cells can live up for years. We may speculate
-that the LSP may be related to those long living structures. However the present span of data from the TBL is insufficient to do more than
-just speculate. It should be noticed that  
- the peak is not present in stokes U, and this may point to a more stochastic event than just long-living structures.
-
 % To summarise what we have done so far: we were able to explain the variability of Betelgeuse using surface convection only. The convective cells play 
@@
 \section{Conclusion}
+\label{Conclusion}
 
-The main observation  presented in this work is that the very same periods identified in the light curve of Betelgeuse are also seen in the intensity and polarization
-spectra measured with the TBL. Periods in the light curve have often been traditionally associated to pulsation due to pressure modes. But 
-such pulsation would in no case affect the polarisation signals nor would explain the fact that the periods can be observed with stronger 
-amplitude at selected wavelengths but no others.  Pulsation appears to offer no coherent explanation to these observations, are hardly to the 
-observed changes in the variability since the great dimming.  The concurrence of observational facts, actual period values, evolution before 
-and after the dimming, the larger amplitude of the period peaks in polarizarion data, points towards a common origin for those variabilities. 
-And such origin cannot be pulsations.
-
+The main observation  presented in this work is that the same periods identified in the light curve of Betelgeuse are also seen in the intensity and polarization
+spectra measured with the TBL. Periods in the light curve have often been traditionally associated to pulsation due to pressure modes.
+%But such pulsation would in no case affect the polarisation signals nor would explain the fact that the periods can be observed with stronger 
+%amplitude at selected wavelengths but no others. % Pulsation appears to offer no coherent explanation to these observations, are hardly to the 
+%observed changes in the variability since the great dimming.  The concurrence of observational facts, actual period values, evolution before 
+%and after the dimming, the larger amplitude of the period peaks in polarizarion data, points towards a common origin for those variabilities. 
+%And such origin cannot be pulsations.
 The interpretation of the linear polarization profiles is made in terms of convective structures in the atmosphere of Betelgeuse. We suggest 
 that it is in these convective structures and their temporal evolution that we will find the true reason for the observed periodicities both 
-in the light curve and in the spectropolarimetric observations. This is the main conclusion of this work: the observed periods are related 
-to convective dynamics and not to radial pulsations of the star.
+in the light curve and in the spectropolarimetric observations.
+This is the main conclusion of this work: since we are able to recover the different periods in the linear polarization profiles, we associate the 200 days periodicyt to convective dynamics while the 400 days periodicity is associated to both stochastic pulsations and convective timescales of the largest granules.%the observed periods are related
+%to convective dynamics and not to radial pulsations of the star.
 
-Although of secondary importance, we may suggest with more or less ease what phenomena are at the origin of the main periods. The 400-day period 
-appears to be strongly related to the appearance of plumes in the back hemisphere of the star rising high enough to be seen above the limb.
-Such plumes have been described in another RSG, $\mu$ Cep, and while less common or prominent, they are also present in Betelgeuse. The irregularity 
-of their presence justifies the 400-day period, which is captured by the Lomb-Scargle periodogram in the polarization signals attributed to 
-this phenomenon. The redshifted wavelengths at which this periodicity is found are justified by the plume moving away of the star and of the Earth while 
-raising in the back hemisphere, while the brightness increase is justified by the extra surface of emitting plasma visible. The 200-day period is less 
-prominent but appears to be correlated to the typical time scales of the evolution of the convective structures themselves on the visible hemisphere. 
-The LSP origin is harder to pinpoint given the insufficient time length of our time series. One can speculate if it may be related to the 
-longest living photospheric structures, but this explanation does not adequately explain all the features associated to this period. Longer time 
-series of spectropolarimetric observations of Betelgeuse will be required to shed light on this.
+Although of secondary importance, we may suggest with more or less ease what phenomena are at the origin of the change of variability before and after the great dimming.
+We speculate that the great dimming event hustled the dynamics of the photosphere, so that the largest granule are destroyed by the turbulent motions caused by the non equilibrium of the photosphere.
+Since this event, the variability of Betelgeuse has shortened and is in adequacy with the timescale of which evolve the smallest convective cells, that is 200 days.
+We expect in the future to see the 400 days variability being the dominant period, the time for the photosphere to go back to equilibrium.
+Longer time
+series of spectropolarimetric observations of Betelgeuse will be required to shed light on this
+
+%Although of secondary importance, we may suggest with more or less ease what phenomena are at the origin of the main periods. The 400-day period 
+%appears to be strongly related to the appearance of plumes in the back hemisphere of the star rising high enough to be seen above the limb.
+%Such plumes have been described in another RSG, $\mu$ Cep, and while less common or prominent, they are also present in Betelgeuse. The irregularity 
+%of their presence justifies the 400-day period, which is captured by the Lomb-Scargle periodogram in the polarization signals attributed to 
+%this phenomenon. The redshifted wavelengths at which this periodicity is found are justified by the plume moving away of the star and of the Earth while 
+%raising in the back hemisphere, while the brightness increase is justified by %the extra surface of emitting plasma visible. The 200-day period is less 
+%prominent but appears to be correlated to the typical time scales of the evolution of the convective structures themselves on the visible hemisphere. 
+%The LSP origin is harder to pinpoint given the insufficient time length of our time series. One can speculate if it may be related to the 
+%longest living photospheric structures, but this explanation does not adequately explain all the features associated to this period. Longer time 
+%series of spectropolarimetric observations of Betelgeuse will be required to shed light on this.
 

### 2024-04-17 main.tex
Add files via upload

@@
 Betelgeuse is a prototypical red supergiant (RSG) known to be a semi-regular variable. Several periods can be found in the 
-literature usually clustered around the so-called long secondary period (LSP) of 2000 days, and other shorter periods of 400 and 200 days.
+literature usually clustered around the so-called long secondary period (LSP) of 2000 days, and other shorter periods of 400 and 200 days \citep{kiss_variability_2006}.
 The origin of the LSP remains poorly known, some tried to explain the LSP by the lifetime of the large convective cells present at the surface of RSG (e.g \cite{stothers_giant_2010}) or by the magnetic field \citep{wood_long_2004}.
@@
 Other theories such as a drop in temperature was also mentioned by \cite{harper_photospheric_2020}, or an increase in the molecular opacity \citep{kravchenko_atmosphere_2021}.
-In this event it was clear that a change in brightness was not due to any kind of pulsation but to a change on the
+In this event it turns out that a change in brightness was not due to any kind of pulsation but to a change on the
 brightness distribution over the stellar disk.
@@
     \includegraphics[width=0.5\textwidth]{Lomb-Scargle Stokes Q.pdf}
-    \caption{Same as figure \ref{LS intensity} for stokes Q. }
+    \caption{Same as figure \ref{LS intensity} for stokes $Q$. }
     \label{LS Q}
@@
     \includegraphics[width=0.5\textwidth]{Lomb-Scargle Stokes U.pdf}
-    \caption{Same as figure \ref{LS intensity} for stokes U.}
+    \caption{Same as figure \ref{LS intensity} for stokes $U$.}
     \label{LS U}
@@
     \includegraphics[width=0.5\textwidth]{Lomb-Scargle linear polarization.pdf}
-    \caption{Same as figure \ref{LS intensity} for linear polarization: $\sqrt{\mathrm{Q^2}+\mathrm{U^2}}$.}
+    \caption{Same as figure \ref{LS intensity} for linear polarization: $\sqrt{Q^2+U^2}$.}
     \label{LS linear polarization}
@@
 From stokes $Q$ and $U$, we recover the 200 days period and the 330 days period is more or less close to the 400 days period found in the literature.
-Figure \ref{LS linear polarization} shows the Lomb-Scargle periodogram of the total linear polarization of Betelgeuse: $\sqrt{\mathrm{Q^2+U^2}}$. 
+Figure \ref{LS linear polarization} shows the Lomb-Scargle periodogram of the total linear polarization of Betelgeuse: $\sqrt{Q^2+U^2}$. 
 Interestingly, the periodogram shows a strong power at both 330 days and 200 days, which are the period that we have found in both stokes $Q$ and $U$.
@@
 
-Betelgeuse is observed in average every two or three weeks by the TBL, and hence
+Betelgeuse is observed in average every month by the TBL, and hence
 we are able to follow its surface activity through polarimetry imaging. This has  allowed in the past the estimation of  the size and the
  lifetime of the convective cells
-at the surface. From these images, we computed a photo-center, which is sensitive to the size and the number of convective cells. 
-An homogeneous star will have a photo-center displacement coinciding with barycenter of the star, whereas a star with one or two huge convective cells will have a
+at the surface \citep{lopez_ariste_convective_2018}. From these images, we computed a photo-center, which is sensitive to the size and the number of convective cells. 
+An homogeneous star will have a photo-center displacement coinciding with barycenter of the star, whereas a star with one or two large convective cells will have a
 photo-center displacement more important, up to a few percent of the stellar radius in the case of red supergiants \citep{chiavassa_probing_2022}.
@@
 Since the photo-center is linked to surface convection, it is worth checking for periods in its 
-dynamics over the 13 years of observations of Betelgeuse at the TBL. Such periods will not 
+dynamics over the 10 years of observations of Betelgeuse at the TBL. Such periods will not 
 be completely unrelated to the ones presented in the previous section, but will capture 
@@
 RSGs encounter mass loss event, where due to stellar winds, plasma is ejected in the interstellar medium.
-Those events have already been seen in RSG like $\mu$Cep, where \cite{lopez_ariste_height_2023} found an excess of linear polarization beyond the limb of the star and associated it to rising plumes of plasma in the back hemisphere of the star.
+Those events have already been inferred in RSG like $\mu$Cep, where \cite{lopez_ariste_height_2023} found an excess of linear polarization beyond the limb of the star and associated it to rising plumes of plasma in the back hemisphere of the star.
 In the case of Betelgeuse, this kind of event happened only once or twice during the last 10 years, with no change in the variability of the star.

### 2024-04-23 main.tex
Add files via upload

@@
 \section{Introduction}
-
 Betelgeuse is a prototypical red supergiant (RSG) known to be a semi-regular variable. Several periods can be found in the 
@@
 
-Besides radial pulsation, surface activity leading to random brightness variation was also mentioned to explain the variability of Betelgeuse, being firstly theorised by \cite{schwarzschild_scale_1975}.
+Besides radial pulsation, surface activity leading to random brightness variation was also mentioned to explain the variability of Betelgeuse, being firstly theorised by \cite{schwarzschild_scale_1975} and \cite{gray_mass_2008} proposed that the 400 days period was a consequence of the convective activity.
 Betelgeuse is known to present large convective cells with lifetimes in the order of one to two years \citep{lopez_ariste_convective_2018} and measurable changes
@@
 variability could be expected to be random, but will present quasi-periodicities related to the typical time scales of the 
-convective patterns. \\
+convective patterns \citep{gray_mass_2008}. \\
 
@@
 This typical timescale has also been observed in the numerical simulations and might play an important role in the determination of stellar distances through the displacement of the photo-center \citep{chiavassa_probing_2022}.
-Once again, the 330 days periodicity being more present in stokes $U$ than $Q$ unravels a stochastic process that is due to convective motions.
+Once again, the 330 days periodicity being more present in stokes $U$ than $Q$ unravels a stochastic process that is due to convective motions, and is consistent with the work of \cite{gray_mass_2008}, who attributed the 400 days period to the large convective cells.
 \
@@
 in the light curve and in the spectropolarimetric observations.
-This is the main conclusion of this work: since we are able to recover the different periods in the linear polarization profiles, we associate the 200 days periodicyt to convective dynamics while the 400 days periodicity is associated to both stochastic pulsations and convective timescales of the largest granules.%the observed periods are related
+This is the main conclusion of this work: since we are able to recover the different periods in the linear polarization profiles, we associate the 200 days periodicyt to convective dynamics while the 400 days periodicity is associated to both stochastic pulsations and convective timescales of the largest granules.
+This conclusion is in agreement with the scenario proposed by \cite{gray_mass_2008}, who attributed the variability of Betelgeuse to surface convection.
+%the observed periods are related 
 %to convective dynamics and not to radial pulsations of the star.

### 2024-05-06 main.tex
Intro proofed

@@
 \section{Introduction}
-Betelgeuse is a prototypical red supergiant (RSG) known to be a semi-regular variable. Several periods can be found in the 
-literature usually clustered around the so-called long secondary period (LSP) of 2000 days, and other shorter periods of 400 and 200 days \citep{kiss_variability_2006}.
-The origin of the LSP remains poorly known, some tried to explain the LSP by the lifetime of the large convective cells present at the surface of RSG (e.g \cite{stothers_giant_2010}) or by the magnetic field \citep{wood_long_2004}.
-The shorter periods of 400 and 200 days seem to have a different origin than the LSP.
-By studying a large sample of RSG, \cite{kiss_variability_2006} attributed the 400 days period to radial pulsation modes, the 200 days period being attributed to the first overtone \citep{joyce_standing_2020}, which was in agreement with the historical work of \cite{stothers_pulsation_1969}.\\
+Betelgeuse is a prototypical red supergiant (RSG) known to be a semi-regular variable. Several periods for this variability can be found in the 
+literature, usually clustered around the so-called long secondary period (LSP) of 2000 days, plus shorter periods of 400 and 200 days \citep{kiss_variability_2006}. 
+The origin of the LSP remains poorly known: some have tried to find the explanation of  the LSP in the lifetime of the large convective cells present at the surface of RSG (e.g \cite{stothers_giant_2010}) while others have invoked a the magnetic field \citep{wood_long_2004}.
+The shorter periods of 400 and 200 days seem to have a different origin than the LSP. By studying a large sample of 
+RSG, \cite{kiss_variability_2006} attributed the 400 days period to radial pulsation modes, while the 200 days period was attributed to the first overtone \citep{joyce_standing_2020}, 
+ in agreement with the historical work of \cite{stothers_pulsation_1969}.\\
 
-Besides radial pulsation, surface activity leading to random brightness variation was also mentioned to explain the variability of Betelgeuse, being firstly theorised by \cite{schwarzschild_scale_1975} and \cite{gray_mass_2008} proposed that the 400 days period was a consequence of the convective activity.
+Besides radial pulsation, surface activity leading to random brightness variation has also been mentioned to explain the variability of Betelgeuse.
+\textbf{Having been firstly theorised by \cite{schwarzschild_scale_1975} and \cite{gray_mass_2008}
+proposed that the 400 days period was a consequence of the convective activity}.
 Betelgeuse is known to present large convective cells with lifetimes in the order of one to two years \citep{lopez_ariste_convective_2018} and measurable changes
@@
 
-At the end of 2019, Betelgeuse reached a historical minimum in its luminosity, called the great dimming \citep{guinan_fall_2020}. 
-From interferometric data, it is speculated that this event was caused by a cloud of dust close to the line of sight \citep{montarges_dusty_2021}. 
-Interferometric images show a drop in luminosity in the southern hemisphere of Betelgeuse, that might be caused by a mass loss event and leading to this dimming \citep{dupree_great_2022}.
-Other theories such as a drop in temperature was also mentioned by \cite{harper_photospheric_2020}, or an increase in the molecular opacity \citep{kravchenko_atmosphere_2021}.
-In this event it turns out that a change in brightness was not due to any kind of pulsation but to a change on the
+At the end of 2019, Betelgeuse reached a historical minimum in its luminosity, called the Great Dimming \citep{guinan_fall_2020}. 
+From interferometric and spectroscopic data, it has been determined that this event was caused by a cloud of dust formed close to the line of 
+sight \citep{montarges_dusty_2021}. 
+Interferometric images show a drop in luminosity in the southern hemisphere of Betelgeuse, that might be caused by a 
+mass loss event and leading to this dimming \citep{dupree_great_2022}.
+Other hypothesis have  also been explored , such as a drop in temperature \citep{harper_photospheric_2020}, or an increase in the molecular opacity \citep{kravchenko_atmosphere_2021}.
+This event is an extreme illustration of  changes in brightness that were not due to any kind of pulsation but to a change on the 
 brightness distribution over the stellar disk.
 %due to the presence of large quantities of dust.
-Since the end of the Great Dimming event, Betelgeuse has continue its random variation of brightness. 
-Interestingly, 
-it has been shown by \cite{jadlovsky_analysis_2023} and \cite{dupree_great_2022} that the periodicity of Betelgeuse has changed after the dimming. Using the light curves from AAVSO, 
-the authors showed that before the dimming, the dominant period of Betelgeuse was the 400 days period, while after the great dimming the main periods of Betelgeuse have shortened, oscillating between 97 days and 230 days \citep{dupree_great_2022}, revealing a change in the behaviour of the atmosphere of Betelgeuse.
+Since the end of the Great Dimming event, Betelgeuse has continued its random variation of brightness. 
+But quite interestingly, 
+it has been shown by \cite{jadlovsky_analysis_2023} and \cite{dupree_great_2022} that the periodicity of Betelgeuse has changed since the dimming. Using the light curves from AAVSO, 
+the cited authors showed that before the dimming, the dominant period of Betelgeuse was the 400 days period, while after the great dimming the main periods of Betelgeuse have shortened, oscillating between 97 days and 230 days \citep{dupree_great_2022}, revealing a change in the behaviour of the atmosphere of Betelgeuse.
 While the great dimming may be seen as 
-a singular event, it brings up the question whether all changes in brightness cannot be due to similar changes in the brightness 
-distribution over the disk, unrelated to any pulsation phenomenon. 
-Our interest in an alternative explanation of the variability
-in terms  of convective patterns arose. In order to address this question we shall seek for such typical periods of 400 and 200 days in  
-observational proxies related to the convective activity.\\
+a singular event, such modifications of the variability periods  bring up the question whether all changes in brightness cannot be due to similar changes 
+in the brightness  distribution over the disk, and unrelated to any pulsation phenomenon. 
+%Our interest in an alternative explanation of the variability
+%in terms  of convective patterns arose. 
+In order to address this question we  seek for the typical periods of 400 and 200 days in  
+observational proxies which are related to the convective activity but not to any pulsation, such as the linear polarization spectra.\\
 %but unrelated to pulsations and which cannot be directly linked to brightness.
@@
 
-Linear polarization in the atomic lines of the spectra of Betelgeuse, discovered by \cite{auriere_discovery_2016}, has been interpreted as the joint action of 
-two mechanisms. First the depolarization of the continuum by atoms which absorb linearly polarized light and re-emit unpolarized light, the 
+Linear polarization in the atomic lines of the spectrum of Betelgeuse, discovered by \cite{auriere_discovery_2016}, has been interpreted as the joint action of 
+two mechanisms. First the depolarization of the continuum by atoms which absorb linearly polarized light from the continuum and re-emit unpolarized light, the 
 continuum photons being polarized by Rayleigh scattering. This depolarization produces signals with an azimuth symmetry over the disk that 
-would  cancel out the net linear polarization on a homogeneous disk. Thus it must be combined with an inhomogenous disk to produce the net linear polarization 
+would  cancel out the net linear polarization on a homogeneous disk.
+Thus, such mechanism of production of polarization must be combined with an inhomogenous disk to produce the net linear polarization
 signal observed. This interpretation suggested the possibility of mapping those brightness inhomogeneities. This has been done by \cite{lopez_ariste_convective_2018} 
-even producing 3-dimensional images of the atmosphere of Betelgeuse \citep{lopez_ariste_three-dimensional_2022} when taking advantage of the different height of formation of 
+even producing 3-dimensional images of the atmosphere of Betelgeuse \citep{lopez_ariste_three-dimensional_2022} when taking advantage of the different heights of formation of 
 different lines in the spectrum of Betelgeuse. The images produced compare very well with contemporaneous images made with interferometric 
-techniques \citep{montarges_close_2016} and show clear convective patterns, akin to solar granulation.\
+techniques \citep{montarges_close_2016} and show clear convective patterns, akin to solar granulation.
 
 Linear polarization observed in the atomic lines is therefore a proxy of convection, unrelated to radial 
-pulsations but linked to the brightness inhomogeneities due to the convective patterns in Betelgeuse. If one can find the aforementioned periods 
-in linear polarization, we must conclude that these periods are related to the convective activity which originates those linear polarization signals.
+pulsations but linked to the brightness inhomogeneities due to the convective patterns in Betelgeuse. If one can find the 
+aforementioned variability periods 
+in linear polarization, we must conclude that these periods are related to the convective activity which originates the linear 
+polarization signals.
 %and unrelated to any radial pulsation phenomena. 
 This is the purpose of the present work.
-In section \ref{Section 2} we describe the dataset of linear polarization obtained with the TBL and the Lomb-Scargle periodogram used to derive periods.
+
+In section \ref{Section 2} we describe the dataset of linear polarization obtained with the Telescope Bernard Lyot (TBL) and the Lomb-Scargle technique used to derive periods. 
 In section \ref{section 3} we associate the 200 days period to convective timescale of the smallest granules and the 330 days periodicity to the largest granules, and we provide an explanation in the change of variability of Betelgeuse before and after the great dimming.

### 2024-05-06 main.tex
Revision

@@
 
-In section \ref{Section 2} we describe the dataset of linear polarization obtained with the Telescope Bernard Lyot (TBL) and the Lomb-Scargle technique used to derive periods. 
-In section \ref{section 3} we associate the 200 days period to convective timescale of the smallest granules and the 330 days periodicity to the largest granules, and we provide an explanation in the change of variability of Betelgeuse before and after the great dimming.
-In section \ref{Conclusion}, we summup our work.
+%In section \ref{Section 2} we describe the dataset of linear polarization obtained with the Telescope Bernard Lyot (TBL) and the 
+%Lomb-Scargle technique used to derive periods. 
+%In section \ref{section 3} we associate the 200 days period to convective timescale of the smallest granules and the 330 days periodicity to the largest granules, and we provide an explanation in the change of variability of Betelgeuse before and after the great dimming.
+%In section \ref{Conclusion}, we summup our work.
 
@@
 
-Betelgeuse has been observed for the last 10 years with Narval and Neo-Narval at the Telescope Bernard Lyot (TBL). These two instruments measure the polarization 
+Betelgeuse has been observed for the last 10 years with Narval and Neo-Narval at the Telescope Bernard Lyot. These two instruments measure the polarization 
 over the visible spectra of Betelgeuese (390-1000 nm) with high spectral resolution (R=65000) and high polarimetric sensitivity. Despite 
-this high sensitivity, the signal-to-noise ratios are not sufficient to measure the weak polarization signals in individual atomic lines of the 
-spectrum. These amplitudes are today known to be of the order of $10^{-4}$ times the continuum intensity, and only exceptionally do they 
+this high sensitivity, the signal-to-noise ratios per spectral bin are not sufficient to measure the weak polarization signals in individual 
+atomic lines. These amplitudes are today known to be of the order of $10^{-4}$ times the continuum intensity, and only exceptionally do they 
 reach amplitudes of $10^{-3}$ the continuum. In these exceptional observations, enough photons can be accumulated per spectral bin to 
-see the linear polarization signal above noise \citep{auriere_discovery_2016}. Most commonly, the amplitudes of linear polarization are below noise levels. In those 
-occasions we have to add up the signals of thousands of lines to reduce the noise and increase the signal-to-noise ratios. This line addition 
+see the linear polarization signal above noise \citep{auriere_discovery_2016}. But most commonly, the amplitudes of linear polarization are 
+below noise levels and measuring them requires the addition of the signals of thousands of lines to reduce  noise and increase the 
+signal-to-noise ratios. This line addition 
 is done through a technique called Least-Squares Deconvolution (LSD) \citep{donati_spectropolarimetric_1997} and assumes that the 
@@
 
-In this work we shall look into the signals produced through these line-addition techniques, and which produce a pseudo-spectral line in intensity 
+In this work we shall look into the signals produced through these line-addition techniques, and which produce a pseudo-line in intensity 
 and linear polarization which does not belong to any atomic species in particular but to an average of all of them present and emitting in the 
@@
 \cite{lopez_ariste_three-dimensional_2022}, which also give details on observing times and conditions as well as more detailed description 
-on the data reduction. \footnote[1]{Beyond a 2-year proprietary embargo, and up to technical issues, all these data is available through 
-PolarBase (http://polarbase.irap.omp.eu/).}
+on the data reduction\footnote[1]{Beyond a 2-year proprietary embargo, and up to technical issues, all these data is available through 
+PolarBase (http://polarbase.irap.omp.eu/).}.
 
@@
 
-\subsection{The variability present in the stokes parameters}
+\subsection{The variability in the Stokes profiles}
 
 Trying to find periods in an astrophysical context carries the hurdle of unevenly spaced observed data. To overcome this issue, 
-we used the Lomb-Scargle periodogram \citep{lomb_least-squares_1976,scargle_studies_1982}. For a given set of data, it fits sinusoidal 
-using least-squares at each frequency between the first and the last observation. The better the fit, the higher the power attributed to the frequency. 
+we used  Lomb-Scargle periodograms \citep{lomb_least-squares_1976,scargle_studies_1982}. For a given set of data, one fits sinusoids to the data
+using least-squares, with frequencies sampled between the first and the last observation. The better the fit, the higher the power attributed to a frequency. 
 We applied this technique to the  LSD profiles of Stokes $I$, $U$, $Q$ and to the total linear polarization of Betelgeuse observed by the TBL during 
@@
 We first computed the Lomb-Scargle periodogram at each wavelength of the LSD profile, between -25 km/s and +60 km/s to span the entire 
-signal present in the LSD profile. \cite{lopez_ariste_convective_2018} found the most blueshifted signal towards -20 km/s, 
-and interpreted it as  the maximum velocity of the rising plasma, seen from the Earth. The most redshifted signal was found at +40 km/s. 
+signal present in the LSD profile. \cite{lopez_ariste_convective_2018} found the most blueshifted signal in linear polarization  towards -20 km/s, 
+and interpreted it as  the maximum velocity of the rising plasma, seen from  Earth. The most redshifted signal was often found at +40 km/s. 
 This velocity is interpreted
 as the rest velocity of the star in our reference frame. Seldom, signals at velocities larger than +40 km/s can be found 
-and can be interpreted as either dark and cold plasma falling back to Betelgeuse, 
+and are interpreted as either dark and cold plasma falling back to Betelgeuse, 
 or to rising plumes of plasma located in the hidden face of Betelgeuse,
@@
 
-Figures \ref{LS intensity},\ref{LS Q},\ref{LS U} and \ref{LS linear polarization} show the Lomb-Scargle periodogram of the LSD profile of respectively 
-the intensity, stokes $Q$, stokes $U$ and the linear polarization. 
+Figures \ref{LS intensity},\ref{LS Q},\ref{LS U} and \ref{LS linear polarization} show the Lomb-Scargle periodogram of the LSD profiles of respectively 
+the intensity, Stokes $Q$, Stokes $U$ and the total linear polarization. 
 %The abscissa is the frequency (in $\mathrm{days^{-1}}$), while the ordinate is the heliocentric radial velocity in km/s. 
-For each figure, the uper panel is the Lomb-Scargle periodogram of each velocity bin (black line) and the total average (red line).
-The green line corresponds to the window function.
-The lower panel is the Lomb-Scargle periodogram represented for each velocity bin, while the white dotted line refer to 400 and 200 days.
-The right panel is the mean profile of each stokes parameter for each velocity bin.
+For each figure, the uper panel is the Lomb-Scargle periodogram of each velocity bin (black line) and the total average (red line). The green line corresponds to the window function. 
+The lower panel is the Lomb-Scargle periodogram represented for each velocity bin, while the white dotted lines mark the 400 and 200 days periods for 
+reference. The right panel shows the mean profile of each Stokes parameter at each velocity bin.
 
@@
     The upper panel is the Lomb-Scargle periodogram for each velocity bin (black line) and the average (red line). The green line is the window function.
-    In the lower pannel, the two white dashed lines represent respectively the 400 days and the 200 days periods. The right panel is the average intensity profile for each velocity bin.}
+    In the lower pannel, the two white dashed lines mark respectively the 400 days and the 200 days periods. The right panel is the average intensity profile for each velocity bin.}
     \label{LS intensity}
@@
     \includegraphics[width=0.5\textwidth]{Lomb-Scargle Stokes Q.pdf}
-    \caption{Same as figure \ref{LS intensity} for stokes $Q$. }
+    \caption{Same as figure \ref{LS intensity} for Stokes $Q$. }
     \label{LS Q}
@@
     \includegraphics[width=0.5\textwidth]{Lomb-Scargle Stokes U.pdf}
-    \caption{Same as figure \ref{LS intensity} for stokes $U$.}
+    \caption{Same as figure \ref{LS intensity} for Stokes $U$.}
     \label{LS U}
@@
     \includegraphics[width=0.5\textwidth]{Lomb-Scargle linear polarization.pdf}
-    \caption{Same as figure \ref{LS intensity} for linear polarization: $\sqrt{Q^2+U^2}$.}
+    \caption{Same as figure \ref{LS intensity} for total linear polarization: $\sqrt{Q^2+U^2}$.}
     \label{LS linear polarization}
@@
 %\subsection{Intensity}
-Figure \ref{LS intensity} shows the Lomb-Scargle periodogram of the LSD of the intensity profile.
-The 400 and 200 days period seem to be recovered by the periodogram.
-However, the 400 days period is uncomfortably close to the peak of the window function, while the 200 days period is not.
-We also notice that the two periods are present at the same HRV.
-Interestingly, \cite{mathias_evolution_2018} also found this 200 days periodicity with a shorter observation period.
+Looking into Fig. \ref{LS intensity}, which shows the Lomb-Scargle periodogram of the LSD of the intensity profile, 
+we notice that the 400 and 200 days periods seem to be recovered by the periodogram. However, the 400 days period is uncomfortably close to 
+the peak of the window function. We also notice that the two periods are present at the same HRV. Interestingly, 
+\cite{mathias_evolution_2018} already found this 200 days periodicity in spectroscopic observations, in spite of a shorter observation period.
 
-Figure \ref{LS Q} and \ref{LS U} show the Lomb-Scargle periodogram of both stokes $Q$ and $U$. It is clear that the periodogram is completely different 
+Figures \ref{LS Q} and \ref{LS U} show the Lomb-Scargle periodograms of Stokes $Q$ and $U$. These periodograms are completely different 
 from the intensity one.
@@
 \ref{LS Q} around 200 days.
-Concerning stokes U in fig.
-\ref{LS U}, it appears that the main period is around $0.003 \ \mathrm{days^{-1}}$ (330 days).
-Other period such as the one at 200 days or 250 days ($0.004 \ \mathrm{days^{-1}}$) are present but difficult to trust.
-From stokes $Q$ and $U$, we recover the 200 days period and the 330 days period is more or less close to the 400 days period found in the literature.
-Figure \ref{LS linear polarization} shows the Lomb-Scargle periodogram of the total linear polarization of Betelgeuse: $\sqrt{Q^2+U^2}$. 
-Interestingly, the periodogram shows a strong power at both 330 days and 200 days, which are the period that we have found in both stokes $Q$ and $U$.
+Concerning Stokes U, Fig.
+\ref{LS U}, it appears on the other hand that the main period is around $0.003 \ \mathrm{days^{-1}}$ (330 days).
+Other periods such as those at 200 days or 250 days ($0.004 \ \mathrm{days^{-1}}$) are present but difficult to trust.
+From both Stokes $Q$ and $U$, we recover the 200 days period, and also a 330 days period which  is more or less close to the 400 days period found in the literature.
+Figure \ref{LS linear polarization} shows the Lomb-Scargle periodogram of $\sqrt{Q^2+U^2}$, the total linear polarization of Betelgeuse. 
+The periodogram recovers the strong powers at both 330 days and 200 days, already found in either Stokes $Q$ and $U$.
 
@@
 
-\subsection{The variability of Betelgeuse inferred from polarimetry imaging}
+\subsection{The variability in polarimetric imaging}
 
-Using linear polarization, \cite{lopez_ariste_convective_2018} have been able to reconstruct images of Betelgeuse that have been successfully compared 
+Using linear polarization, \cite{lopez_ariste_convective_2018} were able to reconstruct images of Betelgeuse that have been successfully compared 
 to inteferometric images of \cite{montarges_close_2016}. The images are produced by finding the brightness distribution that better fits the observed linear 
-polarization LSD profile using a Marquardt-Levenberg minimisation.
+polarization LSD profiles, using a Marquardt-Levenberg minimisation.
 
 Betelgeuse is observed in average every month by the TBL, and hence
-we are able to follow its surface activity through polarimetry imaging. This has  allowed in the past the estimation of  the size and the
- lifetime of the convective cells
-at the surface \citep{lopez_ariste_convective_2018}. From these images, we computed a photo-center, which is sensitive to the size and the number of convective cells. 
-An homogeneous star will have a photo-center displacement coinciding with barycenter of the star, whereas a star with one or two large convective cells will have a
-photo-center displacement more important, up to a few percent of the stellar radius in the case of red supergiants \citep{chiavassa_probing_2022}.
+we are able to follow its surface activity through this technique of polarimetric imaging. This has  allowed in the past the estimation of  
+the size and the lifetime of the convective cells
+at the surface \citep{lopez_ariste_convective_2018}. From these images, we can also compute a photo-center, which is sensitive to the size and the number of convective cells. 
+An homogeneous star will have a photo-center displacement coinciding with the barycenter of the star, whereas a star with one or two large convective cells will have a
+photo-center displacement more important, up to a few percent of the stellar radius in the case of \textbf{red giants?} \citep{chiavassa_probing_2022}.
 
@@
 %We computed the Lomb-Scargle periodogram of the displacement of the photo-center.
-But, before going any further, we have to warn
- about the interpretation of such images. 
-Linear polarization suffers from a $180 ^\circ$ degree ambiguity already mentioned in \cite{auriere_discovery_2016}. For that reason, our images 
-can be 
-rotated by $180^\circ$ degree, and the brightness distribution will still fit the observed LSD polarization profiles. 
+But, before proceeding to search for periods in the movements of the photo-center, we have to warn about the interpretation of such images. 
+Linear polarization suffers from a $180 ^\circ$ degree ambiguity already mentioned in \cite{auriere_discovery_2016}. For that reason, our 
+images can be rotated by $180^\circ$ degrees, and the brightness distribution will still fit the observed LSD polarization profiles. 
 If a particular observation were to be treated independently of previous observations, the solution found by 
 the algorithm could be any of the ambiguous solutions possible. To obtain a continuity 
-between the series of images, 
-for a given day we used the brightness distribution of the previous day as initial point 
-of the  fitting iteration; and for the first image, we start with a 
-random brightness distribution as the beginning of the fit. 
+between the time series of images, for a given day we use the brightness distribution of the previous day as initial point 
+of the  fitting iteration. The first image of the series  starts its minimization iteration with 
+random brightness distribution. 
 Even though we are able to produce a series of images  consistent with each other, it is important 
 to keep in mind that the photo-center displacement computed from the series 
-will be affected by the choice of the first image. 
+will be affected by that first image of the series. 
 To overcome this issue, we decided to compute
-the photo-center displacement from 100 different initial parameters and their subsequent series 
-of images in an effort to recover from the ensemble
+the photo-center displacement from 100 different time serieas, each one starting from a different initial image, in an effort to recover 
+from such ensemble
 properties that are independent of the choice of the first image.
@@
     \includegraphics[width=0.5\textwidth]{Lomb-Scargle Photo-center.pdf}
-    \caption{Lomb-Scargle periodogram of the 100 photo-center displacement of Betelgeuse. Each blue lines correspond to a Lomb-Scargle periodogram of one photo-center
-     displacement. The black line is the average of the 100 periodogram. The red dashed lines represent the 400 and 200 days period respectively 
+    \caption{Lomb-Scargle periodogram of the 100 series of photo-center displacements of Betelgeuse. Each blue line correspond to a Lomb-Scargle periodogram of one series of photo-center
+     displacement. The black line is the average of the 100 periodograms. The red dashed lines mark the 400 and 200 days periods respectively 
      (from left to right). As before, the green line is the window function.}
@@
 %As before, the upper panel shows the Lomb-Scargle periodogram from data before the great dimming only, whereas we used the whole set of data for the lower panel.
-while the black dashed lines mark the 400 and 
-200 days periods respectively.
-As before, the window function is represented by the green line.
-Interestingly, we find two bumps around the 400 days period.
-We cannot directly confirm that we recover the 400 days period from those small bumps, but since polarimetry imaging involves only surface convection, this is an argument that points toward a variability explained by surface convection.
+while the red dashed lines mark the 400 and 
+200 days periods respectively. As before, the window function is represented by the green line. We find two bumps around the 400 days period, which are not 
+sufficiently significant to confirm the presence of such period on their own. However, they concur with the results of the previous section, 
+and since polarimetry imaging involves only surface convection, this is a further argument towards a variability explained by surface convection alone. 
 %First, we find that the Lomb-Scargle power is high around the LSP, though
@@
 
-\section{The variability of the light curve of Betelgeuse}
+\section{Explanations for the variability of the light curve of Betelgeuse}
 \label{section 3}
@@
 The presence of the same 
-periods and, after taking into consideration the time span and sparsity of the TBL data series, the larger amplitude of the peaks in
-polarization data  than in the light curve point to a common origin, one that is rather based upon convective dynamics.
-Since we associate stokes $Q$ and $U$ to proxies of convection, we associate the 200 days period to the typical timescales on which the Stokes parameters evolve.
-The LSD profiles of stokes $Q$ or $U$ will slightly change when we observe Betelgeuse every month, but are completely overhauled 
-on time scales of several months. The Lomb-Scargle periodograms appear to capture this spectral dynamic around periods of 200 days.
+periods and, after taking into consideration the time span and sparsity of the TBL data series, the larger amplitude of the peaks found in
+polarization data  compared to those  in the light curve point to a common origin, one that is rather based upon convective dynamics.
+
+\textbf{The 200-days period appears prominently in the Stokes Q and U profiles. We can }
+associate the 200-days period to the typical timescales on which the Stokes parameters evolve. 
+The LSD profiles of Stokes $Q$ or $U$ will slightly change when we observe Betelgeuse every month, but are completely overhauled 
+on time scales of several months. The Lomb-Scargle periodograms appear to capture these dynamics at periods of about 200 days.
  Such changes in the profiles are directly linked to the dynamics of the surface,  to  the evolution and movements 
 of convective cells over the visible hemisphere. So the 200-day period appears to be related to the evolution of the convective patterns 
-on the surface of Betelgeuse.  The fact that this peak is strongly present in stokes $Q$ but less in $U$ points toward a more stochastic origin, which can fully be explained by convection.\
+on the surface of Betelgeuse.  The fact that this peak is strongly present in Stokes $Q$ but less in $U$ 
+can be interpreted as a temporal situation of how the convective patterns have been distributed in the last 10 years.
+%points toward a more stochastic origin, which can fully be explained by convection.\
 
-The other period found in the linear polarization and stokes $U$, the 330 days periods, is more or less close to the 400 days period often present in the literature (e.g \cite{kiss_variability_2006} found a period $388 \pm 30$ days).
-It has been shown by \cite{lopez_ariste_convective_2018} that large convective cells can live up to one year, thus this period would be associated to the timescale on which the largest granules evolve.
-This typical timescale has also been observed in the numerical simulations and might play an important role in the determination of stellar distances through the displacement of the photo-center \citep{chiavassa_probing_2022}.
-Once again, the 330 days periodicity being more present in stokes $U$ than $Q$ unravels a stochastic process that is due to convective motions, and is consistent with the work of \cite{gray_mass_2008}, who attributed the 400 days period to the large convective cells.
-\
+The other period found in the total linear polarization and Stokes $U$, the 330-days period, is more or less close to the 400 days period 
+often present in the literature (e.g \cite{kiss_variability_2006} found a period $388 \pm 30$ days). It has been found
+by \cite{lopez_ariste_convective_2018} that large convective cells can live  one  to two years, and this period could be associated to the 
+timescales on which the largest granules evolve. This typical timescale has also been observed in the numerical simulations and might 
+play an important role in the determination of stellar distances through the displacement of the photo-center \citep{chiavassa_probing_2022}. 
+Once again we observe this 330-days periodicity to be more present in Stokes $U$ than in $Q$, what may unravel a 
+random situation due to convective motions in the last 10 years. This interpretation appears to be  consistent with the work of \cite{gray_mass_2008}, 
+who attributed the 400 days period to the large convective cells.
 
-Finally, the intensity profile seems to match the 400 days periods, but is uncomfortably close to a peak of the window function of the TBL and a secondary peak close to 200 days.
-Since we are able to recover the 200 days period in the stokes paramters, we attribute this period to the convective timescales, but concerning the 400 days period, it could be a mix of the typical convective timescales of large granules associated with stochastic pulsations exited by convective cavities.
-Up to this point, the model developed to interpret linear polarization does not involve pulsations, that is why we interpret the 200 days period as the convective turnover timescale.
-The 400 days period involve the intensity profile, thus the 400 days period is probably caused by both convective timescales and stochastic pulsations.
+\textbf{Finally, the intensity profiles do  seem to vary in  400 days periods, but the measured peak is uncomfortably close to a 
+peak of the window function of the time series observed with the TBL. This 400-days period, it could be a mix of the typical convective 
+timescales of large granules associated with stochastic pulsations exited by convective cavities. Up to this point, the model developed to 
+interpret linear polarization does not involve pulsations, that is why we interpret the 200 days period as the convective turnover timescale. 
+The 400 days period involve the intensity profile, thus the 400 days period is probably caused by both convective timescales and stochastic 
+pulsations.}
  %As showed by
@@
 
-After examining the periods found by the Lomb-Scargle technique on the polarization data, and its derivatives, obtained by the TBL over the last 
-10 years, it is worth putting them in the context of the periods found traditionally on light curves over the same period of time. Two aspects 
-are of our interest in this comparison: the behaviour of the light curve before and after the great dimming.
+%After examining the periods found by the Lomb-Scargle technique on the polarization data, and its derivatives, obtained by the TBL over the last 
+%10 years, it is worth putting them in the context of the periods found traditionally on light curves over the same period of time. Two aspects 
+%are of  interest in this comparison: the behavior of the light curve before and after the great dimming.
 %and the amplitudes of the peaks 
-%at the main cited periods.
+%at the main cited per%iods.
 
@@
 
-The main observation  presented in this work is that the same periods identified in the light curve of Betelgeuse are also seen in the intensity and polarization
-spectra measured with the TBL. Periods in the light curve have often been traditionally associated to pulsation due to pressure modes.
+The main result  presented in this work is that the same periods identified in the light curve of Betelgeuse are also seen in the intensity and polarization
+spectra, as measured with the TBL. Periods in the light curve have often been traditionally associated to pulsation due to pressure modes.
 %But such pulsation would in no case affect the polarisation signals nor would explain the fact that the periods can be observed with stronger 
@@
 %And such origin cannot be pulsations.
-The interpretation of the linear polarization profiles is made in terms of convective structures in the atmosphere of Betelgeuse. We suggest 
+But the interpretation of the linear polarization profiles is made in terms of convective structures in the atmosphere of Betelgeuse. We suggest 
 that it is in these convective structures and their temporal evolution that we will find the true reason for the observed periodicities both 
-in the light curve and in the spectropolarimetric observations.
-This is the main conclusion of this work: since we are able to recover the different periods in the linear polarization profiles, we associate the 200 days periodicyt to convective dynamics while the 400 days periodicity is associated to both stochastic pulsations and convective timescales of the largest granules.
-This conclusion is in agreement with the scenario proposed by \cite{gray_mass_2008}, who attributed the variability of Betelgeuse to surface convection.
+in the light curve and in the spectropolarimetric observations. This is the main conclusion of this work: since we are able to recover the different periods in the linear polarization profiles, 
+we associate the 200-days periodicity to convective dynamics while the 400-days periodicity may be related \textbf{to both stochastic pulsations and convective timescales of the largest granules.} 
+This conclusion concurs with the scenario proposed by \cite{gray_mass_2008}, who attributed the variability of Betelgeuse to surface convection. 
 %the observed periods are related 
@@
 
-Although of secondary importance, we may suggest with more or less ease what phenomena are at the origin of the change of variability before and after the great dimming.
-We speculate that the great dimming event hustled the dynamics of the photosphere, so that the largest granule are destroyed by the turbulent motions caused by the non equilibrium of the photosphere.
-Since this event, the variability of Betelgeuse has shortened and is in adequacy with the timescale of which evolve the smallest convective cells, that is 200 days.
-We expect in the future to see the 400 days variability being the dominant period, the time for the photosphere to go back to equilibrium.
-Longer time
-series of spectropolarimetric observations of Betelgeuse will be required to shed light on this
+Although of secondary importance, we may suggest with more or less ease what phenomena are at the origin of the change of variability before and after the 
+great dimming. We speculate that the great dimming event hustled the dynamics of the photosphere, and the largest granules were 
+destroyed by the turbulent motions caused by the non equilibrium of the photosphere. Since this event, the variability of Betelgeuse has 
+quickened  in adequacy with the timescales of  smaller convective cells, about 200-days. Were this be true, we shall expect in the 
+coming future to see the 400-days variability re-appear as the dominant period, whenever  the photosphere returns to some equilibrium. 
+Longer time  series of spectropolarimetric observations of Betelgeuse will be required to shed light on this.
 

### 2024-05-15 main.tex
Add files via upload

--- 2024-05-06 main.tex
@@
 \section{Introduction}
-Betelgeuse is a prototypical red supergiant (RSG) known to be a semi-regular variable. Several periods for this variability can be found in the 
-literature, usually clustered around the so-called long secondary period (LSP) of 2000 days, plus shorter periods of 400 and 200 days \citep{kiss_variability_2006}. 
-The origin of the LSP remains poorly known: some have tried to find the explanation of  the LSP in the lifetime of the large convective cells present at the surface of RSG (e.g \cite{stothers_giant_2010}) while others have invoked a the magnetic field \citep{wood_long_2004}.
-The shorter periods of 400 and 200 days seem to have a different origin than the LSP. By studying a large sample of 
-RSG, \cite{kiss_variability_2006} attributed the 400 days period to radial pulsation modes, while the 200 days period was attributed to the first overtone \citep{joyce_standing_2020}, 
- in agreement with the historical work of \cite{stothers_pulsation_1969}.\\
+Betelgeuse is a prototypical red supergiant (RSG), known to be a semi-regular variable. Several periods can be found in the 
+literature, usually clustered around the so-called long secondary period (LSP) of 2000 d, plus shorter periods around 400 and 200 d \citep{kiss_variability_2006}.
+The origin of the LSP remains poorly understood: some have tried to find the explanation of the LSP in the lifetime of the large convective cells present at the surface of RSG \citep[e.g][]{stothers_giant_2010} while others have invoked a magnetic field \citep{wood_long_2004}.
+The shorter periods of 400 and 200 d seem to have a different origin than the LSP.
+By studying a large sample of RSGs, \cite{kiss_variability_2006} attributed the 400 d period to radial pulsation modes, while the 200 d period being attributed to the first overtone \citep{joyce_standing_2020}, in agreement with the historical work of \cite{stothers_pulsation_1969}.\\
 
-Besides radial pulsation, surface activity leading to random brightness variation has also been mentioned to explain the variability of Betelgeuse.
-\textbf{Having been firstly theorised by \cite{schwarzschild_scale_1975} and \cite{gray_mass_2008}
-proposed that the 400 days period was a consequence of the convective activity}.
-Betelgeuse is known to present large convective cells with lifetimes in the order of one to two years \citep{lopez_ariste_convective_2018} and measurable changes
-in the span of one week. Bright convection cells near disk center will increase the integrated brightness of the star when 
-compared with other situations where such cells are found near the edges. So even without calling for the formation of 
-dust or other changes in opacity, one may argue that the simple evolution of convective patterns of the star may create a variability. Such 
+Besides radial pulsation, surface activity leading to random brightness variation has also been mentioned to explain the variability of Betelgeuse, firstly theorised by \cite{schwarzschild_scale_1975}.
+More recently, \cite{gray_mass_2008} proposed that the 400 d period was a consequence of convective activity.
+Betelgeuse is known to present large convective cells with lifetimes of the order of one to two years \citep{lopez_ariste_convective_2018} and measurable changes
+within the span of one week.
+Bright convection cells near the disk center will increase the integrated brightness of the star compared with other situations where such cells are found near the edges.
+So, even without invoking the formation of
+dust or other changes in opacity, one may argue that the simple evolution of convective patterns of the star may create variability. Such 
 variability could be expected to be random, but will present quasi-periodicities related to the typical time scales of the 
@@
 At the end of 2019, Betelgeuse reached a historical minimum in its luminosity, called the Great Dimming \citep{guinan_fall_2020}. 
-From interferometric and spectroscopic data, it has been determined that this event was caused by a cloud of dust formed close to the line of 
-sight \citep{montarges_dusty_2021}. 
-Interferometric images show a drop in luminosity in the southern hemisphere of Betelgeuse, that might be caused by a 
-mass loss event and leading to this dimming \citep{dupree_great_2022}.
-Other hypothesis have  also been explored , such as a drop in temperature \citep{harper_photospheric_2020}, or an increase in the molecular opacity \citep{kravchenko_atmosphere_2021}.
-This event is an extreme illustration of  changes in brightness that were not due to any kind of pulsation but to a change on the 
+From interferometric and spectroscopic data, it has been proposed that this event was caused by the formation of a cloud of dust close to the line of sight \citep{montarges_dusty_2021}. 
+Interferometric images show a drop in luminosity in the southern hemisphere of Betelgeuse, which could be caused by a mass loss event and leading to this dimming \citep{dupree_great_2022}.
+Other hypotheses have also been explored, such as a drop in temperature \citep{harper_photospheric_2020}, or an increase in the molecular opacity \citep{kravchenko_atmosphere_2021}.
+In this event, it turns out that a change in brightness may not be due to pulsation but to a change in the
 brightness distribution over the stellar disk.
 %due to the presence of large quantities of dust.
-Since the end of the Great Dimming event, Betelgeuse has continued its random variation of brightness. 
+Since the end of the Great Dimming event, Betelgeuse has continue its random variation of brightness. 
 But quite interestingly, 
 it has been shown by \cite{jadlovsky_analysis_2023} and \cite{dupree_great_2022} that the periodicity of Betelgeuse has changed since the dimming. Using the light curves from AAVSO, 
-the cited authors showed that before the dimming, the dominant period of Betelgeuse was the 400 days period, while after the great dimming the main periods of Betelgeuse have shortened, oscillating between 97 days and 230 days \citep{dupree_great_2022}, revealing a change in the behaviour of the atmosphere of Betelgeuse.
+the cited authors showed that before the dimming, the dominant period of Betelgeuse was the 400 d period, while after the great dimming, the main periods of Betelgeuse have shortened, oscillating between 97 d and 230 d \citep{dupree_great_2022}, revealing a change in the behaviour of the atmosphere of Betelgeuse.
 While the great dimming may be seen as 
-a singular event, such modifications of the variability periods  bring up the question whether all changes in brightness cannot be due to similar changes 
-in the brightness  distribution over the disk, and unrelated to any pulsation phenomenon. 
-%Our interest in an alternative explanation of the variability
-%in terms  of convective patterns arose. 
-In order to address this question we  seek for the typical periods of 400 and 200 days in  
-observational proxies which are related to the convective activity but not to any pulsation, such as the linear polarization spectra.\\
+a singular event, such modifications of the variability periods bring up the question of whether all changes in brightness can be due to similar changes in the brightness 
+distribution over the disk and unrelated to any pulsation phenomenon. 
+Our interest in an alternative explanation of the variability
+in terms  of convective patterns arose. In order to address this question, we seek such typical periods of 400 and 200 d in  
+observational proxies related to the convective
+activity but not to any pulsation, such as the linear polarization
+spectra.\\
 %but unrelated to pulsations and which cannot be directly linked to brightness.
@@
 Linear polarization in the atomic lines of the spectrum of Betelgeuse, discovered by \cite{auriere_discovery_2016}, has been interpreted as the joint action of 
-two mechanisms. First the depolarization of the continuum by atoms which absorb linearly polarized light from the continuum and re-emit unpolarized light, the 
-continuum photons being polarized by Rayleigh scattering. This depolarization produces signals with an azimuth symmetry over the disk that 
-would  cancel out the net linear polarization on a homogeneous disk.
-Thus, such mechanism of production of polarization must be combined with an inhomogenous disk to produce the net linear polarization
-signal observed. This interpretation suggested the possibility of mapping those brightness inhomogeneities. This has been done by \cite{lopez_ariste_convective_2018} 
-even producing 3-dimensional images of the atmosphere of Betelgeuse \citep{lopez_ariste_three-dimensional_2022} when taking advantage of the different heights of formation of 
-different lines in the spectrum of Betelgeuse. The images produced compare very well with contemporaneous images made with interferometric 
-techniques \citep{montarges_close_2016} and show clear convective patterns, akin to solar granulation.
+two mechanisms. First, the depolarization of the continuum by atoms which absorb linearly polarized light from the continuum and re-emit unpolarized light,
+the continuum photons being polarized by Rayleigh scattering.
+This depolarization produces signals with circular symmetry over the visible disk, which would  cancel out the net linear polarization on a homogeneous disk.
+Thus, such mechanism of polarization must be combined with an inhomogeneous disk to produce the net linear polarization
+signal observed.
+This interpretation suggested the possibility of mapping those brightness inhomogeneities.
+This has been achieved by \cite{lopez_ariste_convective_2018}, who produced even 3-dimensional images of the atmosphere of Betelgeuse \citep{lopez_ariste_three-dimensional_2022} by taking advantage of the different heights of formation of
+different lines in the spectrum of Betelgeuse. The produced images compare very well with contemporaneous images made with interferometric 
+techniques \citep{montarges_close_2016} and show clear convective patterns, akin to solar granulation.\
 
-Linear polarization observed in the atomic lines is therefore a proxy of convection, unrelated to radial 
-pulsations but linked to the brightness inhomogeneities due to the convective patterns in Betelgeuse. If one can find the 
-aforementioned variability periods 
-in linear polarization, we must conclude that these periods are related to the convective activity which originates the linear 
-polarization signals.
+Linear polarization observed in the atomic lines is therefore a proxy of convection, a priori unrelated to radial 
+pulsations but linked to the brightness inhomogeneities due to the convective patterns in Betelgeuse. If one can find the aforementioned variability periods 
+in linear polarization, we must conclude that these periods are related to the convective activity which originates the linear polarization signals.
 %and unrelated to any radial pulsation phenomena. 
-This is the purpose of the present work.
+This is the purpose of the present work.\
 
-In section \ref{Section 2} we describe the dataset of linear polarization obtained with the Telescope Bernard Lyot (TBL) and the Lomb-Scargle technique used to derive periods. 
-In section \ref{section 3} we associate the 200 days period to convective timescale of the smallest granules and the 330 days periodicity to the largest granules, and we provide an explanation in the change of variability of Betelgeuse before and after the great dimming.
-In section \ref{Conclusion}, we summup our work.
+In section \ref{Section 2}, we describe the dataset of linear polarization obtained with Narval and Neo-Narval at TBL, as well as the Lomb-Scargle periodograms used to derive periods.
+In section \ref{section 3}, we seek periods in the LSD profiles of Betelgeuse using the Lomb-Scarlge periodogram.
+In section \ref{section 4}, we associate the 200 d period with the convective timescale of the smallest granules and the 330 d periodicity with the largest granules.
+We speculate on an explanation in the change of variability of Betelgeuse before and after the great dimming.
 
-\section{Searching for periodicities in the polarization spectra of Betelgeuse}
+\section{The polarimetric data}
 \label{Section 2}
 
-Betelgeuse has been observed for the last 10 years with Narval and Neo-Narval at the Telescope Bernard Lyot (TBL). These two instruments measure the polarization 
-over the visible spectra of Betelgeuese (390-1000 nm) with high spectral resolution (R=65000) and high polarimetric sensitivity. Despite 
-this high sensitivity, the signal-to-noise ratios are not sufficient to measure the weak polarization signals in individual atomic lines of the 
-spectrum. These amplitudes are today known to be of the order of $10^{-4}$ times the continuum intensity, and only exceptionally do they 
-reach amplitudes of $10^{-3}$ the continuum. In these exceptional observations, enough photons can be accumulated per spectral bin to 
-see the linear polarization signal above noise \citep{auriere_discovery_2016}. Most commonly, the amplitudes of linear polarization are below noise levels. In those 
-occasions we have to add up the signals of thousands of lines to reduce the noise and increase the signal-to-noise ratios. This line addition 
-is done through a technique called Least-Squares Deconvolution (LSD) \citep{donati_spectropolarimetric_1997} and assumes that the 
-linear polarization signal is similar 
-in all spectral lines up to scale factors in the sampling and amplitude. This technique has been successfully used in the past to measure 
-magnetic field distributions over stellar surfaces and it is now used to produce images of the brightness distributions in the photosphere 
-of Betelgeuse and other red supergiants as cited in the Introduction.
+Betelgeuse has been observed for the last 10 years with Narval and Neo-Narval at the Telescope Bernard Lyot\footnote[1]{https://tbl.omp.eu/}. These two instruments measure the polarization 
+over the visible and near infrared spectra of Betelgeuese (390-1000 nm) with high spectral resolution (R=65000) and high polarimetric sensitivity.
+The signal-to-noise ratios are not sufficient to measure the weak polarization signals in individual atomic lines of the
+spectrum.
+These amplitudes are known to be of the order of $10^{-4}$ times the continuum intensity, and they only exceptionally reach amplitudes of $10^{-3}$.
+In those observations, enough photons can be accumulated per spectral bin to
+detect the linear polarization signal above noise in lines \citep{auriere_discovery_2016}. Most commonly, the amplitudes of linear polarization are below noise levels. On those 
+occasions, we have to add up the signals of thousands of lines to increase the signal-to-noise ratios. This addition 
+is performed through a technique called Least-Squares Deconvolution (LSD) \citep{donati_spectropolarimetric_1997}.% and assumes that the 
+%linear polarization signal is similar 
+%in all spectral lines up to scale factors in the sampling and amplitude.
+This technique has been successfully used in the past to measure 
+magnetic field distributions over stellar surfaces and is now employed to produce images of the brightness distributions in the photosphere 
+of Betelgeuse and other RSGs, as cited in the Introduction.
 
-In this work we shall look into the signals produced through these line-addition techniques, and which produce a pseudo-spectral line in intensity 
-and linear polarization which does not belong to any atomic species in particular but to an average of all of them present and emitting in the 
-photosphere of Betelgeuse. After such procedures of line addition, these signals carry the coherent signals present in the photospheric lines, but the particularities of this or 
-that spectral line are erased. All these data has been presented before by \cite{auriere_discovery_2016}, \cite{mathias_evolution_2018} and 
-\cite{lopez_ariste_three-dimensional_2022}, which also give details on observing times and conditions as well as more detailed description 
-on the data reduction. \footnote[1]{Beyond a 2-year proprietary embargo, and up to technical issues, all these data is available through 
-PolarBase (http://polarbase.irap.omp.eu/).}
+In this work, we shall examine the signals produced through the LSD technique, which produce a pseudo-spectral line in intensity 
+and linear polarization. This pseudo-spectral line does not correspond to any particular atomic species but rather represents an average of all species present and emitting in the 
+photosphere of Betelgeuse.
+Following these procedures of line addition, the resulting profiles carry the coherent signals present in the photospheric lines, but the specific characteristics of individual spectral lines are erased.
+All of this data has been previously presented before by \cite{auriere_discovery_2016}, \cite{mathias_evolution_2018} and
+\cite{lopez_ariste_three-dimensional_2022}, which also provide details on observing times and conditions as well as a more detailed description 
+of the data reduction process\footnote[2]{Beyond a 2-year proprietary embargo, and up to technical issues, all these data is available through 
+PolarBase (http://polarbase.irap.omp.eu/).}.\\
+Note that all the velocities mentioned in the text or in the figures are provided in the heliocentric frame.
 
@@
 
-\subsection{The variability present in the stokes parameters}
+\section{Period search}
+\label{section 3}
 
-Trying to find periods in an astrophysical context carries the hurdle of unevenly spaced observed data. To overcome this issue, 
-we used the Lomb-Scargle periodogram \citep{lomb_least-squares_1976,scargle_studies_1982}. For a given set of data, it fits sinusoidal 
-using least-squares at each frequency between the first and the last observation. The better the fit, the higher the power attributed to the frequency. 
-We applied this technique to the  LSD profiles of Stokes $I$, $U$, $Q$ and to the total linear polarization of Betelgeuse observed by the TBL during 
-that span of 10 years.
+Attempting to identify periods in an astrophysical context faces the challenge of unevenly spaced observed data. To overcome this hurdle, 
+we used the Lomb-Scargle periodogram \citep{lomb_least-squares_1976,scargle_studies_1982}.
+This method involves fitting sinusoids to the data using least-squares with frequency sampled between the first and last observation.
+The quality of the fit determines the power attributed to a frequency.
+We applied this technique to the  LSD profiles of Stokes $I$, $U$, $Q$ and to the total linear polarization of Betelgeuse observed by the TBL over a span of 10 years (2014-2024).
+However, the dataset available post-dimming is insufficient to produce meaningful periodograms.
+Therefore, our focus was on periodograns computed exclusively from data obtained before the great dimming, covering the period from 2014 to the end of 2019.
 
-We first computed the Lomb-Scargle periodogram at each wavelength of the LSD profile, between -25 km/s and +60 km/s to span the entire 
-signal present in the LSD profile. \cite{lopez_ariste_convective_2018} found the most blueshifted signal towards -20 km/s, 
-and interpreted it as  the maximum velocity of the rising plasma, seen from the Earth. The most redshifted signal was found at +40 km/s. 
-This velocity is interpreted
-as the rest velocity of the star in our reference frame. Seldom, signals at velocities larger than +40 km/s can be found 
-and can be interpreted as either dark and cold plasma falling back to Betelgeuse, 
-or to rising plumes of plasma located in the hidden face of Betelgeuse,
-and rising so high that they appear above the limb.
+\subsection{Variability of the Stokes parameters}
+
+We first computed the Lomb-Scargle periodogram at each wavelength of the LSD profile, ranging from -25 km/s to +60 km/s to cover the entire 
+signal present in the LSD profile.
+\cite{lopez_ariste_convective_2018} identified the most blueshifted signal towards -20 km/s and interpreted it as  the maximum velocity of the rising plasma.
+The most redshifted signal was often found at +40 km/s, interpreted
+as the rest velocity of the star.
+Sometimes, signals at velocities greater than +40 km/s can be found, which can be interpreted either as dark and cold plasma falling back to Betelgeuse or as rising plumes of plasma located on the hidden face of Betelgeuse, ascending so high that they appear above the limb \citep[for the case of the RSG $\mu$Cep]{lopez_ariste_height_2023}.
 
@@
 
-Figures \ref{LS intensity},\ref{LS Q},\ref{LS U} and \ref{LS linear polarization} show the Lomb-Scargle periodogram of the LSD profile of respectively 
-the intensity, stokes $Q$, stokes $U$ and the linear polarization. 
+Figures \ref{LS intensity},\ref{LS Q},\ref{LS U} and \ref{LS linear polarization} show the Lomb-Scargle periodogram of the LSD profile for Stokes $I$, Stokes $Q$, Stokes $U$ and the total linear polarization, respectively.
 %The abscissa is the frequency (in $\mathrm{days^{-1}}$), while the ordinate is the heliocentric radial velocity in km/s. 
-For each figure, the uper panel is the Lomb-Scargle periodogram of each velocity bin (black line) and the total average (red line).
-The green line corresponds to the window function.
-The lower panel is the Lomb-Scargle periodogram represented for each velocity bin, while the white dotted line refer to 400 and 200 days.
-The right panel is the mean profile of each stokes parameter for each velocity bin.
+In each figure, the upper panel depicts the Lomb-Scargle periodogram of each velocity bin (black line) along with the total average (red line).
+The green line represents the window function.
+The lower panel shows the Lomb-Scargle periodogram for each velocity bin, with the white dotted lines marking the 400 and 200 d periods for reference.
+The right panel represents the mean profile of each Stokes parameter at each velocity bin.
 
@@
     The upper panel is the Lomb-Scargle periodogram for each velocity bin (black line) and the average (red line). The green line is the window function.
-    In the lower pannel, the two white dashed lines represent respectively the 400 days and the 200 days periods. The right panel is the average intensity profile for each velocity bin.}
+    In the lower panel, the two white dashed lines mark respectively the 400 d and the 200 d periods. The right panel is the average intensity profile for each velocity bin.}
     \label{LS intensity}
@@
     \includegraphics[width=0.5\textwidth]{Lomb-Scargle Stokes Q.pdf}
-    \caption{Same as figure \ref{LS intensity} for stokes $Q$. }
+    \caption{Same as Fig. \ref{LS intensity} for Stokes $Q$. }
     \label{LS Q}
@@
     \includegraphics[width=0.5\textwidth]{Lomb-Scargle Stokes U.pdf}
-    \caption{Same as figure \ref{LS intensity} for stokes $U$.}
+    \caption{Same as Fig. \ref{LS intensity} for Stokes $U$.}
     \label{LS U}
@@
     \includegraphics[width=0.5\textwidth]{Lomb-Scargle linear polarization.pdf}
-    \caption{Same as figure \ref{LS intensity} for linear polarization: $\sqrt{Q^2+U^2}$.}
+    \caption{Same as Fig. \ref{LS intensity} for the total linear polarization: $\sqrt{Q^2+U^2}$.}
     \label{LS linear polarization}
@@
 %\subsection{Intensity}
-Figure \ref{LS intensity} shows the Lomb-Scargle periodogram of the LSD of the intensity profile.
-The 400 and 200 days period seem to be recovered by the periodogram.
-However, the 400 days period is uncomfortably close to the peak of the window function, while the 200 days period is not.
-We also notice that the two periods are present at the same HRV.
-Interestingly, \cite{mathias_evolution_2018} also found this 200 days periodicity with a shorter observation period.
+Examining Fig.\ref{LS intensity}, which shows the Lomb-Scargle periodogram of the LSD of the intensity profile, we observe that the 400 and 200 d periods seem to be captured by the periodogram.
+However, the 400 d period is uncomfortably close to the peak of the window function.
+Furthermore, both periods are located at the same HRV and in the blue wing of the profile.
+Interestingly, \cite{mathias_evolution_2018} previously found this 200 d periodicity in spectroscopic observation, despite of a shorter observation period.
 
-Figure \ref{LS Q} and \ref{LS U} show the Lomb-Scargle periodogram of both stokes $Q$ and $U$. It is clear that the periodogram is completely different 
-from the intensity one.
-A strong signal is present in fig.
-\ref{LS Q} around 200 days.
-Concerning stokes U in fig.
-\ref{LS U}, it appears that the main period is around $0.003 \ \mathrm{days^{-1}}$ (330 days).
-Other period such as the one at 200 days or 250 days ($0.004 \ \mathrm{days^{-1}}$) are present but difficult to trust.
-From stokes $Q$ and $U$, we recover the 200 days period and the 330 days period is more or less close to the 400 days period found in the literature.
-Figure \ref{LS linear polarization} shows the Lomb-Scargle periodogram of the total linear polarization of Betelgeuse: $\sqrt{Q^2+U^2}$. 
-Interestingly, the periodogram shows a strong power at both 330 days and 200 days, which are the period that we have found in both stokes $Q$ and $U$.
+Figures \ref{LS Q} and \ref{LS U} display the Lomb-Scargle periodograms of Stokes $Q$ and $U$.
+These periodograms exhibit significant differences compared to the intensity one.
+In Fig.
+\ref{LS Q}, a prominent signal is evident around 200 d.
+Regarding Stokes $U$ in Fig.
+\ref{LS U}, it appears that the primary period is approximately $0.003 \ \mathrm{d^{-1}}$ (equivalent to 330 d).
+Although other periods, such as those at 200 d or 250 d ($0.004 \ \mathrm{d^{-1}}$) are present, but are difficult to trust.
+From both periodograms of Stokes $Q$ and $U$, we recover the 200 d period, and also the 330 d period is notable, which aligns closely with the 400 d period reported in the literature and is in agreement with the timescale of the hysteresis loop reported by \cite{kravchenko_tomography_2019}.
+Figure \ref{LS linear polarization} shows the Lomb-Scargle periodogram of $\sqrt{Q^2+U^2}$, representing the total linear polarization of Betelgeuse. 
+This periodogram confirms the significant powers at both 330 d and 200 d, consistent with the periodograms of Stokes $Q$ and $U$.
 
@@
 
-\subsection{The variability of Betelgeuse inferred from polarimetry imaging}
+\subsection{Variability of the polarimetric imaging}
 
-Using linear polarization, \cite{lopez_ariste_convective_2018} have been able to reconstruct images of Betelgeuse that have been successfully compared 
-to inteferometric images of \cite{montarges_close_2016}. The images are produced by finding the brightness distribution that better fits the observed linear 
+Using linear polarization, \cite{lopez_ariste_convective_2018} successfully reconstructed images of Betelgeuse, which have been compared favourably to inteferometric images obtained by \cite{montarges_close_2016}.
+The images are produced by finding the brightness distribution that better fits the observed linear
 polarization LSD profile using a Marquardt-Levenberg minimisation.
 
-Betelgeuse is observed in average every month by the TBL, and hence
-we are able to follow its surface activity through polarimetry imaging. This has  allowed in the past the estimation of  the size and the
- lifetime of the convective cells
-at the surface \citep{lopez_ariste_convective_2018}. From these images, we computed a photo-center, which is sensitive to the size and the number of convective cells. 
-An homogeneous star will have a photo-center displacement coinciding with barycenter of the star, whereas a star with one or two large convective cells will have a
-photo-center displacement more important, up to a few percent of the stellar radius in the case of red supergiants \citep{chiavassa_probing_2022}.
+Betelgeuse is observed, on average, every month by the TBL, enabling the tracking of its surface activity through this technique of polarimetric imaging.
+This technique has previously allowed for the estimation of  the size and the
+ lifetime of convective cells on the surface \citep{lopez_ariste_convective_2018}. From these images, we computed a photo-center, a quantity sensitive to the size and number of convective cells. 
+A homogeneous star will have a photo-center displacement coinciding with the barycenter of the star, whereas a star with one or two large convective cells will exhibit a more siginificant photo-center displacement, up to a few percent of the stellar radius in the case of RSGs \citep{chiavassa_probing_2022}.
 
 Since the photo-center is linked to surface convection, it is worth checking for periods in its 
-dynamics over the 10 years of observations of Betelgeuse at the TBL. Such periods will not 
-be completely unrelated to the ones presented in the previous section, but will capture 
-other aspects of the phenomena at work.
+dynamics over the 5 years of observations of Betelgeuse before the dimming.
+While these periods may overlap with those presented in the previous section, they are likely to capture additional aspects of the phenomena at work.
 
 %We computed the Lomb-Scargle periodogram of the displacement of the photo-center.
-But, before going any further, we have to warn
- about the interpretation of such images. 
-Linear polarization suffers from a $180 ^\circ$ degree ambiguity already mentioned in \cite{auriere_discovery_2016}. For that reason, our images 
+However, before proceeding to search for periods in the displacement of the photo-center, it is important to address a key issue regarding the interpretation of such images. 
+Linear polarization suffers from a $180 ^\circ$ degree ambiguity, as mentioned in \cite{auriere_discovery_2016}. Consequently, our images 
 can be 
 rotated by $180^\circ$ degree, and the brightness distribution will still fit the observed LSD polarization profiles. 
-If a particular observation were to be treated independently of previous observations, the solution found by 
-the algorithm could be any of the ambiguous solutions possible. To obtain a continuity 
-between the series of images, 
-for a given day we used the brightness distribution of the previous day as initial point 
-of the  fitting iteration; and for the first image, we start with a 
-random brightness distribution as the beginning of the fit. 
-Even though we are able to produce a series of images  consistent with each other, it is important 
-to keep in mind that the photo-center displacement computed from the series 
-will be affected by the choice of the first image. 
-To overcome this issue, we decided to compute
-the photo-center displacement from 100 different initial parameters and their subsequent series 
-of images in an effort to recover from the ensemble
-properties that are independent of the choice of the first image.
+If each observation were treated independently, the algorithm's solution could be any of the possible ambiguous solutions. To ensure continuity 
+between the image series, 
+for a given day, we use the brightness distribution of the previous day as the initial point 
+of the  fitting iteration. The first image in the series begins its minimisation iteration with a random brightness distribution. 
+Although we can produce a series of consistent images, it is important to keep in mind that the photo-center displacement computed from the series 
+will be affected by the initial image. 
+To overcome this issue, we computed
+the photo-center displacement from 100 different time series, each starting from a different initial image.
+This approach aims to recover ensemble properties independent of the choice of the first image.
 
@@
     \caption{Lomb-Scargle periodogram of the 100 photo-center displacement of Betelgeuse. Each blue lines correspond to a Lomb-Scargle periodogram of one photo-center
-     displacement. The black line is the average of the 100 periodogram. The red dashed lines represent the 400 and 200 days period respectively 
-     (from left to right). As before, the green line is the window function.}
+     displacement. The black line is the average of the 100 periodogram. The red dashed lines represent the 400 and 200 d period respectively 
+     (from left to right). The green line is the window function.}
     \label{LS photocenter}
@@
 
-Figure \ref{LS photocenter} shows the Lomb-Scargle periodogram of the photo-center displacement of each one of those 100 series (blue lines) and the average Lomb-Scargle periodogram (black line), 
+Figure \ref{LS photocenter} shows the Lomb-Scargle periodogram of the photo-center displacement for each of the 100 series (blue lines) and the average Lomb-Scargle periodogram (black line). 
 %As before, the upper panel shows the Lomb-Scargle periodogram from data before the great dimming only, whereas we used the whole set of data for the lower panel.
-while the black dashed lines mark the 400 and 
-200 days periods respectively.
-As before, the window function is represented by the green line.
-Interestingly, we find two bumps around the 400 days period.
-We cannot directly confirm that we recover the 400 days period from those small bumps, but since polarimetry imaging involves only surface convection, this is an argument that points toward a variability explained by surface convection.
-%First, we find that the Lomb-Scargle power is high around the LSP, though
-%uncomfortably close to the auto-correlation peak.
-%We also notice a decrease in the Lomb-Scargle power from  before to after the dimming, in 
-%concurrence to what was seen in the periodograms of the LSD profiles. Before it, there are two smalls bumps around the 400 days period. 
-%But those bumps almost completely disappear after the dimming. Even though it is 
-%hard to identify the 400 days period from those bumps alone,
-%we still find that there is a difference before and after the dimming. 
-%And this is concluded  from images that exclusively involve surface convection.
-%The Lomb-Scargle periodogram of the photo-center displacement points towards a variability explained by convection only,
-%and this is consistent with what we obtained from linear polarization.
-%Both the photo-center displacement and the LS of the linear polarization point toward a variability explained by surface convection.
+The red dashed lines indicate the 400 and 
+200 d periods, respectively.
+Similar to previous figures, the window function is represented by the green line.
+Interestingly, we find two peaks around the 400 d period in the periodogram, although they are not individually significant to confirm the presence of such period.
+However, these peaks are consistent with the findings of the previous section.
+Since polarimetry imaging involves only surface convection, this provides further support for a variability being explained by surface convection alone.
 
-\section{The variability of the light curve of Betelgeuse}
-\label{section 3}
+\subsection{Variability of the light curve}
+
+After examining the periods identified by the Lomb-Scargle technique in the polarization data obtained by the TBL over the last years before the dimming, it is worth contextualizing them alongside the periods traditionally identified in light curves over the same period of time.
+Two aspects
+are of our interest in this comparison: the behaviour of the light curve before and after the great dimming.
+
+\begin{figure}[!h]
+    \centering
+    \includegraphics[width=0.5\textwidth]{Light_curve_Betelgeuse.pdf}
+    \caption{Light curve of Betelgeuse from AAVSO in the V-band.}
+    \label{light curve Betelgeuse}
+\end{figure}
+
+We have retrieved the light curve of Betelgeuse in the visible from the AAVSO database for the past 10 years, as shown in Fig. \ref{light curve Betelgeuse}.
+Before the great dimming, Betelgeuse's magnitude exhibited variations on a yearly timescale, whereas after the dimming, its variability has shortened. 
+Figure. \ref{light curve Betelgeuse} clearly illustrates that the 
+variability of Betelgeuse now occurs on a timescale shorter than one year. This qualitative change has been pointed out before,
+before the great dimming, the primary period of Betelgeuse was approximately 400 d \citep{kiss_variability_2006}, often associated with the fundamental pressure mode.
+However, after the dimming, this period seems to have vanished, and only timescales shorter than 230 d are visible since \citep{dupree_great_2022}.
+No explanations have been put forth regarding this change of variability after, or perhaps because of the dimming.
+Consequently, we can afford proposing alternative scenarios for the observed periodicities in the light curve (see Section \ref{section 4}).\\
+
+We utilized the AAVSO database to search for periodicities in the light curve of Betelgeuse using the Lomb-Scargle periodogram, aiming to identify periods previously reported in the litterature.
+In Fig.\ref{Lomb Scargle AAVSO}, we computed the periodogram from the light curve of Betelgeuse spanning from 1990 to 2024 (blue line), alongside the periodogram computed using AAVSO observations corresponding to the observation dates of the TBL (orange line) within a range of 5 days.
+For the periodogram since 1990, we binned the observations with an interval of 10 days, as Betelgeuse has been more observed in the 21st century \citep{kiss_variability_2006}.
+Our analysis reveals the recovery of the 400 d period and also a period close to 200 d, while the other peaks are due to the windowing effect.
+Thus, the Lomb-Scargle periodogram succssfully identifies periods mentioned in the literature.
+However, when focusing on the periodogram derived from the light curve corresponding to the TBL observation dates, we fail to retrieve the periods mentioned in the literature.
+Hence, the small peaks present in Fig.
+\ref{LS photocenter} are attributed to the limited number of TBL observations.
+
+\begin{figure}[!h]
+    \centering
+    \includegraphics[width=0.5\textwidth]{Lomb-Scargle AAVSO.pdf}
+    \caption{Lomb-Scargle periodogram of the light curve of Betelgeuse since 1990 (blue line) and from the light curve where the dates correspond to the observation of the TBL.
+The two black dashed lines are respectively the 400 and 200 d periods.}
+    \label{Lomb Scargle AAVSO}
+\end{figure}
+
+\section{Discussion}
+\label{section 4}
 %Before summing up the conclusions of this work, we can afford proposing alternative scenarios for the observed periodicities in the light curve, 
@@
 The presence of the same 
-periods and, after taking into consideration the time span and sparsity of the TBL data series, the larger amplitude of the peaks in
-polarization data  than in the light curve point to a common origin, one that is rather based upon convective dynamics.
-Since we associate stokes $Q$ and $U$ to proxies of convection, we associate the 200 days period to the typical timescales on which the Stokes parameters evolve.
-The LSD profiles of stokes $Q$ or $U$ will slightly change when we observe Betelgeuse every month, but are completely overhauled 
-on time scales of several months. The Lomb-Scargle periodograms appear to capture this spectral dynamic around periods of 200 days.
- Such changes in the profiles are directly linked to the dynamics of the surface,  to  the evolution and movements 
-of convective cells over the visible hemisphere. So the 200-day period appears to be related to the evolution of the convective patterns 
-on the surface of Betelgeuse.  The fact that this peak is strongly present in stokes $Q$ but less in $U$ points toward a more stochastic origin, which can fully be explained by convection.\
+periods and, considering the time span and sparsity of the TBL data series, the larger amplitude of the peaks found in
+polarization data compared to those in the light curve point to a common origin, one that is rather based upon convective dynamics.
+Since we associate Stokes $Q$ and $U$ to proxies of convection, we associate the 200 d period to the typical timescales on which the Stokes parameters evolve.
+The LSD profiles of Stokes $Q$ or $U$ undergo slight changes when observing Betelgeuse monthly, but are completely overhauled 
+on time scales of several months.
+The Lomb-Scargle periodograms seem to capture this spectral dynamic at periods of aproximately 200 d.
+Such changes in the profiles are directly linked to the dynamics of the surface,  to  the evolution, and movements
+of convective cells across the visible hemisphere. Therefore, the 200 d period appears to be related to the evolution of the convective patterns 
+on the surface of Betelgeuse.  The stronger presence of this peak in Stokes $Q$ compared to $U$ could be interpreted as a temporal situation of how convective patterns have been distributed in the
+last years.\
 
-The other period found in the linear polarization and stokes $U$, the 330 days periods, is more or less close to the 400 days period often present in the literature (e.g \cite{kiss_variability_2006} found a period $388 \pm 30$ days).
-It has been shown by \cite{lopez_ariste_convective_2018} that large convective cells can live up to one year, thus this period would be associated to the timescale on which the largest granules evolve.
-This typical timescale has also been observed in the numerical simulations and might play an important role in the determination of stellar distances through the displacement of the photo-center \citep{chiavassa_probing_2022}.
-Once again, the 330 days periodicity being more present in stokes $U$ than $Q$ unravels a stochastic process that is due to convective motions, and is consistent with the work of \cite{gray_mass_2008}, who attributed the 400 days period to the large convective cells.
+The other period found in the total linear polarization and Stokes $U$, the 330 d periods, is not so far from the 400 days period often mentioned in the literature (e.g \cite{kiss_variability_2006} found a period of $388 \pm 30$ d).
+It has been shown by \cite{lopez_ariste_convective_2018} that large convective cells can persist for one to two years.
+Thus, the 330 d period could be associated to the timescale on which the largest granules evolve.
+This characteristic timescale has also been observed in the numerical simulations and might play an important role in determining stellar distances through the displacement of the photo-center \citep{chiavassa_probing_2022}.
+Once again, we observe the 330 d periodicity to be more prominent in Stokes $U$ than in $Q$, what may unravel a random situation due to convective motions in the last years.
+This interpretation aligns with the work of \cite{gray_mass_2008}, who attributed the 400 d period to large convective cells.
 \
 
-Finally, the intensity profile seems to match the 400 days periods, but is uncomfortably close to a peak of the window function of the TBL and a secondary peak close to 200 days.
-Since we are able to recover the 200 days period in the stokes paramters, we attribute this period to the convective timescales, but concerning the 400 days period, it could be a mix of the typical convective timescales of large granules associated with stochastic pulsations exited by convective cavities.
-Up to this point, the model developed to interpret linear polarization does not involve pulsations, that is why we interpret the 200 days period as the convective turnover timescale.
-The 400 days period involve the intensity profile, thus the 400 days period is probably caused by both convective timescales and stochastic pulsations.
+Regarding the Stokes $I$ profile, it seems to correspond to the 400 d period, but is uncomfortably close to a peak of the window function of the TBL, making this peak difficult to trust.
+A secondary peak close to 200 d is also present in the Stokes $I$ profile.
+Since we are able to recover the 200 d period in the Stokes parameters, we attribute this period to convective timescales.
+Up to this point, the model developed to interpret linear polarization involves only surface convection, which is why we interpret the 200 d period as the convective timescale of the smallest structures, while the 400 d period corresponds to the timescale of the largest structures.\\
  %As showed by
@@
 
-\subsection{The light curve of Betelgeuse from the point of view of convection}
-
-After examining the periods found by the Lomb-Scargle technique on the polarization data, and its derivatives, obtained by the TBL over the last 
-10 years, it is worth putting them in the context of the periods found traditionally on light curves over the same period of time. Two aspects 
-are of our interest in this comparison: the behaviour of the light curve before and after the great dimming.
-%and the amplitudes of the peaks 
-%at the main cited periods.
-
-% In the literature, it is reported that before the great dimming the main period of Betelgeuse was of 400 days \citep{kiss_variability_2006}. 
-% This pulsation is often associated to the fundamental pressure mode. After the dimming, however this period appears to have disappeared 
-% and only the period at 200 days,  associated to the first overtone of the fundamental mode, is visible since. 
-% No explanations based upon pulsations have been ofered concerning this change of variability after or perhaps because of the dimming. But this 
-% is a behaviour that
-
-We have recovered the light curve of Betelgeuse in the visible from the AAVSO database for the last 10 years (Figure \ref{light curve Betelgeuse}).
-Before the great dimming, the magnitude of Betelgeuse evolves on a timescale of a year, whereas after the dimming, the variability of Betelgeuse has shortened. 
-It is clear from fig. \ref{light curve Betelgeuse} that the 
-variability of Betelgeuse is evolving on a timescale shorter than one year. This qualitative change has been pointed out before,
-before the great dimming the main period of Betelgeuse was of 400 days \citep{kiss_variability_2006}. 
-This period is often associated to the fundamental pressure mode. After the dimming however, this period appears to have disappeared 
-and only timescales shorter than 230 days are visible since \citep{dupree_great_2022}. 
-No explanations have been offered concerning this change of variability after or perhaps because of the dimming, thus we can afford proposing alternative scenarios for the observed periodicities in the light curve.
+Before summing up the results of our work, we can afford to explain the variability of Betelgeuse after the great dimming with surface convection.
+RSGs experience mass loss event, where plasma is ejected into the interstellar medium \citep{josselin_atmospheric_2007}.
+These events have been inferred in RSG such as $\mu$Cep, where \cite{lopez_ariste_height_2023} found an excess of linear polarization beyond the limb of the star, attributing it to rising plumes of plasma in the back hemisphere of the star.
+In the case of Betelgeuse, the excess of linear polarization beyond the limb is rare and not as pronounced as those observed in $\mu$Cep.
+Also, they do not change the variability of the star.
+However, since the great dimming, the variability of Betelgeuse has changed.
+This suggests that the great dimming differed significantly from the usual mass loss event, that it hustled the dynamics of the photosphere.
 \
 
-RSGs encounter mass loss event, where due to stellar winds, plasma is ejected in the interstellar medium.
-Those events have already been inferred in RSG like $\mu$Cep, where \cite{lopez_ariste_height_2023} found an excess of linear polarization beyond the limb of the star and associated it to rising plumes of plasma in the back hemisphere of the star.
-In the case of Betelgeuse, this kind of event happened only once or twice during the last 10 years, with no change in the variability of the star.
-However, since the great dimming, the variability of Betelgeuse has changed, meaning that this event was such different from the usual mass loss event, that it hustled the dynamics of the photosphere.
-\
-
-Considering a scenario where the variability of Betelgeuse is mostly governed by convection, before the great dimming, the variability of the star was linked to the typical timescale on which convection occurs, that is 200 days for the small structures and roughly 350 days for the large structures.
-The great dimming impacted the behaviour of the photosphere, such that the largest structures do not evolve on timescale of the order of 350 days but rather 200 days, that is the smallest convective structures.
-We hypothesise that since the dimming, the photosphere is not in equilibrium and the largest convective structures are destroyed by the turbulent motion of the photosphere.
-Therefore, it is hard to identify periods after the great dimming, which \cite{dupree_great_2022} noticed.
-We expect to recover the 400 days variability within the next years, which corresponds to the time for the photosphere to go back to equilibrium.
+In a scenario where the variability of Betelgeuse is primarily governed by convection, before the great dimming, the variability of the star was associated to the typical timescale at which convection occurs: approximately 200 d for the small structures and roughly 350 d for the large structures.
+The great dimming affected the behaviour of the photosphere in such a way that the largest structures no longer evolved on a timescale around 350 d, but rather on the timescale of the smallest convective structures, around 200 d.
+We hypothesise that since the great dimming event, the photosphere has not returned to equilibrium, and the largest convective structures have been disrupted by the turbulent motion of the photosphere.
+Consequently, identifying periods after the great dimming is hard, which \cite{dupree_great_2022} noticed.
+We expect that the 400 d variability will gradually reappear in the coming years as the photosphere returns to equilibrium.
+However, it is essential to note that this scenario is speculative but provides a rough explanation of how convective activity could lead to a change in variability since the great dimming.
+A more detailed explanation of this variability change falls beyond the scope of this paper and is left for future research.
 %But this 
@@
 %strong peaks seen at selected wavelengths in the LSD profiles, strong when compared to the peaks at similar periods arising from the light curve.
-
-\begin{figure}[!h]
-    \centering
-    \includegraphics[width=0.5\textwidth]{Light_curve_Betelgeuse.png}
-    \caption{Light curve of Betelgeuse from AAVSO in the V-band.}
-    \label{light curve Betelgeuse}
-\end{figure}
 
@@
 
-The main observation  presented in this work is that the same periods identified in the light curve of Betelgeuse are also seen in the intensity and polarization
-spectra measured with the TBL. Periods in the light curve have often been traditionally associated to pulsation due to pressure modes.
+The main result presented in this work is that the same periods identified in the light curve of Betelgeuse are also observed in the intensity and polarization
+spectra measured with the TBL. Traditionally, periods in the light curve have been associated to pulsations due to pressure modes.
 %But such pulsation would in no case affect the polarisation signals nor would explain the fact that the periods can be observed with stronger 
@@
 %And such origin cannot be pulsations.
-The interpretation of the linear polarization profiles is made in terms of convective structures in the atmosphere of Betelgeuse. We suggest 
-that it is in these convective structures and their temporal evolution that we will find the true reason for the observed periodicities both 
-in the light curve and in the spectropolarimetric observations.
-This is the main conclusion of this work: since we are able to recover the different periods in the linear polarization profiles, we associate the 200 days periodicyt to convective dynamics while the 400 days periodicity is associated to both stochastic pulsations and convective timescales of the largest granules.
-This conclusion is in agreement with the scenario proposed by \cite{gray_mass_2008}, who attributed the variability of Betelgeuse to surface convection.
+However, the interpretation of the linear polarization profiles is made in terms of convective structures in the atmosphere of Betelgeuse.
+We propose that the true reason for the observed periodicities in both the light curve and spectropolarimetric observations lies within these convective structures and their temporal evolution.
+This is the main conclusion of our work: since we are able to recover the different periods in the linear polarization profiles, we associate the 200 d periodicity to convective dynamics, while the 400 d periodicity is linked to convective timescales of the largest granules.
+This conclusion concurs with the scenario proposed by \cite{gray_mass_2008}, attributing the variability of Betelgeuse to surface convection.
 %the observed periods are related 
@@
 
-Although of secondary importance, we may suggest with more or less ease what phenomena are at the origin of the change of variability before and after the great dimming.
-We speculate that the great dimming event hustled the dynamics of the photosphere, so that the largest granule are destroyed by the turbulent motions caused by the non equilibrium of the photosphere.
-Since this event, the variability of Betelgeuse has shortened and is in adequacy with the timescale of which evolve the smallest convective cells, that is 200 days.
-We expect in the future to see the 400 days variability being the dominant period, the time for the photosphere to go back to equilibrium.
+Although of secondary importance, we can speculate on the phenomena underlying the change in variability before and after the great dimming event.
+We hypothesize that the great dimming event disrupted the dynamics of the photosphere, leading to the destruction of the largest granules by the turbulent motions caused by the non-equilibrium of the photosphere.
+Since this event, the variability of Betelgeuse has quickened in adequacy with the timescale of smaller convective cells, around 200 d.
+If this scenario is correct, we shall expect in the coming future, the re-emergence of the 400 d variability as the dominant period once the photosphere returns to some form of equilibrium.
 Longer time
-series of spectropolarimetric observations of Betelgeuse will be required to shed light on this
+series of spectropolarimetric observations of Betelgeuse will be necessary to provide further insights into this phenomenon.
 

### 2024-05-15 main.tex
test

@@
 literature, usually clustered around the so-called long secondary period (LSP) of 2000 d, plus shorter periods around 400 and 200 d \citep{kiss_variability_2006}.
-The origin of the LSP remains poorly understood: some have tried to find the explanation of the LSP in the lifetime of the large convective cells present at the surface of RSG \citep[e.g][]{stothers_giant_2010} while others have invoked a magnetic field \citep{wood_long_2004}.
+The origin of the LSP remains a source of discussion: some have tried to find the explanation of the LSP in the lifetime of the large convective cells present at the surface of RSG \citep[e.g][]{stothers_giant_2010} while others have invoked a magnetic field \citep{wood_long_2004}.
 The shorter periods of 400 and 200 d seem to have a different origin than the LSP.

### 2024-05-15 main.tex
test 2

@@
 The origin of the LSP remains a source of discussion: some have tried to find the explanation of the LSP in the lifetime of the large convective cells present at the surface of RSG \citep[e.g][]{stothers_giant_2010} while others have invoked a magnetic field \citep{wood_long_2004}.
-The shorter periods of 400 and 200 d seem to have a different origin than the LSP.
+The shorter periods of 400 and 200 d are given a different origin than the LSP.
 By studying a large sample of RSGs, \cite{kiss_variability_2006} attributed the 400 d period to radial pulsation modes, while the 200 d period being attributed to the first overtone \citep{joyce_standing_2020}, in agreement with the historical work of \cite{stothers_pulsation_1969}.\\
 
-Besides radial pulsation, surface activity leading to random brightness variation has also been mentioned to explain the variability of Betelgeuse, firstly theorised by \cite{schwarzschild_scale_1975}.
+Besides radial pulsation, surface activity leading to random brightness variations has also been mentioned to explain the variability of Betelgeuse, firstly theorised by \cite{schwarzschild_scale_1975}.
 More recently, \cite{gray_mass_2008} proposed that the 400 d period was a consequence of convective activity.
@@
 Other hypotheses have also been explored, such as a drop in temperature \citep{harper_photospheric_2020}, or an increase in the molecular opacity \citep{kravchenko_atmosphere_2021}.
-In this event, it turns out that a change in brightness may not be due to pulsation but to a change in the
+In this event, it turns out that a change in brightness was not  due to pulsation but to a change in the 
 brightness distribution over the stellar disk.
@@
 the continuum photons being polarized by Rayleigh scattering.
-This depolarization produces signals with circular symmetry over the visible disk, which would  cancel out the net linear polarization on a homogeneous disk.
+This depolarization produces signals with azimuthal symmetry over the visible disk, which would  cancel out the net linear polarization on a homogeneous disk.
 Thus, such mechanism of polarization must be combined with an inhomogeneous disk to produce the net linear polarization

### 2024-05-15 main.tex
test 5

@@
 
-   \title{The variability  of Betelgeuse explained by surface convection}
+   \title{The variability  of Betelgeuse explained by surface convection\thanks{Based on observations obtained at the T\'elescope Bernard Lyot
+   (TBL) at Observatoire du Pic du Midi, CNRS/INSU and Universit\'e de
+   Toulouse, France.}}
 
@@
 These amplitudes are known to be of the order of $10^{-4}$ times the continuum intensity, and they only exceptionally reach amplitudes of $10^{-3}$.
-In those observations, enough photons can be accumulated per spectral bin to
+When these high amplitudes are present, enough photons can be accumulated per spectral bin to
 detect the linear polarization signal above noise in lines \citep{auriere_discovery_2016}. Most commonly, the amplitudes of linear polarization are below noise levels. On those 
@@
 %The abscissa is the frequency (in $\mathrm{days^{-1}}$), while the ordinate is the heliocentric radial velocity in km/s. 
-In each figure, the upper panel depicts the Lomb-Scargle periodogram of each velocity bin (black line) along with the total average (red line).
+In each figure, the upper panel depicts the Lomb-Scargle periodogram of each velocity bin (grey lines) along with the total average (red line).
 The green line represents the window function.
@@
     \caption{Lomb-Scargle periodogram of the LSD profile of intensity.
-    The upper panel is the Lomb-Scargle periodogram for each velocity bin (black line) and the average (red line). The green line is the window function.
+    The upper panel is the Lomb-Scargle periodogram for each velocity bin (grey lines) and the average (red line). The green line is the window function.
     In the lower panel, the two white dashed lines mark respectively the 400 d and the 200 d periods. The right panel is the average intensity profile for each velocity bin.}
@@
 \ref{LS U}, it appears that the primary period is approximately $0.003 \ \mathrm{d^{-1}}$ (equivalent to 330 d).
-Although other periods, such as those at 200 d or 250 d ($0.004 \ \mathrm{d^{-1}}$) are present, but are difficult to trust.
+Other periods, such as those at 200 d or 250 d ($0.004 \ \mathrm{d^{-1}}$) are present, but are difficult to trust.
 From both periodograms of Stokes $Q$ and $U$, we recover the 200 d period, and also the 330 d period is notable, which aligns closely with the 400 d period reported in the literature and is in agreement with the timescale of the hysteresis loop reported by \cite{kravchenko_tomography_2019}.
@@
 can be 
-rotated by $180^\circ$ degree, and the brightness distribution will still fit the observed LSD polarization profiles. 
+rotated by $180^\circ$ degrees, and the brightness distribution will still fit the observed LSD polarization profiles. 
 If each observation were treated independently, the algorithm's solution could be any of the possible ambiguous solutions. To ensure continuity 
@@
     \includegraphics[width=0.5\textwidth]{Lomb-Scargle Photo-center.pdf}
-    \caption{Lomb-Scargle periodogram of the 100 photo-center displacement of Betelgeuse. Each blue lines correspond to a Lomb-Scargle periodogram of one photo-center
-     displacement. The black line is the average of the 100 periodogram. The red dashed lines represent the 400 and 200 d period respectively 
-     (from left to right). The green line is the window function.}
+    \caption{Lomb-Scargle periodogram of the 100 photo-center displacement series of Betelgeuse. Each blue line corresponds to a Lomb-Scargle periodogram of one photo-center
+     displacement. The black line is the average of the 100 periodograms. The red dashed lines represent the 400 and 200 d period respectively. The green line is the window function.}
     \label{LS photocenter}
@@
 
-We have retrieved the light curve of Betelgeuse in the visible from the AAVSO database for the past 10 years, as shown in Fig. \ref{light curve Betelgeuse}.
+We have retrieved the light curve of Betelgeuse in the visible from the AAVSO database for the past 10 years. It is shown in Fig. \ref{light curve Betelgeuse}.
 Before the great dimming, Betelgeuse's magnitude exhibited variations on a yearly timescale, whereas after the dimming, its variability has shortened. 
-Figure. \ref{light curve Betelgeuse} clearly illustrates that the 
+Figure \ref{light curve Betelgeuse} clearly illustrates that the 
 variability of Betelgeuse now occurs on a timescale shorter than one year. This qualitative change has been pointed out before,
@@
 However, after the dimming, this period seems to have vanished, and only timescales shorter than 230 d are visible since \citep{dupree_great_2022}.
-No explanations have been put forth regarding this change of variability after, or perhaps because of the dimming.
-Consequently, we can afford proposing alternative scenarios for the observed periodicities in the light curve (see Section \ref{section 4}).\\
+No explanations have been put forth regarding this change of variability after, or perhaps because of the dimming.% In  Section \ref{section 4},  we  afford proposing alternative scenarios for the observed periodicities in the light curve.\\
 
-We utilized the AAVSO database to search for periodicities in the light curve of Betelgeuse using the Lomb-Scargle periodogram, aiming to identify periods previously reported in the litterature.
-In Fig.\ref{Lomb Scargle AAVSO}, we computed the periodogram from the light curve of Betelgeuse spanning from 1990 to 2024 (blue line), alongside the periodogram computed using AAVSO observations corresponding to the observation dates of the TBL (orange line) within a range of 5 days.
+ In Fig.\ref{Lomb Scargle AAVSO}, we computed the periodogram from that light curve of Betelgeuse (blue line), alongside the periodogram computed exclusively with AAVSO observations made less than 2 days from our observation dates of the TBL (orange line).
  For the periodogram since 1990, we binned the observations with an interval of 10 days, as Betelgeuse has been more observed in the 21st century \citep{kiss_variability_2006}.
-Our analysis reveals the recovery of the 400 d period and also a period close to 200 d, while the other peaks are due to the windowing effect.
-Thus, the Lomb-Scargle periodogram succssfully identifies periods mentioned in the literature.
+This periodogram  recovers the 400 d period and also a period close to 200 d.
+Other visible peaks are rather
+ due to the windowing effect.
 However, when focusing on the periodogram derived from the light curve corresponding to the TBL observation dates, we fail to retrieve the periods mentioned in the literature.
-Hence, the small peaks present in Fig.
-\ref{LS photocenter} are attributed to the limited number of TBL observations.
+The periodicities found in the TBL observations
+ gain validity through this comparison: they were evident in the TBL data, but barely recovered in the AAVSO data when pruned to the same observing dates.
+We can anticipated that extended spectropolarimetric observations
+ of Betelgeuse will make these periodicities seen in linear polarization, and arising from convective structures, clearer and clearer.
 
@@
     \includegraphics[width=0.5\textwidth]{Lomb-Scargle AAVSO.pdf}
-    \caption{Lomb-Scargle periodogram of the light curve of Betelgeuse since 1990 (blue line) and from the light curve where the dates correspond to the observation of the TBL.
-The two black dashed lines are respectively the 400 and 200 d periods.}
+    \caption{Lomb-Scargle periodogram of the light curve of Betelgeuse since 1990 as retrieved from the AAVSO database (blue line). When only dates corresponding to TBL observations are retained in 
+    that light curve, the periodogram in orange color is obtaine. The two black dashed lines are respectively the 400 and 200 d periods for visual reference.}
     \label{Lomb Scargle AAVSO}
@@
 periods and, considering the time span and sparsity of the TBL data series, the larger amplitude of the peaks found in
-polarization data compared to those in the light curve point to a common origin, one that is rather based upon convective dynamics.
+polarization data when compared to those in the light curve point to a common origin in either observation, one that is rather based upon convective dynamics.
 Since we associate Stokes $Q$ and $U$ to proxies of convection, we associate the 200 d period to the typical timescales on which the Stokes parameters evolve.
@@
 Once again, we observe the 330 d periodicity to be more prominent in Stokes $U$ than in $Q$, what may unravel a random situation due to convective motions in the last years.
-This interpretation aligns with the work of \cite{gray_mass_2008}, who attributed the 400 d period to large convective cells.
-\
+This interpretation aligns with the work of \cite{gray_mass_2008}, who attributed the 400 d period to
+large convective cells.
+Whether this 330 d period and the commonly cited 400 d period are one and the same will have to wait for extended spectropolarimetric observations with the TBL that smooth the windowing function around these frequencies.
 
-Regarding the Stokes $I$ profile, it seems to correspond to the 400 d period, but is uncomfortably close to a peak of the window function of the TBL, making this peak difficult to trust.
-A secondary peak close to 200 d is also present in the Stokes $I$ profile.
-Since we are able to recover the 200 d period in the Stokes parameters, we attribute this period to convective timescales.
-Up to this point, the model developed to interpret linear polarization involves only surface convection, which is why we interpret the 200 d period as the convective timescale of the smallest structures, while the 400 d period corresponds to the timescale of the largest structures.\\
+Regarding the Stokes $I$ profile, a peak closer to the 400 d period appears, but it is once again uncomfortably close 
+to a peak of the window function of the TBL, making it difficult to trust.
+A secondary peak close to 200 d is also present in the Stokes $I$ profile, that we fully assimilate to the one found in the Stokes parameters, and that  we attribute to convective timescales.
+%Up to this point, the model developed to interpret linear polarization involves only surface convection, which is why we interpret the 200 d period as the convective timescale of the smallest structures, while the 400 d period corresponds to the timescale of the largest structures.\\
  %As showed by
@@
 
-Before summing up the results of our work, we can afford to explain the variability of Betelgeuse after the great dimming with surface convection.
-RSGs experience mass loss event, where plasma is ejected into the interstellar medium \citep{josselin_atmospheric_2007}.
-These events have been inferred in RSG such as $\mu$Cep, where \cite{lopez_ariste_height_2023} found an excess of linear polarization beyond the limb of the star, attributing it to rising plumes of plasma in the back hemisphere of the star.
-In the case of Betelgeuse, the excess of linear polarization beyond the limb is rare and not as pronounced as those observed in $\mu$Cep.
-Also, they do not change the variability of the star.
+Before summing up the results of our work, we can afford to propose an explanation for the variability of Betelgeuse after the great dimming with surface convection.
+RSGs experience mass loss events when plasma is ejected into the interstellar medium \citep{josselin_atmospheric_2007}.
+These events have been inferred in RSG such as $\mu$Cep at photospheric heights, where \cite{lopez_ariste_height_2023} found an excess of linear polarization beyond the limb of the star, attributing it to rising plumes of plasma in the back hemisphere of the star.
+In the case of Betelgeuse, the excess of linear polarization beyond the limb is rare and not as pronounced as those observed in $\mu$Cep. They do not appear to contribute to 
+ the variability of the star.
 However, since the great dimming, the variability of Betelgeuse has changed.
-This suggests that the great dimming differed significantly from the usual mass loss event, that it hustled the dynamics of the photosphere.
+This suggests that the great dimming differed significantly from the usual mass loss event, and perhaps it hustled the  photosphere.
 \
 
-In a scenario where the variability of Betelgeuse is primarily governed by convection, before the great dimming, the variability of the star was associated to the typical timescale at which convection occurs: approximately 200 d for the small structures and roughly 350 d for the large structures.
-The great dimming affected the behaviour of the photosphere in such a way that the largest structures no longer evolved on a timescale around 350 d, but rather on the timescale of the smallest convective structures, around 200 d.
-We hypothesise that since the great dimming event, the photosphere has not returned to equilibrium, and the largest convective structures have been disrupted by the turbulent motion of the photosphere.
-Consequently, identifying periods after the great dimming is hard, which \cite{dupree_great_2022} noticed.
-We expect that the 400 d variability will gradually reappear in the coming years as the photosphere returns to equilibrium.
-However, it is essential to note that this scenario is speculative but provides a rough explanation of how convective activity could lead to a change in variability since the great dimming.
-A more detailed explanation of this variability change falls beyond the scope of this paper and is left for future research.
+In a scenario where the variability of Betelgeuse is primarily governed by convection,  the visible periods of such variability  were associated, before the great dimming, to the 
+typical timescales at which convection occurs: approximately 200 d for the small structures and roughly 350 d for the large structures.
+The great dimming affected the behaviour of the photosphere in such a way that the largest structures
+no longer evolved on a timescale around 350 d, but rather on the shorter timescales of the smallest convective structures, around 200 d.
+We hypothesise that after the great dimming event, the photosphere has not yet returned to equilibrium, and the largest convective structures
+have been disrupted by the turbulent motions present of the photosphere in this period.
+This would explain why identifying the longer periods after the great dimming has been hard, as \cite{dupree_great_2022} noticed.
+We may expect that the 400 d
+period will gradually reappear in the coming years as the photosphere returns to equilibrium. However speculative, this scenario 
+provides a rough explanation of how convective activity could lead to a change in variability since the great dimming that future observations and 
+research will confirm or infirm.
+%A more detailed explanation of this variability change falls beyond the scope of this paper and is left for future research.
 %But this 
@@
 We propose that the true reason for the observed periodicities in both the light curve and spectropolarimetric observations lies within these convective structures and their temporal evolution.
-This is the main conclusion of our work: since we are able to recover the different periods in the linear polarization profiles, we associate the 200 d periodicity to convective dynamics, while the 400 d periodicity is linked to convective timescales of the largest granules.
+This is the main conclusion of our work: since we are able to recover the different periods in the linear polarization profiles, we can exclude pulsation phenomena as the explanation for those periodicities in the light curve.
 This conclusion concurs with the scenario proposed by \cite{gray_mass_2008}, attributing the variability of Betelgeuse to surface convection.
+ We further find hints in the polarized spectra that lead us to associate the 200 d periodicity to convective dynamics, while the 400 d periodicity may be linked to convective timescales of the largest granules.
 %the observed periods are related 
@@
 Although of secondary importance, we can speculate on the phenomena underlying the change in variability before and after the great dimming event.
-We hypothesize that the great dimming event disrupted the dynamics of the photosphere, leading to the destruction of the largest granules by the turbulent motions caused by the non-equilibrium of the photosphere.
-Since this event, the variability of Betelgeuse has quickened in adequacy with the timescale of smaller convective cells, around 200 d.
+We hypothesize that the great dimming event disrupted the dynamics of the photosphere,
+leading to the destruction of the largest granules by the turbulent motions present in the non-equilibrium of the photosphere after such event.
+Since this happened, the variability of Betelgeuse has quickened in adequacy with the timescale of smaller convective cells, around 200 d.
 If this scenario is correct, we shall expect in the coming future, the re-emergence of the 400 d variability as the dominant period once the photosphere returns to some form of equilibrium.

### 2024-05-17 main.tex
test 6
test 7
fix typos

@@
 \section{Introduction}
+\label{sec:introduction}
 Betelgeuse is a prototypical red supergiant (RSG), known to be a semi-regular variable. Several periods can be found in the 
 literature, usually clustered around the so-called long secondary period (LSP) of 2000 d, plus shorter periods around 400 and 200 d \citep{kiss_variability_2006}.
-The origin of the LSP remains a source of discussion: some have tried to find the explanation of the LSP in the lifetime of the large convective cells present at the surface of RSG \citep[e.g][]{stothers_giant_2010} while others have invoked a magnetic field \citep{wood_long_2004}.
+The origin of the LSP remains a source of discussion: some have tried to find the explanation of the LSP in the lifetime of the large convective cells present at the surface of RSGs \citep[e.g][]{stothers_giant_2010} while others have invoked a magnetic field \citep{wood_long_2004}.
 The shorter periods of 400 and 200 d are given a different origin than the LSP.
@@
 
-Besides radial pulsation, surface activity leading to random brightness variations has also been mentioned to explain the variability of Betelgeuse, firstly theorised by \cite{schwarzschild_scale_1975}.
+Besides radial pulsation, surface activity leading to random brightness variations has also been mentioned to explain the variability of Betelgeuse, first theorised by \cite{schwarzschild_scale_1975}.
 More recently, \cite{gray_mass_2008} proposed that the 400 d period was a consequence of convective activity.
@@
 occasions, we have to add up the signals of thousands of lines to increase the signal-to-noise ratios. This addition 
-is performed through a technique called Least-Squares Deconvolution (LSD) \citep{donati_spectropolarimetric_1997}.% and assumes that the 
+is performed through a technique called Least-Squares Deconvolution \citep[LSD;][]{donati_spectropolarimetric_1997}.% and assumes that the 
 %linear polarization signal is similar 
@@
 magnetic field distributions over stellar surfaces and is now employed to produce images of the brightness distributions in the photosphere 
-of Betelgeuse and other RSGs, as cited in the Introduction.
+of Betelgeuse and other RSGs, as cited in section~\ref{sec:introduction}.
 
@@
 as the rest velocity of the star.
-Sometimes, signals at velocities greater than +40 km/s can be found, which can be interpreted either as dark and cold plasma falling back to Betelgeuse or as rising plumes of plasma located on the hidden face of Betelgeuse, ascending so high that they appear above the limb \citep[for the case of the RSG $\mu$Cep]{lopez_ariste_height_2023}.
+Sometimes, signals at velocities greater than +40 km/s can be found, which can be interpreted either as dark and cold plasma falling back to Betelgeuse or as rising plumes of plasma located on the hidden face of Betelgeuse, ascending so high that they appear above the limb \citep[for the case of the RSG $\mu$~Cep]{lopez_ariste_height_2023}.
 
@@
 RSGs experience mass loss events when plasma is ejected into the interstellar medium \citep{josselin_atmospheric_2007}.
-These events have been inferred in RSG such as $\mu$Cep at photospheric heights, where \cite{lopez_ariste_height_2023} found an excess of linear polarization beyond the limb of the star, attributing it to rising plumes of plasma in the back hemisphere of the star.
-In the case of Betelgeuse, the excess of linear polarization beyond the limb is rare and not as pronounced as those observed in $\mu$Cep. They do not appear to contribute to 
+These events have been inferred in RSG such as $\mu$~Cep at photospheric heights, where \cite{lopez_ariste_height_2023} found an excess of linear polarization beyond the limb of the star, attributing it to rising plumes of plasma in the back hemisphere of the star.
+In the case of Betelgeuse, the excess of linear polarization beyond the limb is rare and not as pronounced as those observed in $\mu$~Cep. They do not appear to contribute to 
  the variability of the star.

### 2024-05-17 main.tex
lay out TBL acronym on first use

@@
 
-In section \ref{Section 2}, we describe the dataset of linear polarization obtained with Narval and Neo-Narval at TBL, as well as the Lomb-Scargle periodograms used to derive periods.
+In section \ref{Section 2}, we describe the dataset of linear polarization obtained with Narval and Neo-Narval at the 2-metre T\'elescope Bernard Lyot (TBL), as well as the Lomb-Scargle periodograms used to derive periods.
 In section \ref{section 3}, we seek periods in the LSD profiles of Betelgeuse using the Lomb-Scarlge periodogram.

### 2024-05-17 main.tex
add citations for LSD and magnetic field

@@
 occasions, we have to add up the signals of thousands of lines to increase the signal-to-noise ratios. This addition 
-is performed through a technique called Least-Squares Deconvolution \citep[LSD;][]{donati_spectropolarimetric_1997}.% and assumes that the 
+is performed through the Least-Squares Deconvolution technique \citep[LSD;][]{donati_spectropolarimetric_1997,2010A&A...524A...5K}.% and assumes that the 
 %linear polarization signal is similar 
@@
 This technique has been successfully used in the past to measure 
-magnetic field distributions over stellar surfaces and is now employed to produce images of the brightness distributions in the photosphere 
+magnetic field distributions over stellar surfaces \citep[e.g][]{donatiLandstreet2009,reiners2012,kochukhov2021} and is now employed to produce images of the brightness distributions in the photosphere
 of Betelgeuse and other RSGs, as cited in section~\ref{sec:introduction}.

### 2024-05-17 main.tex
clarify inhomogenous (+bright) disk

@@
 the continuum photons being polarized by Rayleigh scattering.
-This depolarization produces signals with azimuthal symmetry over the visible disk, which would  cancel out the net linear polarization on a homogeneous disk.
-Thus, such mechanism of polarization must be combined with an inhomogeneous disk to produce the net linear polarization
+This depolarization produces signals with azimuthal symmetry over the visible disk, which would  cancel out the net linear polarization on a homogeneously bright disk.
+Thus, such mechanism of polarization must be combined with an inhomogeneously bright disk to produce the net linear polarization
 signal observed.

### 2024-05-17 main.tex
fix typos/grammar

@@
 %due to the presence of large quantities of dust.
-Since the end of the Great Dimming event, Betelgeuse has continue its random variation of brightness. 
+Since the end of the Great Dimming event, Betelgeuse has continued its random variation of brightness. 
 But quite interestingly, 
@@
 
-In this work, we shall examine the signals produced through the LSD technique, which produce a pseudo-spectral line in intensity 
+In this work, we examine the signals produced through the LSD technique, which produce a pseudo-spectral line in intensity 
 and linear polarization. This pseudo-spectral line does not correspond to any particular atomic species but rather represents an average of all species present and emitting in the 

### 2024-05-17 main.tex
soften tone: can conclude instead of must conclude

@@
 pulsations but linked to the brightness inhomogeneities due to the convective patterns in Betelgeuse. If one can find the aforementioned variability periods 
-in linear polarization, we must conclude that these periods are related to the convective activity which originates the linear polarization signals.
+in linear polarization, we can conclude that these periods are related to the convective activity which originates the linear polarization signals.
 %and unrelated to any radial pulsation phenomena. 

### 2024-05-17 main.tex
fix tense

@@
  gain validity through this comparison: they were evident in the TBL data, but barely recovered in the AAVSO data when pruned to the same observing dates.
-We can anticipated that extended spectropolarimetric observations
+We can anticipate that extended spectropolarimetric observations
  of Betelgeuse will make these periodicities seen in linear polarization, and arising from convective structures, clearer and clearer.

### 2024-05-27 main.tex
Add files via upload

@@
 %
+
+\def\kms {km\,s$^{-1}$ }
 \begin{document}
@@
   \abstract
+ % context heading (optional)
+  {Betelgeuse is a red supergiant (RSG) that is known to vary semi-regularly on both short and long timescales.}
+  % {} leave it empty if necessary  
+   {The origin of the short period of Betelgeuse has often been associated to radial pulsations but could also be due to the convection motions present at the surface of RSGs.
+Here, we investigate the link between surface activity and the variability of the star.
+}
+  % aims heading (mandatory)
+   {Linear polarization in Betelgeuse is a proxy of convection which is unrelated to pulsations.
+From the 10 years of spectropolarimetric data of Betelgeuse, we seek for periodicities in the least-squares deconvolution (LSD) profiles of Stokes $I$, $Q$, $U$ and the total linear polarization using the Lomb-Scargle periodogram.
+}
+  % methods heading (mandatory)
+   {We find that the 400 d period is present in the LSD profile of Stokes $I$ but is close to a peak of the window function of the TBL.
+Two periods of 330 d and 200 d are present in the periodogram of Stokes $Q$ and $U$, showing that we recover the variability of Betelgeuse using surface convection.
+We also speculate that change in variability of Betelgeuse after the Great Dimming is due to the non-equilibrium of the photosphere, destroying the large convective cells and leading to shorter timescales in the light curve of the star.
+}
+     % results heading (mandatory)
+   {}
+  % conclusions heading (optional), leave it empty if necessary
 
@@
 \section{Introduction}
-\label{sec:introduction}
+\label{sec: Introduction}
 Betelgeuse is a prototypical red supergiant (RSG), known to be a semi-regular variable. Several periods can be found in the 
@@
 
-Besides radial pulsation, surface activity leading to random brightness variations has also been mentioned to explain the variability of Betelgeuse, first theorised by \cite{schwarzschild_scale_1975}.
+Besides radial pulsation, surface activity leading to random brightness variation has also been mentioned to explain the variability of Betelgeuse, first theorised by \cite{schwarzschild_scale_1975}.
 More recently, \cite{gray_mass_2008} proposed that the 400 d period was a consequence of convective activity.
@@
 
-In section \ref{Section 2}, we describe the dataset of linear polarization obtained with Narval and Neo-Narval at the 2-metre T\'elescope Bernard Lyot (TBL), as well as the Lomb-Scargle periodograms used to derive periods.
+In section \ref{Section 2}, we describe the dataset of linear polarization obtained with Narval and Neo-Narval at the 2-meter Télescope Bernard Lyot (TBL), as well as the Lomb-Scargle periodograms used to derive periods.
 In section \ref{section 3}, we seek periods in the LSD profiles of Betelgeuse using the Lomb-Scarlge periodogram.
@@
 occasions, we have to add up the signals of thousands of lines to increase the signal-to-noise ratios. This addition 
-is performed through the Least-Squares Deconvolution technique \citep[LSD;][]{donati_spectropolarimetric_1997,2010A&A...524A...5K}.% and assumes that the 
+is performed through the Least-Squares Deconvolution (LSD) \citep[LSD;][]{donati_spectropolarimetric_1997,kochukhov_least-squares_2010}.% and assumes that the 
 %linear polarization signal is similar 
@@
 magnetic field distributions over stellar surfaces \citep[e.g][]{donatiLandstreet2009,reiners2012,kochukhov2021} and is now employed to produce images of the brightness distributions in the photosphere
-of Betelgeuse and other RSGs, as cited in section~\ref{sec:introduction}.
+of Betelgeuse and other RSGs, as cited in section~\ref{sec: Introduction}.
 
-In this work, we examine the signals produced through the LSD technique, which produce a pseudo-spectral line in intensity 
+In this work, we shall examine the signals produced through the LSD technique, which produce a pseudo-spectral line in intensity 
 and linear polarization. This pseudo-spectral line does not correspond to any particular atomic species but rather represents an average of all species present and emitting in the 
@@
 
-We first computed the Lomb-Scargle periodogram at each wavelength of the LSD profile, ranging from -25 km/s to +60 km/s to cover the entire 
+We first computed the Lomb-Scargle periodogram at each wavelength of the LSD profile, ranging from -25~\kms to +60~\kms with a step of 1.8~\kms to cover the entire 
 signal present in the LSD profile.
-\cite{lopez_ariste_convective_2018} identified the most blueshifted signal towards -20 km/s and interpreted it as  the maximum velocity of the rising plasma.
-The most redshifted signal was often found at +40 km/s, interpreted
+\cite{lopez_ariste_convective_2018} identified the most blueshifted signal towards -20~\kms and interpreted it as  the maximum velocity of the rising plasma.
+The most redshifted signal was often found at +40~\kms, interpreted
 as the rest velocity of the star.
-Sometimes, signals at velocities greater than +40 km/s can be found, which can be interpreted either as dark and cold plasma falling back to Betelgeuse or as rising plumes of plasma located on the hidden face of Betelgeuse, ascending so high that they appear above the limb \citep[for the case of the RSG $\mu$~Cep]{lopez_ariste_height_2023}.
+Sometimes, signals at velocities greater than +40~\kms can be found, which can be interpreted either as dark and cold plasma falling back to Betelgeuse or as rising plumes of plasma located on the hidden face of Betelgeuse, ascending so high that they appear above the limb \citep[for the case of the RSG $\mu$~Cep]{lopez_ariste_height_2023}.
 
@@
 %The abscissa is the frequency (in $\mathrm{days^{-1}}$), while the ordinate is the heliocentric radial velocity in km/s. 
-In each figure, the upper panel depicts the Lomb-Scargle periodogram of each velocity bin (grey lines) along with the total average (red line).
+In each figure, the upper panel depicts the Lomb-Scargle periodogram of each velocity bin (black line) along with the total average (red line).
 The green line represents the window function.
@@
 can be 
-rotated by $180^\circ$ degrees, and the brightness distribution will still fit the observed LSD polarization profiles. 
+rotated by $180^\circ$ degree, and the brightness distribution will still fit the observed LSD polarization profiles. 
 If each observation were treated independently, the algorithm's solution could be any of the possible ambiguous solutions. To ensure continuity 
@@
     \includegraphics[width=0.5\textwidth]{Lomb-Scargle Photo-center.pdf}
-    \caption{Lomb-Scargle periodogram of the 100 photo-center displacement series of Betelgeuse. Each blue line corresponds to a Lomb-Scargle periodogram of one photo-center
-     displacement. The black line is the average of the 100 periodograms. The red dashed lines represent the 400 and 200 d period respectively. The green line is the window function.}
+    \caption{Lomb-Scargle periodogram of the 100 photo-center displacement series of Betelgeuse. Each blue lines correspond to a Lomb-Scargle periodogram of one photo-center
+     displacement. The black line is the average of the 100 periodogram. The red dashed lines represent the 400 and 200 d period respectively.
+     The green line is the window function. The black horizontal line marks the 2$\sigma$ confidence level.}
     \label{LS photocenter}
@@
 200 d periods, respectively.
+The black horizontal line represents the 2$\sigma$ confidence level.
 Similar to previous figures, the window function is represented by the green line.
-Interestingly, we find two peaks around the 400 d period in the periodogram, although they are not individually significant to confirm the presence of such period.
-However, these peaks are consistent with the findings of the previous section.
+Interestingly, we find two peaks around the 400 d period in the periodogram, one at 0.0024 $\mathrm{d^{-1}}$ and one at 0.003 $\mathrm{d^{-1}}$ which are above the 2$\sigma$ confidence level.
+Although they are not individually significant to confirm the presence of such period, these peaks are consistent with the findings of the previous section.
 Since polarimetry imaging involves only surface convection, this provides further support for a variability being explained by surface convection alone.
@@
 However, after the dimming, this period seems to have vanished, and only timescales shorter than 230 d are visible since \citep{dupree_great_2022}.
-No explanations have been put forth regarding this change of variability after, or perhaps because of the dimming.% In  Section \ref{section 4},  we  afford proposing alternative scenarios for the observed periodicities in the light curve.\\
+No explanations have been put forth regarding this change of variability after, or perhaps because of the dimming. \
 
- In Fig.\ref{Lomb Scargle AAVSO}, we computed the periodogram from that light curve of Betelgeuse (blue line), alongside the periodogram computed exclusively with AAVSO observations made less than 2 days from our observation dates of the TBL (orange line).
+In Fig.\ref{Lomb Scargle AAVSO}, we computed the periodogram from the light curve of Betelgeuse from 1990 to 2024 (blue line), alongside the periodogram computed exclusively with AAVSO observations made less than 5 days from our observation dates of the TBL (orange line).
 For the periodogram since 1990, we binned the observations with an interval of 10 days, as Betelgeuse has been more observed in the 21st century \citep{kiss_variability_2006}.
@@
 However, when focusing on the periodogram derived from the light curve corresponding to the TBL observation dates, we fail to retrieve the periods mentioned in the literature.
-The periodicities found in the TBL observations
- gain validity through this comparison: they were evident in the TBL data, but barely recovered in the AAVSO data when pruned to the same observing dates.
+The periodicities found in the TBL observations gain validity through this comparison: they were evident in the TBL data, but barely recovered in the AAVSO data when pruned to the same observing dates.
 We can anticipate that extended spectropolarimetric observations
- of Betelgeuse will make these periodicities seen in linear polarization, and arising from convective structures, clearer and clearer.
+of Betelgeuse will make these periodicities seen in linear polarization. Due to convective structures, these periodicities will be clearer and clearer.
 
@@
     \includegraphics[width=0.5\textwidth]{Lomb-Scargle AAVSO.pdf}
-    \caption{Lomb-Scargle periodogram of the light curve of Betelgeuse since 1990 as retrieved from the AAVSO database (blue line). When only dates corresponding to TBL observations are retained in 
-    that light curve, the periodogram in orange color is obtaine. The two black dashed lines are respectively the 400 and 200 d periods for visual reference.}
+    \caption{Lomb-Scargle periodogram of the light curve of Betelgeuse since 1990 as retrived from the AAVSO database (blue line).
+When only dates corresponding to TBL observations are retained in that light curve, the periodogram in orange color is obtained.
+The two black dashed lines are respectively the 400 and 200 d periods for visual reference.}
     \label{Lomb Scargle AAVSO}
@@
 of convective cells across the visible hemisphere. Therefore, the 200 d period appears to be related to the evolution of the convective patterns 
-on the surface of Betelgeuse.  The stronger presence of this peak in Stokes $Q$ compared to $U$ could be interpreted as a temporal situation of how convective patterns have been distributed in the
+on the surface of Betelgeuse.
+The stronger presence of this peak in Stokes $Q$ compared to $U$ was already pointed out by \cite{mathias_evolution_2018}, and could be interpreted as a temporal situation of how convective patterns have been distributed in the
 last years.\
@@
 Once again, we observe the 330 d periodicity to be more prominent in Stokes $U$ than in $Q$, what may unravel a random situation due to convective motions in the last years.
-This interpretation aligns with the work of \cite{gray_mass_2008}, who attributed the 400 d period to
-large convective cells.
+Convective cells could arise in peculiar positions, where the Stokes $U$ profile remains unchanged for month but where Stokes $Q$ could change on shorter timescales.
+This interpretation aligns with the work of \cite{gray_mass_2008}, who attributed the 400 d period to large convective cells.
 Whether this 330 d period and the commonly cited 400 d period are one and the same will have to wait for extended spectropolarimetric observations with the TBL that smooth the windowing function around these frequencies.
+\
 
-Regarding the Stokes $I$ profile, a peak closer to the 400 d period appears, but it is once again uncomfortably close 
-to a peak of the window function of the TBL, making it difficult to trust.
-A secondary peak close to 200 d is also present in the Stokes $I$ profile, that we fully assimilate to the one found in the Stokes parameters, and that  we attribute to convective timescales.
-%Up to this point, the model developed to interpret linear polarization involves only surface convection, which is why we interpret the 200 d period as the convective timescale of the smallest structures, while the 400 d period corresponds to the timescale of the largest structures.\\
+Regarding the Stokes $I$ profile, a peak closer to the 400 d period appears, but is once again uncomfortably close to a peak of the window function of the TBL, making it difficult to trust.
+A secondary peak close to 200 d is also present in the Stokes $I$ profile, that we fully assimilate to the one found in the Stokes parameters, and that we attribute to convective timescales.\\
  %As showed by
@@
 Before summing up the results of our work, we can afford to propose an explanation for the variability of Betelgeuse after the great dimming with surface convection.
-RSGs experience mass loss events when plasma is ejected into the interstellar medium \citep{josselin_atmospheric_2007}.
+RSGs experience mass loss event, where plasma is ejected into the interstellar medium \citep{josselin_atmospheric_2007}.
 These events have been inferred in RSG such as $\mu$~Cep at photospheric heights, where \cite{lopez_ariste_height_2023} found an excess of linear polarization beyond the limb of the star, attributing it to rising plumes of plasma in the back hemisphere of the star.
-In the case of Betelgeuse, the excess of linear polarization beyond the limb is rare and not as pronounced as those observed in $\mu$~Cep. They do not appear to contribute to 
- the variability of the star.
+In the case of Betelgeuse, the excess of linear polarization beyond the limb is rare and not as pronounced as those observed in $\mu$~Cep.
+They do not appear to contribute to the variability of the star.
 However, since the great dimming, the variability of Betelgeuse has changed.
-This suggests that the great dimming differed significantly from the usual mass loss event, and perhaps it hustled the  photosphere.
+This suggests that the great dimming differed significantly from the usual mass loss event, that it hustled the dynamics of the photosphere.
 \
 
-In a scenario where the variability of Betelgeuse is primarily governed by convection,  the visible periods of such variability  were associated, before the great dimming, to the 
-typical timescales at which convection occurs: approximately 200 d for the small structures and roughly 350 d for the large structures.
-The great dimming affected the behaviour of the photosphere in such a way that the largest structures
-no longer evolved on a timescale around 350 d, but rather on the shorter timescales of the smallest convective structures, around 200 d.
-We hypothesise that after the great dimming event, the photosphere has not yet returned to equilibrium, and the largest convective structures
-have been disrupted by the turbulent motions present of the photosphere in this period.
-This would explain why identifying the longer periods after the great dimming has been hard, as \cite{dupree_great_2022} noticed.
-We may expect that the 400 d
-period will gradually reappear in the coming years as the photosphere returns to equilibrium. However speculative, this scenario 
-provides a rough explanation of how convective activity could lead to a change in variability since the great dimming that future observations and 
-research will confirm or infirm.
-%A more detailed explanation of this variability change falls beyond the scope of this paper and is left for future research.
+In a scenario where the variability of Betelgeuse is primarily governed by convection, the visible periods of such variability were associated, before the great dimming, to the typical timescales at which convection occurs: approximately 200 d for the small structures and roughly 350 d for the large structures.
+The great dimming affected the behaviour of the photosphere in
+such a way that the largest structures no longer evolved on a
+timescale around 350 d, but rather on the shorter timescales of
+the smallest convective structures, around 200 d. We hypothesise that after the great dimming event, the photosphere has not
+yet returned to equilibrium, and the largest convective structures
+have been disrupted by the turbulent motions present of the pho-
+tosphere in this period. This would explain why identifying the
+longer periods after the great dimming has been hard, as \cite{dupree_great_2022} noticed. We may expect that the 400 d period will
+gradually reappear in the coming years as the photosphere returns to equilibrium. However speculative, this scenario provides
+a rough explanation of how convective activity could lead to a
+change in variability since the great dimming that future observations and research will confirm or infirm.
 %But this 
@@
 We propose that the true reason for the observed periodicities in both the light curve and spectropolarimetric observations lies within these convective structures and their temporal evolution.
-This is the main conclusion of our work: since we are able to recover the different periods in the linear polarization profiles, we can exclude pulsation phenomena as the explanation for those periodicities in the light curve.
+This is the main conclusion of our work: since we are able to recover the different periods in the linear polarization profiles, we can exclude pulsation phenomena as the explanation for those periodicities in the
+light curve.
 This conclusion concurs with the scenario proposed by \cite{gray_mass_2008}, attributing the variability of Betelgeuse to surface convection.
- We further find hints in the polarized spectra that lead us to associate the 200 d periodicity to convective dynamics, while the 400 d periodicity may be linked to convective timescales of the largest granules.
+We further find hints in the polarized spectra that lead us to associate the 200 d periodicity to convective dynamics, while the 400 d periodicty may be linked to convective timescales of the largest granules.
 %the observed periods are related 
@@
 Although of secondary importance, we can speculate on the phenomena underlying the change in variability before and after the great dimming event.
-We hypothesize that the great dimming event disrupted the dynamics of the photosphere,
-leading to the destruction of the largest granules by the turbulent motions present in the non-equilibrium of the photosphere after such event.
+We hypothesize that the great dimming event disrupted the dynamics of the photosphere, leading to the destruction of the largest granules by the turbulent motions present in the non-equilibrium of the photosphere after such event.
 Since this happened, the variability of Betelgeuse has quickened in adequacy with the timescale of smaller convective cells, around 200 d.

### 2024-06-04 main.tex
Correct IRAP affiliation
Last modif

@@
 
-   \institute{IRAP, Universit\'e de Toulouse, CNRS, CNES, UPS.  14, Av. E. Belin. 31400 Toulouse, France 
+   \institute{
+            Institut de Recherche en Astrophysique et Plan\'etologie,
+            Universit\'e de Toulouse, CNRS, IRAP/UMR 5277,
+            14 avenue Edouard Belin, F-31400, Toulouse, France  
              }
@@
  % context heading (optional)
-  {Betelgeuse is a red supergiant (RSG) that is known to vary semi-regularly on both short and long timescales.}
+  {Betelgeuse is a red supergiant (RSG) that is known to vary semi-regularly on both short and long timescales.
+The origin of the short period of Betelgeuse has often been associated to radial pulsations but could also be due to the convection motions present at the surface of RSGs.}
   % {} leave it empty if necessary  
-   {The origin of the short period of Betelgeuse has often been associated to radial pulsations but could also be due to the convection motions present at the surface of RSGs.
-Here, we investigate the link between surface activity and the variability of the star.
-}
+   { We investigate the link between surface activity and the variability of the star, ignoring pulsations.   }
   % aims heading (mandatory)
    {Linear polarization in Betelgeuse is a proxy of convection which is unrelated to pulsations.
-From the 10 years of spectropolarimetric data of Betelgeuse, we seek for periodicities in the least-squares deconvolution (LSD) profiles of Stokes $I$, $Q$, $U$ and the total linear polarization using the Lomb-Scargle periodogram.
+Using  10 years of spectropolarimetric data of Betelgeuse, we seek for periodicities in the least-squares deconvolution (LSD) profiles of Stokes $I$, $Q$, $U$ and the total linear polarization using  Lomb-Scargle periodograms.
 }
   % methods heading (mandatory)
-   {We find that the 400 d period is present in the LSD profile of Stokes $I$ but is close to a peak of the window function of the TBL.
-Two periods of 330 d and 200 d are present in the periodogram of Stokes $Q$ and $U$, showing that we recover the variability of Betelgeuse using surface convection.
-We also speculate that change in variability of Betelgeuse after the Great Dimming is due to the non-equilibrium of the photosphere, destroying the large convective cells and leading to shorter timescales in the light curve of the star.
+   {We find similar periods in linear polarization signals than in photometric variability.
+   The 400 d period is too close to a peak of the window function of our data.
+But the two periods of 330 d and 200 d are present in the periodogram of Stokes $Q$ and $U$, showing that  the variability of Betelgeuse can be interpreted as due to  surface convection.
 }
      % results heading (mandatory)
-   {}
+   {Since linear polarization in the spectrum of Betelgeuse is not known to vary with pulsations, but is linked to surface convection, and 
+   since similar periods are found in time series of photometric measurements and spectropolarimetry, we conclude that the photometric 
+   variability is due to the surface convective structures, and not to any pulsation phenomenon.}
   % conclusions heading (optional), leave it empty if necessary

### 2024-09-04 main.tex
Add files via upload

@@
 
-    \author{{ Q.~Pilate}\inst{1},{ A.~L{\'o}pez Ariste}\inst{1},{ A.~Lavail}\inst{1},{ Ph. Mathias}\inst{1} }
+    \author{{ Q.~Pilate}\inst{1},{ A.~López Ariste}\inst{1},{ A.~Lavail}\inst{1},{ Ph. Mathias}\inst{1} }
 
@@
      % results heading (mandatory)
-   {Since linear polarization in the spectrum of Betelgeuse is not known to vary with pulsations, but is linked to surface convection, and 
-   since similar periods are found in time series of photometric measurements and spectropolarimetry, we conclude that the photometric 
-   variability is due to the surface convective structures, and not to any pulsation phenomenon.}
+   {Since linear polarization in the spectrum of Betelgeuse is not known to vary with pulsations, but is linked to surface convection, and since similar periods are found in time series of photometric measurements and spectropolarimetry, we conclude that the photometric variability is due to the surface convective structures, and not to any pulsation phenomenon.}
   % conclusions heading (optional), leave it empty if necessary
@@
 The origin of the LSP remains a source of discussion: some have tried to find the explanation of the LSP in the lifetime of the large convective cells present at the surface of RSGs \citep[e.g][]{stothers_giant_2010} while others have invoked a magnetic field \citep{wood_long_2004}.
-The shorter periods of 400 and 200 d are given a different origin than the LSP.
-By studying a large sample of RSGs, \cite{kiss_variability_2006} attributed the 400 d period to radial pulsation modes, while the 200 d period being attributed to the first overtone \citep{joyce_standing_2020}, in agreement with the historical work of \cite{stothers_pulsation_1969}.\\
+\textbf{The shorter periods of 400 and 200 d are given a different origin than the LSP, and are traditionally attributed to radial pulsation driven by the $\kappa$-mechanism, the 400 d period  being the fundamental mode and the 200 d the first overtone.
+By studying a large sample of RSGs, \cite{kiss_variability_2006} found a period of 388 $\pm$ 30 d, which was later confirmed by \cite{chatys_periodluminosity_2019}.
+Further works have been carried out to identify periods in Betelgeuse, for instance, using the observations from the AAVSO database between 1995 and 2014, \cite{montarges_close_2016} found a period of 423.59 $\pm$ 39 d for observations before January 2012 and a period of 422.01 $\pm$ 88 d for observations after January 2012.
+Furthermore, \cite{jadlovsky_analysis_2023} found a photometric period of 417 $\pm$ 17 d, while \cite{joyce_standing_2020} found a period of 416 $\pm$ 24 d and identified another shorter period of 185 $\pm$ 13.5 d being attributed to the first overtone, everything in agreement with the historical work of \cite{stothers_pulsation_1969}.}\\
 
@@
 within the span of one week.
-Bright convection cells near the disk center will increase the integrated brightness of the star compared with other situations where such cells are found near the edges.
+\textbf{Those typical convective timescales are also seen in numerical simulations of RSGs, where the smallest convective granules live on a timescale of a few weeks/month and the largest on a timescale of a few months up to a few years \citep{chiavassa_probing_2022}.} Bright convection cells near the disk center will increase the integrated brightness of the star compared with other situations where such cells are found near the edges.
 So, even without invoking the formation of
@@
 convective patterns \citep{gray_mass_2008}. \\
-
-%more or less well defined 
-%periods around 400 and 200 days. 
-%Such periods are often linked to radial pulsation modes,   and have been used to model 
-%the characteristics of the resonant cavity, hence the size of the star and eventually its evolutionary stage \citep{saio_evolutionary_2023}.
 
@@
 brightness distribution over the stellar disk.
-%due to the presence of large quantities of dust.
 Since the end of the Great Dimming event, Betelgeuse has continued its random variation of brightness. 
@@
 spectra.\\
-%but unrelated to pulsations and which cannot be directly linked to brightness.
-%Faced with the dramatic change in stellar parameters and age that such change in a pulsation period would 
-%require,
 
@@
 in linear polarization, we can conclude that these periods are related to the convective activity which originates the linear polarization signals.
-%and unrelated to any radial pulsation phenomena. 
 This is the purpose of the present work.\
@@
 In section \ref{Section 2}, we describe the dataset of linear polarization obtained with Narval and Neo-Narval at the 2-meter Télescope Bernard Lyot (TBL), as well as the Lomb-Scargle periodograms used to derive periods.
-In section \ref{section 3}, we seek periods in the LSD profiles of Betelgeuse using the Lomb-Scarlge periodogram.
+In section \ref{section 3}, we seek periods in the \textbf{Least-Squares Deconvolution (LSD)} profiles of Betelgeuse using the \textbf{Lomb-Scargle} periodogram.
 In section \ref{section 4}, we associate the 200 d period with the convective timescale of the smallest granules and the 330 d periodicity with the largest granules.
@@
 
-Betelgeuse has been observed for the last 10 years with Narval and Neo-Narval at the Telescope Bernard Lyot\footnote[1]{https://tbl.omp.eu/}. These two instruments measure the polarization 
+Betelgeuse has been observed \textbf{since 2013} with Narval and Neo-Narval at the \textbf{Télescope} Bernard Lyot\footnote[1]{https://tbl.omp.eu/}.
+\textbf{ The polarimeter Narval observed Betelgeuse on average every month from November 2013 to April 2019, except during summer, representing a total of 56 observations over 5.5 years.
+During this period, Betelgeuse was observed 43\% of the time between January and April and 57\% between September and December.
+Before the great dimming of Betelgeuse, Narval was replaced by Neo-Narval which is now observing Betelgeuse since 2020.
+} These two instruments measure the polarization
 over the visible and near infrared spectra of Betelgeuese (390-1000 nm) with high spectral resolution (R=65000) and high polarimetric sensitivity.
@@
 occasions, we have to add up the signals of thousands of lines to increase the signal-to-noise ratios. This addition 
-is performed through the Least-Squares Deconvolution (LSD) \citep[LSD;][]{donati_spectropolarimetric_1997,kochukhov_least-squares_2010}.% and assumes that the 
-%linear polarization signal is similar 
-%in all spectral lines up to scale factors in the sampling and amplitude.
+is performed through the Least-Squares Deconvolution \citep[LSD;][]{donati_spectropolarimetric_1997,kochukhov_least-squares_2010}.
 This technique has been successfully used in the past to measure 
@@
 of Betelgeuse and other RSGs, as cited in section~\ref{sec: Introduction}.
+\textbf{ To produce the LSD profiles, we used a solar abundance line mask, similar to the one used by \cite{auriere_discovery_2016} and \cite{mathias_evolution_2018}, which was calculated from the data provided by VALD \citep{kupka_vald2_1999}, for an effective temperature of 3750 K, $\log g =0.0$ and a microturbulence of 4 \kms, aligning with the physical paramters of \cite{josselin_atmospheric_2007}.
+The mask includes approximately 15,000 lines with depths exceeding 40\% of the continuum.}
 
@@
 
-%\subsection{Lomb-Scargle periodograms of the profiles}
-
 \section{Period search}
@@
 we used the Lomb-Scargle periodogram \citep{lomb_least-squares_1976,scargle_studies_1982}.
-This method involves fitting sinusoids to the data using least-squares with frequency sampled between the first and last observation.
+This method involves fitting sinusoids to the data using least-squares with frequency sampled between the first and last observation \textbf{\citep{the_astropy_collaboration_astropy_2018}}.
 The quality of the fit determines the power attributed to a frequency.
-We applied this technique to the  LSD profiles of Stokes $I$, $U$, $Q$ and to the total linear polarization of Betelgeuse observed by the TBL over a span of 10 years (2014-2024).
+We applied this technique to the  LSD profiles of Stokes $I$, $U$, $Q$ and to the total linear polarization of Betelgeuse observed by the TBL over a span of 10 years (\textbf{2013}-2024).
 However, the dataset available post-dimming is insufficient to produce meaningful periodograms.
-Therefore, our focus was on periodograns computed exclusively from data obtained before the great dimming, covering the period from 2014 to the end of 2019.
+Therefore, our focus was on periodograns computed exclusively from data obtained before the great dimming, covering the period from \textbf{ November 2013} to \textbf{April} 2019.
+\textbf{We only used data obtained from Narval, with a total of 56 observations}.
 
@@
 as the rest velocity of the star.
-Sometimes, signals at velocities greater than +40~\kms can be found, which can be interpreted either as dark and cold plasma falling back to Betelgeuse or as rising plumes of plasma located on the hidden face of Betelgeuse, ascending so high that they appear above the limb \citep[for the case of the RSG $\mu$~Cep]{lopez_ariste_height_2023}.
-
-% The LS periodogram  computed for each frequency 
-% in the LSD profile must show periodicities around the LSP, but also on the 400 days and 200 days period. If the LS periodogram finds a periodicity at a
-% given wavelength, it would appear as a strong LS power. However, we expect a change in the periodicity after the dimming. 
-% This would lead to a decrease in the Lomb-Scargle power. By comparing LS periodogram before and after the dimming, we expect to see a change in the LS power,
-% especially at the 400 days period.
+Sometimes, signals at velocities greater than +40~\kms can be found, \textbf{with peak in the linear polarization signal up to 50~\kms \citep{lopez_ariste_three-dimensional_2022}}, which can be interpreted either as dark and cold plasma falling back to Betelgeuse or as rising plumes of plasma located on the hidden face of Betelgeuse, ascending so high that they appear above the limb \citep[for the case of the RSG $\mu$~Cep]{lopez_ariste_height_2023}.
+\textbf{The +60~\kms cut off in the Lomb-Scargle periodograms was set to include the occasional linear polarization peaks beyond the velocity of the star.}
 
@@
 %The abscissa is the frequency (in $\mathrm{days^{-1}}$), while the ordinate is the heliocentric radial velocity in km/s. 
-In each figure, the upper panel depicts the Lomb-Scargle periodogram of each velocity bin (black line) along with the total average (red line).
+In each figure, the upper panel depicts the Lomb-Scargle periodogram of each velocity bin (\textbf{gray lines}) along with the total average (red line).
 The green line represents the window function.
-The lower panel shows the Lomb-Scargle periodogram for each velocity bin, with the white dotted lines marking the 400 and 200 d periods for reference.
+\textbf{The spectral window reﬂects the pattern caused by the structure of gaps in the time string.
+The peaks pointed out by this window cannot be interpreted as ones related to the star.} The lower panel shows the Lomb-Scargle periodogram for each velocity bin, with the white dotted lines marking the 400 and 200 d periods for reference.
 The right panel represents the mean profile of each Stokes parameter at each velocity bin.
@@
     The upper panel is the Lomb-Scargle periodogram for each velocity bin (grey lines) and the average (red line). The green line is the window function.
-    In the lower panel, the two white dashed lines mark respectively the 400 d and the 200 d periods. The right panel is the average intensity profile for each velocity bin.}
+    In the lower panel, the two white dashed lines mark respectively the 400 d and the 200 d periods.
+\textbf{The y-axis represents the heliocentric radial velocity (HRV).
+The color axis marks the power of the fit, red shows a higher power than dark blue.} The right panel is the average intensity profile for each velocity bin.}
     \label{LS intensity}
@@
 
-%\section{Recovering the variability of Betelgeuse in the LSD profile}
-
-%\subsection{Intensity}
 Examining Fig.\ref{LS intensity}, which shows the Lomb-Scargle periodogram of the LSD of the intensity profile, we observe that the 400 and 200 d periods seem to be captured by the periodogram.
@@
 Regarding Stokes $U$ in Fig.
-\ref{LS U}, it appears that the primary period is approximately $0.003 \ \mathrm{d^{-1}}$ (equivalent to 330 d).
+\ref{LS U}, it appears that the primary \textbf{frequency} is approximately $0.003 \ \mathrm{d^{-1}}$ (equivalent to 330 d \textbf{period}).
 Other periods, such as those at 200 d or 250 d ($0.004 \ \mathrm{d^{-1}}$) are present, but are difficult to trust.
@@
 This periodogram confirms the significant powers at both 330 d and 200 d, consistent with the periodograms of Stokes $Q$ and $U$.
-
-%Nonetheless, the power of this period peaks at a HRV of around 40 km/s. We recall that this velocity is the rest velocity of the star. 
-%Every signal beyond this velocity is either due to  plasma sinking in Betelgeuse, or to plumes of plasma rising beyond and above the limb of the star. 
-%Such plumes from the back hemisphere have been often seen in other red supergiants like $\mu$ Cep \citep{lopez_ariste_height_2023}. 
-%This suggests that both signals at these period of 400 days found in intensity and total linear polarisation 
-%could well be due to the periodic appearance of plumes in the back hemisphere rising high enough to appear above
-%the visible limb and generating a supplementary source of brightness. Once again we find a plausible explanation 
-%of these periods unrelated to pulsations, but rather to the convective dynamics of the star.
-
-% Up to this point, we want to pinpoint the two elements that are a clue 
-% to understand the variability of Betelgeuse: the 400 days period is present in the linear polarization signal and also at a HRV above 40 km/s.
-% In any case, from the linear polarization and the intensity profile, we recover the different periods of Betelgeuse. Since linear polarization is 
-% associated to convection and because we recover the periods of Betelgeuse, the main mechanism behind the variability of Betelgeuse points toward surface convection.
 
@@
 This technique has previously allowed for the estimation of  the size and the
- lifetime of convective cells on the surface \citep{lopez_ariste_convective_2018}. From these images, we computed a photo-center, a quantity sensitive to the size and number of convective cells. 
+ lifetime of convective cells on the surface \citep{lopez_ariste_convective_2018}.
+From these images, we computed a photo-center, a quantity sensitive to the size and number of convective cells.
+\textbf{The photo-center was computed from the equations given by \cite{chiavassa_probing_2022}.}
 A homogeneous star will have a photo-center displacement coinciding with the barycenter of the star, whereas a star with one or two large convective cells will exhibit a more siginificant photo-center displacement, up to a few percent of the stellar radius in the case of RSGs \citep{chiavassa_probing_2022}.
@@
 While these periods may overlap with those presented in the previous section, they are likely to capture additional aspects of the phenomena at work.
-
-%We computed the Lomb-Scargle periodogram of the displacement of the photo-center.
 However, before proceeding to search for periods in the displacement of the photo-center, it is important to address a key issue regarding the interpretation of such images. 
@@
      displacement. The black line is the average of the 100 periodogram. The red dashed lines represent the 400 and 200 d period respectively.
-     The green line is the window function. The black horizontal line marks the 2$\sigma$ confidence level.}
+     The green line is the window function. The black horizontal line marks the \textbf{1}$\sigma$ confidence level.}
     \label{LS photocenter}
@@
 Figure \ref{LS photocenter} shows the Lomb-Scargle periodogram of the photo-center displacement for each of the 100 series (blue lines) and the average Lomb-Scargle periodogram (black line).
-%As before, the upper panel shows the Lomb-Scargle periodogram from data before the great dimming only, whereas we used the whole set of data for the lower panel.
 The red dashed lines indicate the 400 and
 200 d periods, respectively.
-The black horizontal line represents the 2$\sigma$ confidence level.
+The black horizontal line represents the \textbf{1}$\sigma$ confidence level\textbf{, which was computed from the standard deviation of the average Lomb-Scarlge periodogram}.
 Similar to previous figures, the window function is represented by the green line.
-Interestingly, we find two peaks around the 400 d period in the periodogram, one at 0.0024 $\mathrm{d^{-1}}$ and one at 0.003 $\mathrm{d^{-1}}$ which are above the 2$\sigma$ confidence level.
+Interestingly, we find two peaks around the 400 d period in the periodogram, one at 0.0024 $\mathrm{d^{-1}}$ and one at 0.003 $\mathrm{d^{-1}}$ which are above the \textbf{1}$\sigma$ confidence level.
 Although they are not individually significant to confirm the presence of such period, these peaks are consistent with the findings of the previous section.
@@
 
-In Fig.\ref{Lomb Scargle AAVSO}, we computed the periodogram from the light curve of Betelgeuse from 1990 to 2024 (blue line), alongside the periodogram computed exclusively with AAVSO observations made less than 5 days from our observation dates of the TBL (orange line).
-For the periodogram since 1990, we binned the observations with an interval of 10 days, as Betelgeuse has been more observed in the 21st century \citep{kiss_variability_2006}.
+In Fig.\ref{Lomb Scargle AAVSO}, \textbf{the upper panel is the periodogram of the mean intensity profile (red line) shown in Fig.\ref{LS intensity}.
+In the middle panel, we computed the periodogram from the light curve of Betelgeuse from 1990 to 2024, represented with the blue line.
+The orange line in lower panel depicts the periodogram computed exclusively with AAVSO observations made less than 2 days from our observation dates of the TBL.
+In every panels, the green line represents the window function.} For the periodogram since 1990, we binned the observations with an interval of 10 days, as Betelgeuse has been more observed in the 21st century \citep{kiss_variability_2006}.
 This periodogram  recovers the 400 d period and also a period close to 200 d.
@@
 due to the windowing effect.
-However, when focusing on the periodogram derived from the light curve corresponding to the TBL observation dates, we fail to retrieve the periods mentioned in the literature.
-The periodicities found in the TBL observations gain validity through this comparison: they were evident in the TBL data, but barely recovered in the AAVSO data when pruned to the same observing dates.
-We can anticipate that extended spectropolarimetric observations
-of Betelgeuse will make these periodicities seen in linear polarization. Due to convective structures, these periodicities will be clearer and clearer.
+\textbf{When focusing on the periodogram derived from the light curve corresponding to the TBL observation dates (lower panel), we recover the 400 d period, which is once again, very close to a peak of the window function.
+A smaller peak is also present around 200 d.
+The comparison between this periodogram and the one obtained from Stokes $I$ is important to validate our results.
+Stokes $I$ is linked to the total amount of light, that is the light curve.
+Hence, by taking observations of the AAVSO that correspond to the TBL observation dates of Betelgeuse, we should recover the same periods that the ones found in the Stokes $I$ periodogram.
+Since we found the 400 d and the 200 d periods in both periodograms, this validates our results.
+In both cases, the peak at 400 d is too close to a peak of the window function to be interpreted as a real peak.
+On the other hand, the second peak around 200 d, could be interpreted as a real peak.
+In the future, extended spectropolarimetric observations of Betelgeuse will be necessary to smooth the window function of the TBL and perhaps, disentangle the 400 d period from the window function.}
 
@@
     \includegraphics[width=0.5\textwidth]{Lomb-Scargle AAVSO.pdf}
-    \caption{Lomb-Scargle periodogram of the light curve of Betelgeuse since 1990 as retrived from the AAVSO database (blue line).
-When only dates corresponding to TBL observations are retained in that light curve, the periodogram in orange color is obtained.
-The two black dashed lines are respectively the 400 and 200 d periods for visual reference.}
+    \caption{\textbf{Upper panel: Lomb-Scargle periodogram of the mean intensity profile presented in Fig.\ref{LS intensity} (red line).
+    Middle panel: Periodogram of the light curve since 1990 as retrived from the AAVSO database (blue line).
+    Lower panel: the orange line represents the periodogram of the light curve from the AAVSO when the dates correspond to the TBL observations.
+In every panels, the green line marks the associated window function.
+}}
     \label{Lomb Scargle AAVSO}
@@
 \label{section 4}
-%Before summing up the conclusions of this work, we can afford proposing alternative scenarios for the observed periodicities in the light curve, 
-%the LSD profiles and the displacement of the photo-center of the inferred photospheric image reconstructions. 
-The presence of the same 
-periods and, considering the time span and sparsity of the TBL data series, the larger amplitude of the peaks found in
-polarization data when compared to those in the light curve point to a common origin in either observation, one that is rather based upon convective dynamics.
-Since we associate Stokes $Q$ and $U$ to proxies of convection, we associate the 200 d period to the typical timescales on which the Stokes parameters evolve.
+
+\textbf{The presence of the same 
+periods in the peridogram derived from the AAVSO and the ones obtained from linear polarization and, considering the time span and sparsity of the TBL data series point to a common origin, one that is rather based upon convective dynamics.} Since we associate Stokes $Q$ and $U$ to proxies of convection, we associate the 200 d period to the typical timescales on which the Stokes parameters evolve.
 The LSD profiles of Stokes $Q$ or $U$ undergo slight changes when observing Betelgeuse monthly, but are completely overhauled 
@@
 This characteristic timescale has also been observed in the numerical simulations and might play an important role in determining stellar distances through the displacement of the photo-center \citep{chiavassa_probing_2022}.
-Once again, we observe the 330 d periodicity to be more prominent in Stokes $U$ than in $Q$, what may unravel a random situation due to convective motions in the last years.
-Convective cells could arise in peculiar positions, where the Stokes $U$ profile remains unchanged for month but where Stokes $Q$ could change on shorter timescales.
+Once again, we observe the 330 d periodicity to be more prominent in Stokes $U$ than in $Q$, \textbf{which} may unravel a random situation due to convective motions in the last years.
+Convective cells could arise in peculiar positions, where the Stokes $U$ profile remains unchanged for \textbf{months} but where Stokes $Q$ could change on shorter timescales.
 This interpretation aligns with the work of \cite{gray_mass_2008}, who attributed the 400 d period to large convective cells.
@@
 A secondary peak close to 200 d is also present in the Stokes $I$ profile, that we fully assimilate to the one found in the Stokes parameters, and that we attribute to convective timescales.\\
- %As showed by
-%\cite{lopez_ariste_convective_2018}, some of those convective cells can live up for years. We may speculate
-%that the LSP may be related to those long living structures. However the present span of data from the TBL is insufficient to do more than
-%just speculate. It should be noticed that  
-% the peak is not present in stokes U, and this may point to a more stochastic event than just long-living structures.
 
@@
 change in variability since the great dimming that future observations and research will confirm or infirm.
-%But this 
-%is a behaviour that fully concurs with what we find in the polarization data, and which cannot be related to any stellar pulsation but rather 
-%convective dynamics. Beyond the actual values of the periods found in a Lomb-Scargle periodogram, we also see a correspondence between polarization 
-%and the light curve in these changes in the observed periods before and after the great dimming. Such concurrence points to an explanation 
-%in terms of convective dynamics also for the variability of the light curve.
-
-%It may be argued that the periodograms presented in the previous sections and, in particular, those  of the photo-center displacements, 
-%present low amplitude peaks at the referred main periods. One may wonder whether we are over-interpreting mere fluctuations of the periodograms, 
-%while the peaks seen in the literature from analysis of the light curve are unmistakable. It is therefore worth to take advantage of the 
-%availability of the AAVSO data to perform a Lomb-Scargle periodogram of the light curve in the same period covered by our polarization data and,
-%furthermore, taking into consideration data around the dates for which TBL data is available. Periodograms  with such constraints are 
-%shown in Fig.\ref{LSAAVSO}. Even when the full dataset for the period is taken into account (red line), the periods present small amplitudes, and are 
-%fully and favourably comparable with the periodograms presented above. This is even more dramatic when the periodogram is made from data at the exclusive 
-%dates at which TBL data is available (blue line). It stands out that the low amplitude peaks are rather due to both the sparse cadence and short span of 
-%time of the TBL data series rather than to a physical absence. Actually, the present tests puts more emphasis on the importance of the 
-%strong peaks seen at selected wavelengths in the LSD profiles, strong when compared to the peaks at similar periods arising from the light curve.
-
-%\begin{figure}[!h]
-%    \centering
-%    \includegraphics[width=\linewidth]{Lomb-Scargle-TBL.png}
-%    \caption{Lomb-Scargle periodogram from the AAVSO data between 2013 and 2023 (red line) and from the AAVSO data taken near TBL observations (blue line).}
-%    \label{LSAAVSO}
-%\end{figure}
-
-% constant, which is not consistent with a pulsating scenario. The constant magnitude of Betelgeuse points toward another scenario than pulsation. Using surface 
-% convection, we can fully explain the variability of Betelgeuse. The oscillations in the light curve are due to two mechanisms: surface convection and rising 
-% plumes of plasma beyond the limb of the star. This last mechanism being random, it happens from time to time. The timescale on which convective cells evolve
-% is between one and two years, which is also consistent with the 400 days period found in the literature.  After the dimming, the change of periodicity in the
-% light curve is explained by surface convection, the random variation of brightness distribution could lead to the 200 days periodicity found in the literature. 
-% The constant brightness of Betelgeuse for the last 6 months is explainable by the same random brightness distribution: the surface of Betelgeuse is calm, 
-% leading to this constant brightness. In the future, we expect to see the light curve of Betelgeuse to change, maybe with another main period, but which 
-% is in fact caused by surface convection. Thus, the main mechanism in the variability of Betelgeuse is surface convection, which is sufficient to explain the 
-% variation in the magnitude of Betelgeuse.
-
-%and not to radial or 
-%other global pulsations.
-
-% We recall that while the nature of the LSP is still in debate, the nature of the lower periods is quite commonly 
-% attributed to  pressure modes 
-% (e.g \cite{kiss_variability_2006}). The mechanism behind the pulsation of Betelgeuse is attributed to the $\kappa$-mechanism. In this 
-% section, we attempt to 
-% find an explanation to the 400 days periodicity of Betelgeuse. 
-% As mentioned before, the LS periodogram of linear polarization shows a periodicity 
-% around 400 days. Furthermore, the region where the LS power is high is around 40km/s. 
-% This wavelength corresponds to most redshifted signal present in Betelgeuse.
-
-%The 400-day period appears prominently in both the intensity and the total polarization profiles. And in both cases, it is mostly present 
-%at wavelengths in the red wing, often beyond the 40 km/s value.
-%We recall that in the present scenario for the formation of the polarization signal, this signal is supposed to remain between 
-%-20 and +40km/s. Signals above 40km/s are uncommon and can be associated to
- %plasma rising behind the limb of Betelgeuse. 
-%Such scenario has been observed in the RSG $\mu$ Cep and was introduced by \cite{lopez_ariste_height_2023} to explain the excess of 
-%linear polarization beyond  the rest velocity of the star, which for Betelgeuse is fixed at +40km/s. 
-%In $\mu$ Cep, such  linear polarization signal beyond the limit of the rest velocity of the star is very strong and particularly 
-%common, two features that prompted a dedicated explanation. In Betelgeuse, such signals, when present, are weaker. However, 
-%they are visible from time to time, with regularities that are captured by  the Lomb-Scargle periodogram at 400 days.
-%Without further justification, we may propose that convective plumes in the hidden hemisphere of the star and 
-%rising above the limb from time to time are responsible for the  400 days period in the light curve. When such plumes rise high enough 
-%to be geometrically visible above the limb they show signals in intensity and polarization at redshifted wavelengths. They also
- %increase the emitting surface of the star and  the number of photons that we receive,
-%leading to an increase of the brightness visible in the light curve. After a few weeks or months, the plume falls back to Betelgeuse, 
-%disappearing from our point of view,  leading to a decrease of the brightness.
-
-%The appearance of such high-rising plumes is not necessarily periodic. And the star can easily shift from active periods when such 
-%plumes are a common feature to calmer periods when no plume is visible for months. $\mu$ Cep may be in one such active period presently,
-%while Betelgeuse may have experienced recently one great such event in the direction of the Earth that, after dust formed, became the great 
-%dimming, but since then no other high-rising plume has appeared, explaining the absence of the 400-day period since the great dimming. 
-%From the point of view of convective dynamics we can offer an explanation of the 400-day period which not only explains the observation of 
-%this period both in the light curve and in the polarization profiles, but also its presence at redshifted wavelengths and its present (and probably 
-%temporary) disappearance since the great dimming.
-
-% This is exactly what we see in the LSD profile of the linear polarization of Betelgeuse: 
-% from time to time, there is a net signal above 40km/s, which is present for a few weeks, and then disappears. Since this signal is not always present, 
-% it is found by the LS periodogram as a periodic event. However, it is not periodic at all: because the plumes rise at random moments, sometimes they 
-% appear every years, sometime they do not. This would explain why we only see it before the dimming. After the dimming, this kind of event never happened 
-% for the moment. Thus, the LS periodogram, at that precise frequency, finds that there is a periodic signal before the dimming, but not after, leading to
-% a lower LS power at this frequency.
-
-% To summarise what we have done so far: we were able to explain the variability of Betelgeuse using surface convection only. The convective cells play 
-% the most important role in the variability of Betelgeuse. Our work does not need pulsation to explain its variability. This does not mean that there are 
-% no pulsation in Betelgeuse. What we claim is that the most important factor in the variability of Betelgeuse is the timescale on which the convective 
-% cells evolve. Pulsation could exist on Betelgeuse, but they are not the main contributor to the variability.
 
@@
 The main result presented in this work is that the same periods identified in the light curve of Betelgeuse are also observed in the intensity and polarization
-spectra measured with the TBL. Traditionally, periods in the light curve have been associated to pulsations due to pressure modes.
-%But such pulsation would in no case affect the polarisation signals nor would explain the fact that the periods can be observed with stronger 
-%amplitude at selected wavelengths but no others. % Pulsation appears to offer no coherent explanation to these observations, are hardly to the 
-%observed changes in the variability since the great dimming.  The concurrence of observational facts, actual period values, evolution before 
-%and after the dimming, the larger amplitude of the period peaks in polarizarion data, points towards a common origin for those variabilities. 
-%And such origin cannot be pulsations.
+spectra measured with the TBL.
+Traditionally, periods in the light curve have been associated to pulsations due to pressure modes.
 However, the interpretation of the linear polarization profiles is made in terms of convective structures in the atmosphere of Betelgeuse.
@@
 We further find hints in the polarized spectra that lead us to associate the 200 d periodicity to convective dynamics, while the 400 d periodicty may be linked to convective timescales of the largest granules.
-%the observed periods are related 
-%to convective dynamics and not to radial pulsations of the star.
 
@@
 series of spectropolarimetric observations of Betelgeuse will be necessary to provide further insights into this phenomenon.
-
-%Although of secondary importance, we may suggest with more or less ease what phenomena are at the origin of the main periods. The 400-day period 
-%appears to be strongly related to the appearance of plumes in the back hemisphere of the star rising high enough to be seen above the limb.
-%Such plumes have been described in another RSG, $\mu$ Cep, and while less common or prominent, they are also present in Betelgeuse. The irregularity 
-%of their presence justifies the 400-day period, which is captured by the Lomb-Scargle periodogram in the polarization signals attributed to 
-%this phenomenon. The redshifted wavelengths at which this periodicity is found are justified by the plume moving away of the star and of the Earth while 
-%raising in the back hemisphere, while the brightness increase is justified by %the extra surface of emitting plasma visible. The 200-day period is less 
-%prominent but appears to be correlated to the typical time scales of the evolution of the convective structures themselves on the visible hemisphere. 
-%The LSP origin is harder to pinpoint given the insufficient time length of our time series. One can speculate if it may be related to the 
-%longest living photospheric structures, but this explanation does not adequately explain all the features associated to this period. Longer time 
-%series of spectropolarimetric observations of Betelgeuse will be required to shed light on this.
 

### 2024-09-04 answerreferee.tex
Add files via upload

\documentclass{article}
\usepackage{graphicx} % Required for inserting images

\begin{document}

We would like to thank the anonymous referee for the comments. Every changes in the article were marked by using boldface. \\

\textbf{Major comments:}\\

Q: Given that this paper presents an analysis of a polarimetric dataset to try to find periodic
behaviors in the data, I feel that the introduction section of the paper would benefit from a more
extended discussion of the periodicities found in Betelgeuse that currently exists in the
literature, including uncertainties on those numbers. Currently some of this information is
scattered throughout the paper and not included in the introduction, but it is important for the
reader to have a thorough understanding of what the current state of our understanding of its
periodicities are so that they can put the results of this paper in context.\\

A: We added extra information in the introduction about the different periods found in the literature and their uncertainties. Especially, we did mention the uncertainties found by Jadlovsky et al.2023 and Joyce et al.2020. We also added another reference to Montargès et al.2016 and Chiavassa et al. 2022.\\

Q: I feel that section 2 of the paper would benefit from more discussion on the details of the
dataset. I know that the data have been previously published in Aurière et al. (2016), Mathias et
al. (2018), and López Ariste et al. (2022) and each of those papers includes both a table of the
data they used and the details of the observations, but I think a brief summary of those details in
this paper would help it better stand alone for readers attempting to understand the details of
this work. For example, I think it would be helpful to include information like the total number of
observations used (84 observations are in all three papers, but the paper does not state exactly
what data was cut out of this analysis so that only data from before the great dimming was
used), and how they are divided up between Narval and Neo-Narval (28 of the 84 observations
are from Neo-Narval, but again, it s not easily clear how much of that was not included).
Additionally, I think including a date/timeframe for when the switch between Narval and
Neo-Narval happened is important as changes in instrumentation are important. I also think this
is an appropriate place to discuss how often the data were taken on average, as this information
affects the periods that this dataset would be sensitive to, as well as including the range of dates
over which this dataset was taken (for example, saying something like the data were taken
between November 2013 and December 2019 ). In general, I feel that saying Betelgeuse has
been observed for the last 10 years is vague and if someone is reading this a few years in the
future, they may easily mistake the time period over which the data was taken.\\

A: We added the total number of observations that we used to compute the periodograms. There is a total of 56 observations, we precised that 43\% of the observations were taken between September and December while the other 57\% are taken between January and April. Only Narval data were used in this study, since Neo-Narval was operational after the Great Dimming of Betelgeuse.\\

Q: Why was a Lomb-Scargle periodogram not computed for the position angle of the polarization
and shown? I understand that this information is really already tied into the Q and plots, but
so is the \% polarization. In general, I feel that it s best to include \% polarization and position angle
in papers using polarimetric data because that s easier for people not used to working with
Stokes vectors to understand and the combination of the two is the equivalent of showing all of the polarimetric data (just showing \% polarization doesn’t give the full polarimetric details of the
Stokes parameters). If the reason it wasn’t included is because there were no periods found in
the position angle, then I think including it is actually even more important. Section 4, paragraph
2 of the paper states, Once again, we observe the 330 d periodicity to be more prominent in
Stokes than in Q, what may unravel a random situation due to convective motions in the last
years. Convective cells could arise in peculiar positions, where Stokes profile remains
unchanged for month but where Stokes Q could change on shorter timescales. I believe a
periodogram of position angle might help support this statement.\\

A: \\

Q: How was the relative scaling between each 1.8 km/s step maintained in the periodograms to
produce figures 1 through 4?\\

A: For a given periodogram, we computed the highest power of each velocity bin. Then, every other power was compared to this highest value. The closer the power is to the highest power, the redder it is.  \\

Q: In several places, starting in Section 3 and going through the end of the paper, the manuscript
mentions the window function of the Lomb-Scargle periodograms. I think there needs to be a
brief explanation on what the window function is and how that can be used in the interpretation
of the results because only people who are experts in Lomb-Scargle analysis are likely know
what this is.\\

A: We added the following sentence to be sure that readers that are not experts in LS analysis understand the window function "The spectral window reﬂects the pattern caused by the structure of gaps in the time string. The peaks pointed out by this window cannot be interpreted as ones related to the star."\\

Q: In section 3.2, a Lomb-Scargle periodogram is computed from the photo-center displacements
from polarimetric reconstructed imaging that was presented in López Ariste et al. (2022). The
details of how the photo-center displacements are calculated need to be stated.\\

A: The reference of how the photo-center was computed was added, it was computed using the equations of Chiavassa et al 2022.\\

Q: At first glance, Section 3.3 seems a bit redundant because the paper already is doing a
Lomb-Scargle analysis of the Stokes I parameter. Because Stokes I is total light, it is the light
curve of Betelgeuse. But because the AAVSO light curve and the Stokes I data should be
identical, I think a comparison between the two is important because it could validate the results
of the paper. This section of the paper states that the 400 d and 200 d periods are only
recovered when the full AAVSO light curve is used, but when the AAVSO light curve is reduced
to just the nights where TBL data were obtained we fail to retrieve the periods mentioned in the
literature . However, the Stokes I periodogram does recover the 400 d and 200 d periods. There
is very little discussion about what may cause this difference. Given that they are both
periodograms derived from the total light of the system on the same days, I would have
expected the periodogram of the AAVSO light curve that was trimmed to just include
observations on days the TBL data were taken to match the results from the Stokes I
periodogram. The fact that they do not produce the same periodicities deserves more
discussion than is currently present in the paper. What could be causing this? What does that
mean for the interpretation of the polarimetric results? The paper then goes on to say We can
anticipate that extended polarimetric observations of Betelgeuse will make these periodicities
seen in linear polarization. Due to convective structures, these periodicities will be clearer and
clearer. I don t understand these two sentences because the 200 d and 400 d periods are
already seen in the polarimetric data (there is a ~330 day periodicity in Stokes and
\% polarization, and what appears to be a ~200 day period in Stokes Q and \% polarization). More
observations will likely make their detection levels stronger, but that is unlikely to fix the reason
the Stokes I periodogram isn’t currently reproducing the AAVSO light curve periodogram.\\

A: We computed the periodogram where AAVSO observations correspond to the TBL observation dates. We sum up it in figure 7, where the upper panel represents the average periodogram of Stokes I, the middle panel the periodogram of the AAVSO since 1990 and the lower panel is the periodogram computed from AAVSO observations that correspond to TBL observation dates. We see that the same periods are present in the periodogram from Stokes I and the one where the AAVSO matches TBL observations. Hence, this confirms our results. We mentioned that in both cases, the 400 d period is close to a peak of the window function while the 200 d period is not. \\

\textbf{Minor comments:}\\

Q: Section 1:\\
In the last paragraph of section 1, Lomb-Scargle is misspelled the second time it is used. (The
misspelling is in the sentence that starts In section 3, we seek )\\

A: Corrected.\\

Q: In the last paragraph of section 1 of the paper (line 99 of the manuscript in the reviewer’s copy), the abbreviation LSD is used before the reader is told what it stands for (currently that s in the first paragraph of section 2, lines 122 and 123). I suggest flipping these so what LSD stands for
is clarified the first time it is mentioned.\\

A : We flipped the two sentences.\\

Q: Section 2:\\
In the first sentence of section 2, the acute accent over the first e in the word Télescope is missing.\\

A: Corrected.\\

Q: From López Ariste et al. (2022), I understand that the upgrade between Narval and Neo-Narval
happened during Betelgeuse s great dimming. Therefore, was any neo-Narval data actually
included in the Lomb-Scargle analysis shown? If not, then maybe Neo-Narval shouldn t even be
mentioned in section 2 of the paper.\\

A: We only used Narval data in our paper, it was clarified in the text, at the beginning of Section 3. \\

Q: The total number of lines that go into the LSD profile of each observation should be mentioned,
including the details of how lines are chosen to be included. (e.g. the details of the mask, cut off
for the central depth of the lines compared to the continuum for them to be included, etc.)\\

A: We added information about the mask to produce the LSD profile, those information can also be found in Mathias et al 2018 and Aurière et al 2016.\\

Q: Section 3:\\
How was the Lomb-Scargle analysis performed? Was it using the Lomb-Scargle class in astropy
or some other fitting software (such as Period04)?\\

A: We used the Lomb-Scargle class from Astropy and added a reference in the text.\\

Q: "Sometimes, signals at velocities greater than +40 km/s can be found..." is mentioned in the first paragraph of section 3. How much greater than +40 km/s are they? Since the explanation for
signals greater than this velocity is convective processes, is +60 km/s enough to still include
them? If so, maybe just stating that the +60 km/s cut off was used to ensure you always include
the occasional +40 km/s signal is appropriate.\\

A: The peaks in the linear polarization signal do not exceed +50km/s, so the cut off at +60km/s is to add the occasional signal above +40km/s. We precised it in the paper. \\

Q: In the second paragraph of Section 3.1, the paper states that the upper panel depicts the
Lomb-Scargle periodogram of each velocity bin (black line)... , but I believe that this should be
grey lines.\\

A: Thank you for the precision. \\

Q: Figures 1 through 4 all have HRV listed on the y-axis of the bottom plots, but this abbreviation is not defined anywhere. Perhaps it could be stated that it means heliocentric radial velocity in the
caption of figure 1. Additionally, all four plots have the individual periodograms in the top panels,
but there is no y-axis label. Is that axis power on those panels? If so, I believe it should be labeled. Similarly, the color axis in the bottom panels is not explained. (e.g. Red shows a higher power than dark blue).\\

A: The abbreviation of HRV was included in the legend of Figure 1, as well as the signification of the color axis. The label of the y-axis (power) was added in the figures.\\

Q: Line 196 (Section 3.1 paragraph 4), states "...it appears that the primary period is approximately 0.003 d\^-1 (equivalent to 330 d). I believe it should read ...it appears that the primary frequency is approximately 0.003 d\^-1 (equivalent to a 330 d period)." Section 4, paragraph 2 of the paper states, "Once again, we observe the 330 d periodicity to be more prominent in Stokes than in Q, what may unravel a random situation due to convective motions in the last years. Convective cells could arise in peculiar positions, where Stokes profile remains unchanged for month but where Stokes Q could change on shorter timescales." I think the authors mean Once again, we observe the 330 d periodicity to be more prominent in Stokes than in Q, which may unravel a random situation due to convective motions in the last years. Convective cells could arise in peculiar positions, where Stokes profile remains unchanged for months but where Stokes Q could change on shorter timescales.\\

A: We corrected the sentences. \\

Q: Why is a confidence level given for only one periodogram? How was it determined? In my
understanding of Lomb-Scargle analysis, confidence levels aren’t very reliable.\\

A: We computed the confidence level by computing the standard deviation of the mean value of each 100 periodogram. Since the images that we produce from polarimetry imaging are ambiguous, we performed a statistical analysis to see that the variability found in the photo-center displacement was due to statistical reasons, up to 1 sigma (not 2 as mentioned in the text). Concerning the other periodograms, we computed the LS for each velocity bin and we wanted to see if a variability shows up or not. But there is no reason to compute the confidence level in the observations since we are not performing a statistical analysis on the observations.

\end{document}

### 2024-09-09 answerreferee.tex
Add files via upload

@@
 
-Q: Given that this paper presents an analysis of a polarimetric dataset to try to find periodic
+Given that this paper presents an analysis of a polarimetric dataset to try to find periodic
 behaviors in the data, I feel that the introduction section of the paper would benefit from a more
@@
 A: We added extra information in the introduction about the different periods found in the literature and their uncertainties.
-Especially, we did mention the uncertainties found by Jadlovsky et al.2023 and Joyce et al.2020.
-We also added another reference to Montargès et al.2016 and Chiavassa et al.
-2022.\\
+Especially, we did mention the uncertainties found by Jadlovsky et al.
+(2023) and Joyce et al.
+(2020).
+We also added another reference to Montargès et al.
+(2016) and Chiavassa et al.
+(2022).\\
 
-Q: I feel that section 2 of the paper would benefit from more discussion on the details of the
+I feel that section 2 of the paper would benefit from more discussion on the details of the
 dataset. I know that the data have been previously published in Aurière et al. (2016), Mathias et
@@
 
-Q: Why was a Lomb-Scargle periodogram not computed for the position angle of the polarization
+Why was a Lomb-Scargle periodogram not computed for the position angle of the polarization
 and shown? I understand that this information is really already tied into the Q and plots, but
@@
 
-A: \\
+A: Contrary to the other questions and suggestions made by the referee, we mostly disagree in this point with the remark of the referee on the interest of showing the results for the polar angle.
+And this disagreement requires a longer response than the other questions.
+For a starter we strongly disagree in that giving a degree of polarization and a polar angle is a good thing.\\
+The purpose of the referee was to point out that showing Q/U  may reveal information not apparent in the separate Q and U various plots.
+We must not delay ourselves in the difficulty of numerically analysing a ratio Q/U where U can be zero, but rather insist on the possible interpretations of such ratio.
+The referee recalls a sentence of the present manuscript where we suggest a possible connection between the differences in Q and U and a possible preference of the position angle of convective patterns on the disk of Betelgeuse during the periods of observation.
+This rises the doubt on whether by polar angle the referee meant all the time the actual polar angle of particular features inferred on the disk of Betelgeuse rather than the angle of the polarization ellipse.
+Whether this is the actual intended meaning of "polar angle" or not, the response is the same: this is a derived quantity so, as the referee points out, there is no further information in it that is not in Q and U separately.
+And whatever this information is, there is no other manner of interpreting it than in terms of the azimuthal distribution of features over the disk.
+But such distribution is tributary of not just this angle but of the whole brightness distribution model used in the inversion algorithm and it is risky to interpret any periodicities in Q/U as azimuth anisotropies.
+We agree that in such case our own sentence is perhaps too speculative and ambiguous.
+Perhaps it should be retired.
+But we thought that 1) it is important to mention that we are aware that Q and U show differences and that these differences could be interpreted as azimuth anisotropies, but 2) this is not the purpose of the present paper.\\
+It is the purpose of the present paper to show that the very same periodicities that have been historically interpreted as oscillations or pulsations can be re-interpreted as due to the convection dynamics alone, and that indeed this is the most appropriate explanation since the very same periodicities are seen in observables unrelated to pulsation or oscillatory phenomena as are the Stokes parameters.
+Whether there is further information on the convection of Betelgeuse hidden in those peridocities is beyond our present purpose, but that would be the only interest on comparing Q and U periodicities, through the ratio Q/U or else, to further explore such dynamics.
+This is not our purpose now, however interesting.\\
 
-Q: How was the relative scaling between each 1.8 km/s step maintained in the periodograms to
+How was the relative scaling between each 1.8 km/s step maintained in the periodograms to
 produce figures 1 through 4?\\
@@
 
-Q: In several places, starting in Section 3 and going through the end of the paper, the manuscript
+In several places, starting in Section 3 and going through the end of the paper, the manuscript
 mentions the window function of the Lomb-Scargle periodograms. I think there needs to be a
@@
 
-Q: In section 3.2, a Lomb-Scargle periodogram is computed from the photo-center displacements
+In section 3.2, a Lomb-Scargle periodogram is computed from the photo-center displacements
 from polarimetric reconstructed imaging that was presented in López Ariste et al. (2022). The
@@
 
-A: The reference of how the photo-center was computed was added, it was computed using the equations of Chiavassa et al 2022.\\
+A: The reference of how the photo-center was computed was added, it was computed using the equations 2 and 3 from Chiavassa et al. (2022).\\
 
-Q: At first glance, Section 3.3 seems a bit redundant because the paper already is doing a
+At first glance, Section 3.3 seems a bit redundant because the paper already is doing a
 Lomb-Scargle analysis of the Stokes I parameter. Because Stokes I is total light, it is the light
@@
 
-Q: Section 1:\\
+Section 1:\\
 In the last paragraph of section 1, Lomb-Scargle is misspelled the second time it is used. (The
-misspelling is in the sentence that starts In section 3, we seek )\\
+misspelling is in the sentence that starts "In section 3, we seek..." )\\
 
@@
 
-Q: In the last paragraph of section 1 of the paper (line 99 of the manuscript in the reviewer’s copy), the abbreviation LSD is used before the reader is told what it stands for (currently that s in the first paragraph of section 2, lines 122 and 123).
+In the last paragraph of section 1 of the paper (line 99 of the manuscript in the reviewer’s copy), the abbreviation LSD is used before the reader is told what it stands for (currently that s in the first paragraph of section 2, lines 122 and 123).
 I suggest flipping these so what LSD stands for
@@
 
-A : We flipped the two sentences.\\
+A: We flipped the two sentences.\\
 
-Q: Section 2:\\
+Section 2:\\
 In the first sentence of section 2, the acute accent over the first e in the word Télescope is missing.\\
@@
 
-Q: From López Ariste et al. (2022), I understand that the upgrade between Narval and Neo-Narval
+From López Ariste et al. (2022), I understand that the upgrade between Narval and Neo-Narval
 happened during Betelgeuse s great dimming. Therefore, was any neo-Narval data actually
@@
 
-Q: The total number of lines that go into the LSD profile of each observation should be mentioned,
+The total number of lines that go into the LSD profile of each observation should be mentioned,
 including the details of how lines are chosen to be included. (e.g. the details of the mask, cut off
@@
 
-A: We added information about the mask to produce the LSD profile, those information can also be found in Mathias et al 2018 and Aurière et al 2016.\\
+A: We added information about the mask to produce the LSD profile, those information can also be found in Mathias et al. (2018) and Aurière et al. (2016).\\
 
-Q: Section 3:\\
+Section 3:\\
 How was the Lomb-Scargle analysis performed? Was it using the Lomb-Scargle class in astropy
@@
 
-Q: "Sometimes, signals at velocities greater than +40 km/s can be found..." is mentioned in the first paragraph of section 3. How much greater than +40 km/s are they? Since the explanation for
+"Sometimes, signals at velocities greater than +40 km/s can be found..." is mentioned in the first paragraph of section 3. How much greater than +40 km/s are they? Since the explanation for
 signals greater than this velocity is convective processes, is +60 km/s enough to still include
@@
 
-Q: In the second paragraph of Section 3.1, the paper states that the upper panel depicts the
+In the second paragraph of Section 3.1, the paper states that the upper panel depicts the
 Lomb-Scargle periodogram of each velocity bin (black line)... , but I believe that this should be
@@
 
-Q: Figures 1 through 4 all have HRV listed on the y-axis of the bottom plots, but this abbreviation is not defined anywhere.
-Perhaps it could be stated that it means heliocentric radial velocity in the
+Figures 1 through 4 all have HRV listed on the y-axis of the bottom plots, but this abbreviation is not defined anywhere. Perhaps it could be stated that it means heliocentric radial velocity in the
 caption of figure 1. Additionally, all four plots have the individual periodograms in the top panels,
@@
 
-Q: Line 196 (Section 3.1 paragraph 4), states "...it appears that the primary period is approximately 0.003 d\^-1 (equivalent to 330 d).
+Line 196 (Section 3.1 paragraph 4), states "...it appears that the primary period is approximately 0.003 d\^-1 (equivalent to 330 d).
 I believe it should read ...it appears that the primary frequency is approximately 0.003 d\^-1 (equivalent to a 330 d period)." Section 4, paragraph 2 of the paper states, "Once again, we observe the 330 d periodicity to be more prominent in Stokes than in Q, what may unravel a random situation due to convective motions in the last years.
@@
 
-Q: Why is a confidence level given for only one periodogram? How was it determined? In my
+Why is a confidence level given for only one periodogram? How was it determined? In my
 understanding of Lomb-Scargle analysis, confidence levels aren’t very reliable.\\
 
-A: We computed the confidence level by computing the standard deviation of the mean value of each 100 periodogram.
+A: We computed the confidence level by computing the standard deviation of the mean value of the 100 periodograms (black line Fig.5).
 Since the images that we produce from polarimetry imaging are ambiguous, we performed a statistical analysis to see that the variability found in the photo-center displacement was due to statistical reasons, up to 1 sigma (not 2 as mentioned in the text).

### 2024-09-09 main.tex
Add files via upload

@@
 Attempting to identify periods in an astrophysical context faces the challenge of unevenly spaced observed data. To overcome this hurdle, 
-we used the Lomb-Scargle periodogram \citep{lomb_least-squares_1976,scargle_studies_1982}.
-This method involves fitting sinusoids to the data using least-squares with frequency sampled between the first and last observation \textbf{\citep{the_astropy_collaboration_astropy_2018}}.
+we used the Lomb-Scargle periodogram \citep{lomb_least-squares_1976,scargle_studies_1982} \textbf{as implemented in astropy \citep{the_astropy_collaboration_astropy_2018}}.
+This method involves fitting sinusoids to the data using least-squares with frequency sampled between the first and last observation.
 The quality of the fit determines the power attributed to a frequency.
-We applied this technique to the  LSD profiles of Stokes $I$, $U$, $Q$ and to the total linear polarization of Betelgeuse observed by the TBL over a span of 10 years (\textbf{2013}-2024).
-However, the dataset available post-dimming is insufficient to produce meaningful periodograms.
-Therefore, our focus was on periodograns computed exclusively from data obtained before the great dimming, covering the period from \textbf{ November 2013} to \textbf{April} 2019.
-\textbf{We only used data obtained from Narval, with a total of 56 observations}.
+\textbf{We applied this technique to the LSD profiles of Stokes $I$, $Q$, $U$ and to the total linear polarization of Betelgeuse observed by the TBL with Narval from 2013 to 2019, with a total of 56 observations.
+We also looked at the periodograms obtained from the dataset available post-dimming, from 2020 to 2024.
+However, this dataset is insufficient to produce meaningful periodograms.
+Therefore, our focus was on periodograms computed exclusively from data obtained before the great dimming with Narval.}
 
@@
 From these images, we computed a photo-center, a quantity sensitive to the size and number of convective cells.
-\textbf{The photo-center was computed from the equations given by \cite{chiavassa_probing_2022}.}
+\textbf{The photo-center was computed from the equations 2 and 3 given by \cite{chiavassa_probing_2022}.}
 A homogeneous star will have a photo-center displacement coinciding with the barycenter of the star, whereas a star with one or two large convective cells will exhibit a more siginificant photo-center displacement, up to a few percent of the stellar radius in the case of RSGs \citep{chiavassa_probing_2022}.
